\chapter*{The nineteenth-century saga, according to BL Add. 11109}
\addcontentsline{toc}{chapter}{The nineteenth-century saga}
Manuscripts used for this edition: 
\begin{itemize}[noitemsep]
	\item BL Add. 11109 (B11109) - base text
	\item Lbs 679 4to (L679)
	\item Lbs 1572 4to (L1572)
	\item Lbs 2404 8vo (L2404)
\end{itemize}
\newpage
\beginnumbering \pstart \ledouternote{106v} 
\section*{Sagan af  Hrómundi Greipssyni.}
\subsection*{I kap\textit{ítuli}}
 Sva hefr Søgo þ\textit{ess}a, a\textit{t} á þ\textit{ei}m tímu\textit{m} s\textit{em} \edtext{m\textit{ar}g\textit{ir} Stólk\textit{on}g\textit{ar}}{\Afootnote{fylkis kongar  L1572.}} v\textit{or}o í Norvegi, Dan\-m\textit{or}co o\textit{c} Svíþjóþ, g\textit{jỏr}þost m\textit{ar}g\textit{ir} \edtext{Hỏfþíngj\textit{ar}}{\Afootnote{÷  L2404, L1572.}} \edtext{o\textit{c}}{\Afootnote{÷  L679, L2404, L1572.}}  st\textit{or}\-mektug\textit{ir}  \edtext{Herr\textit{ar}}{\Afootnote{÷  L679.}}, Greif\textit{ar} oc J\textit{ar}l\textit{ar}, s\textit{em}  bruto   s\textit{ic}
t\textit{il} r\textit{ic}ja o\textit{c} landa,   h\textit{er}juþo   \edtext{sva v\textit{el}}{\Afootnote{jafnvel  L2404.}} á vetr\textit{om} \edtext{s\textit{em}}{\Afootnote{og  L1572.}} sumr\textit{om}, 
o\textit{c} ỏbluþo sér sva fjár o\textit{c} \edtext{frægþa}{\Afootnote{frægdar  L679, L1572.}}.  
A m\textit{ed}al þ\textit{ess}ara  \edtext{o\textit{c} þ\textit{ví}líkra}{\Afootnote{÷  L2404, L1572.}} v\textit{ar} ein\textit{n} m\textit{ic}ils  \edtext{v\textit{er}þr}{\Afootnote{háttar  L679, L2404.}} K\textit{ón}gr,  \edtext{s\textit{em}}{\Afootnote{er  L679, L1572.}} Olafr h\textit{ie}t, h\textit{ann} v\textit{ar} \edtext{ólatasti}{\Afootnote{hinn ólatasti  L2404, L1572.}} t\textit{il} b\textit{ar}daga,  
ỏr af fé v\textit{it} vini sína, en\textit{n} \edtext{str\textit{an}gr o\textit{c}}{\Afootnote{÷  L1572.}} \edtext{hefnigj\textit{ar}n}{\Afootnote{hefndargiarn  L1572.}}
v\textit{it} Ovini; 
h\textit{ann}  r\textit{ac} af  \edtext{r\textit{ic}j\textit{om} }{\Afootnote{riki L1572.}} m\textit{ar}ga K\textit{on}ga o\textit{c} hỏfþíngja, o\textit{c} hafþi  vetrseto  \edtext{í Norvegi,  sérdeilis}{\Afootnote{÷  L2404.}} á Háloga l\textit{an}di, þ\textit{ar} sat K\textit{ón}gr optast.  
\edtext{K\textit{óngr}}{\Afootnote{hann  L1572.}} átti ij systr, s\textit{em} fram\textit{m} t\textit{óc}o  \edtext{ỏll\textit{om}}{\Afootnote{÷  L679.}} ỏþr\textit{om} mey\textit{om}, a\textit{t} kost\textit{om}
o\textit{c}  qven\textit{n}príþi,   hyggind\textit{om}  \edtext{s\textit{em}}{\Afootnote{og  L679, L1572.}} \edtext{h\textit{an}dvirþ\textit{om}}{\Afootnote{hamingu  L679, handyrdum  L1572, handiþnum  L2404.}},\footnote{
	An example of an archaic word \textit{handvirði} (neuter noun), registered in OH from the 17th cent.; the variants \textit{handiðn} (feminine noun) registered from 18th-20th cent., and \textit{handyrð} (feminine noun) registered from 17th-19th cent. In Blöndal's dictionary \textit{handvyrðir} is a synonym of \textit{hannyrðir}, which is translated as female handicraft (Da. \textit{kvindeligt håndarbejde}).} 
á millum Vícr
o\textit{c} Hálogal\textit{an}ds, h\textit{ie}t ỏn\textit{n}r Dagný
e\textit{n} ỏn\textit{n}r Svanhvít;  \edtext{Svanh\textit{vít}}{\Afootnote{hún  L679.}} b\textit{ar} af systr sin\textit{n}i  \edtext{of}{\Afootnote{um  L679, L2404, L1572.}}\footnote{An archaic use of the preposition \textit{of} with acc., which has the same meaning as \textit{um}.}  flesta luti, þó v\textit{or}o báþ\textit{ar} hin\textit{ar} \edtext{tignarl\textit{e}gosto}{\Afootnote{tig[a]glegustu L679, görfilegustu  L2404.}} \edtext{Jómfrúr}{\Afootnote{meýar  L2404.}}. Tveir b\textit{ræþ}r v\textit{or}o m\textit{eþ} K\textit{óngi} er
h\textit{ann} h\textit{afþi} mest\textit{ar}  mætr  á,  \edtext{o\textit{c} h\textit{ie}t ein\textit{n}}{\Afootnote{hiet annar  L679, L1572.}} Bíldr, e\textit{n} an\textit{n}\textit{ar} Vole, þ\textit{ei}r v\textit{or}o full\textit{ir} \edtext{fjỏlcíngis}{\Afootnote{fjólkingis og  L679, L2404.}}, galdra o\textit{c} \edtext{g\textit{jỏr}nínga}{\Afootnote{÷  L2404.}},  \edtext{íllsko}{\Afootnote{÷  L2404, L1572.}} \edtext{o\textit{c} und\textit{ir}f\textit{ỏr}ulh\textit{ei}ta}{\Afootnote{÷  L1572, og undirfỏrulir  L2404.}};  \edtext{þ\textit{ei}m  q\textit{vo}mo}{\Afootnote{komu þeim  L1572.}} fá\textit{ir}  lut\textit{ir}   óv\textit{or}om, o\textit{c} íllt lỏgþo
þ\textit{ei}r t\textit{il} fleyrstra mála.  Maþr n\textit{oc}r bjó í nálægþ K\textit{ón}gi, er Greipr
h\textit{ie}t, h\textit{ann} v\textit{ar}  auþugr  \edtext{a\textit{t}}{\Afootnote{af  L679, L1572.}} fé, o\textit{c} \edtext{hin\textit{n}  gỏfgasti  af bænd\textit{om}}{\Afootnote{gófugastr bondi  L679.}} \edtext{þ\textit{ar} of  pláts}{\Afootnote{÷  L2404, L1572, þar um plátss  L679.}}; h\textit{ann} v\textit{ar} qvæntr m\textit{aþr} \edtext{o\textit{c}}{\Afootnote{hann  L679.}} átti d\textit{óttur} Hróks hins Svarta, er 
 \edtext{Vallaug h\textit{ie}t}{\Afootnote{Vallaug hét /:al: Gunnlod :/ modir hennar Brinhildur dottir Haka konungs i Skáney  L1572, Vallaug (alii Gunnlỏd :) móþir hennar var Brynhildr dóttir Haca konungs á Skáneý  L2404.}}, 
hún  v\textit{ar} qven\textit{n}kostr e\textit{n} \edtext{betsti}{\Afootnote{vænsti  L679.}}, o\textit{c} v\textit{el} a\textit{t} sér  \edtext{of}{\Afootnote{um  L679, L2404, L1572.}}  m\textit{ar}ga luti.  Greipr  un\textit{n}i kono sin\textit{n}i m\textit{ic}it,
o\textit{c}  ól   v\textit{it} hen\textit{n}i ix suno;
h\textit{ie}t ein\textit{n} \edtext{Héþin\textit{n}}{\Afootnote{Haki  L1572.}},  an\textit{n}\textit{ar} \edtext{Haci}{\Afootnote{Hedinn  L1572.}}, þriþji Gautr, fjórþi Hỏgni, fim\textit{m}ti Þrỏstr,  sjỏtti
Angantýr, sjỏundi  \edtext{Logi}{\Afootnote{Lodinn  L679, L1572.}}, áttundi Helgi, níundi  Hrómundr, h\textit{ann} v\textit{ar}
eldstr þ\textit{eir}ra b\textit{ræþ}ra o\textit{c} hin\textit{n} frægasti m\textit{aþr}  at ||\ledouternote{107r} \edtext{hreysti o\textit{c} ỏll\textit{om} fræcleica}{\Afootnote{óllum jþrottum og hreisti  L679.}},  sva h\textit{ann} b\textit{ar} l\textit{an}gt af m\textit{ỏnnom} þ\textit{ar} í  \edtext{byggþ\textit{om}}{\Afootnote{sveitbigdum  L679.}}, þ\textit{ei}r b\textit{ræþ}r  \edtext{all\textit{ir} v\textit{or}o}{\Afootnote{voru  L679, voru allir  L1572.}} \edtext{af alþýþo kallaþ\textit{ir} Hróc\textit{ar}}{\Afootnote{kalladir þar i bigdum af óllum Hrokar  L679, af alþíþo Hrókar kallaþir  L2404.}}, þ\textit{ví} þeir  v\textit{or}o af Hrócs ætt q\textit{vo}mn\textit{ir}.  Hrómundr v\textit{ar} bj\textit{ar}tr á hár,   hýr   í t\textit{il}liti,  \edtext{e\textit{n}}{\Afootnote{og  L2404.}}  sn\textit{ar}  í aug\textit{om}, breiþr á  
h\textit{er}þ\textit{ar}  \edtext{o\textit{c}}{\Afootnote{en  L679.}} \edtext{st\textit{or}m\textit{enn}i}{\Afootnote{stór  L1572, mikill  L2404.}} \edtext{a\textit{t}}{\Afootnote{í  L679.}} vexti,  h\textit{ann} gaf s\textit{ic} alldrei f\textit{yrir}
und eþr  ákomo, o\textit{c}
v\textit{it} enga   æþru  
v\textit{ar} h\textit{ann} kéndr; þ\textit{ei}r b\textit{ræþ}r h\textit{an}s v\textit{or}o  \edtext{o\textit{c}}{\Afootnote{÷  L2404.}} all\textit{ir} \edtext{m\textit{ic}l\textit{ir} m\textit{enn}}{\Afootnote{mikilmenni  L679.}} t\textit{il} \edtext{allra m\textit{an}\textit{n}burþa}{\Afootnote{hreisti  L2404.}}.
\pend \endnumbering  

\beginnumbering \pstart  
\vspace{5mm}\subsection*{II kap\textit{ítuli}} 
Þat v\textit{ar} einusin\textit{n}i s\textit{em} \edtext{opt\textit{ar}}{\Afootnote{÷  L679.}}, at Olafr K\textit{ón}gr lætr blása t\textit{il} h\textit{er}f\textit{er}þ\textit{ar}, 
heimt\textit{ir}  sam\textit{an} liþ sitt,
o\textit{c} \edtext{segist}{\Afootnote{qvaþst L2404.}} vilja h\textit{afa} h\textit{er} úti,  
ept\textit{ir} v\textit{an}a sín\textit{om},  \edtext{a\textit{t}}{\Afootnote{og  L679.}} \edtext{ávin\textit{n}a}{\Afootnote{vinna  L679, [ava]  L2404.}} sér o\textit{c} kỏpp\textit{om} sín\textit{om} kost o\textit{c} uppheldi, 
sva  \edtext{frægþ af f\textit{ar}a megi}{\Afootnote{fara meigi frægd af  L679, frægd af fengist  L1572.}}, 
e\textit{n} óhróþr niþr leggjast, dreif nú t\textit{il} K\textit{ón}gs fjỏldi  \edtext{af fólki}{\Afootnote{÷  L679, fólks  L1572.}} or ýms\textit{om} \edtext{átt\textit{om}}{\Afootnote{lóndum  L679.}},  heldr  \edtext{þá}{\Afootnote{÷  L679, L1572.}} K\textit{óngr} þeg\textit{ar} á haf m\textit{eþ} m\textit{ỏr}g\textit{om}  t\textit{il}faung\textit{om}. 
M\textit{aþr} v\textit{ar} \edtext{sá}{\Afootnote{÷  L679.}} m\textit{eþ} Ol\textit{afi} K\textit{óngi} \edtext{s\textit{em}}{\Afootnote{er  L679, L2404, L1572.}} Kári h⟨ie⟩t, 
h\textit{ann} v\textit{ar} \edtext{ágjætis}{\Afootnote{ágjætr  L679.}} m\textit{aþr} a\textit{t} ỏll\textit{om} drengscap,
h\textit{ann} v\textit{ar} fóstbr\textit{óþir} Hrómund\textit{ar} o\textit{c}  ástfólgin\textit{n} vinr;
br\textit{óþir} h\textit{an}s  h\textit{ie}t Ørnólfr,  \edtext{o\textit{c}}{\Afootnote{hann  L679.}} v\textit{ar} m\textit{aþr} fræcin\textit{n}. 
Kári v\textit{ar} ramr  a\textit{t} abli, \edtext{han\textit{n}}{\Afootnote{og  L679, L1572.}} stóþ jafn\textit{an} í Kóngsins eyrind\textit{om},  \edtext{o\textit{c}}{\Afootnote{hann  L679.}} v\textit{ar} stafnbúi han\textit{n}s.  Nú s\textit{em} K\textit{óngr} v\textit{ar} á haf  sigldr, 
g\textit{jỏr}þost h\textit{ono}m  \edtext{óhagstæþ\textit{ir} vind\textit{ar}}{\Afootnote{óhagstædr vindr  L679, L1572.}},  
sva K\textit{óngr} \edtext{hlaut a\textit{t} halda austr}{\Afootnote{vard ad sig hallda  L679.}} t\textit{il} \edtext{Svíþjóþ\textit{ar}}{\Afootnote{Svijþjodar austr  L679.}}, o\textit{c} a\textit{t} eino qveldi sigldo þ\textit{ei}r und\textit{ir} Ey\textit{ar} þ\textit{ær} er  \edtext{Ulfaskér h\textit{ei}ta}{\Afootnote{heita Elfar skier  L679.}}, þ\textit{ar} létto  þ\textit{ei}r  \edtext{atkér\textit{om}}{\Afootnote{atkér síga  L679.}}, o\textit{c} t\textit{óc}o  náþ\textit{ir}. 
Þá m\textit{ælir} K\textit{óngr} v\textit{it} Kára:
nú skulot þit b\textit{ræþ}r  \edtext{g\textit{an}ga á l\textit{an}d}{\Afootnote{ganga hier á land L679, a land ganga  L1572.}} 
o\textit{c}  þv\textit{er}t   yf\textit{ir} þ\textit{ess}a eyo  \edtext{s\textit{em}}{\Afootnote{er  L679, L1572.}} v\textit{it} er\textit{om} \edtext{v\textit{it}}{\Afootnote{ad  L679, L1572.}} komn\textit{ir}, 
o\textit{c} vita hv\textit{or}t þ\textit{ér}
v\textit{er}þ\textit{it} e\textit{ck}i v\textit{ar}ir v\textit{it},  \edtext{a\textit{t}}{\Afootnote{÷  L679.}} \edtext{vícíng\textit{ar} nein\textit{ir} séo}{\Afootnote{neina vijkinga  L679, vijkingar seu  L1572.}} hino  megin\textit{n} Ey\textit{ar}in\textit{n}\textit{ar}.
Kári m\textit{ælir} sva skul\textit{om} v\textit{it} g\textit{jỏr}a s\textit{em} þér t\textit{il}mælit,  \edtext{Herra!}{\Afootnote{herra segir hann  L679, L2404, L1572.}} t\textit{ac}a þ\textit{ei}r b\textit{ræþ}r vapn sín o\textit{c}
g\textit{an}ga á l\textit{an}d  \edtext{upp}{\Afootnote{÷  L679.}} o\textit{c} yfir eyona, o\textit{c} \edtext{s\textit{em}}{\Afootnote{er  L679.}} þ\textit{ei}r q\textit{vo}mo á ham\textit{ar} n\textit{oc}orn, sjá þ\textit{ei}r  \edtext{vj}{\Afootnote{vij  L679.}} h\textit{er}scip ligg\textit{ia} und\textit{ir} Eyun\textit{n}i,  \edtext{m\textit{ic}it}{\Afootnote{mjóg  L679.}} st\textit{or} o\textit{c} skrautl\textit{e}g, 
þó b\textit{ar} \edtext{þ\textit{ar}}{\Afootnote{÷  L679.}} eitt af ỏll\textit{om}, þ\textit{at} v\textit{ar} dreci  \edtext{sva vegl\textit{e}gr}{\Afootnote{mikill og vænlegr  L679.}}, 
a\textit{t} Kári þóktist ei  \edtext{an\textit{n}\textit{an}}{\Afootnote{÷  L679.}} slíc\textit{an} sjeþ h\textit{afa}, 
||
\ledouternote{107v} 
h\textit{ann} v\textit{ar}  scygþr   o\textit{c}  scorin\textit{n}   stafna   á mill\textit{om}. 
Kári  \edtext{settist niþr o\textit{c}}{\Afootnote{÷  L679.}} qv\textit{aþ} m\textit{ar}g\textit{ar} vísr af ágæti  \edtext{scip\textit{an}a}{\Afootnote{skipsins  L679.}}; 
síþ\textit{an} kallaþi h\textit{ann} t\textit{il} þ\textit{eir}ra er \edtext{láo f\textit{yrir}}{\Afootnote{fijrir láu  L679.}}, 
o\textit{c} spyr  \edtext{hv\textit{or}j\textit{ir} v\textit{ær}o?}{\Afootnote{hvór fijrir væri  L679.}} sá nemndi s\textit{ic} er f\textit{yrir} þ\textit{ei}m  v\textit{ar} o\textit{c} \edtext{sagþist}{\Afootnote{kvadst  L679, L1572.}} Hraungviþr h\textit{ei}ta; 
Kári þóktist e\textit{ck}i  \edtext{sjeþ  h\textit{afa}}{\Afootnote{hafa sied  L679.}} \edtext{sva  gild\textit{an}}{\Afootnote{gildari  L679.}}  þussa   o\textit{c} \edtext{íllúþl\textit{e}gri}{\Afootnote{illudligann  L1572.}} \edtext{s\textit{em}}{\Afootnote{enn  L679, L2404.}} Hraungviþ, þ\textit{ví} h\textit{ann} v\textit{ar} bæþi blár o\textit{c} sv\textit{ar}tr; Kári m\textit{ælir} ertú Vícíngr eþr Kaupm\textit{aþr}.  
Hraungv\textit{iþr} m\textit{ælir} e\textit{ck}i er þ\textit{ér} a\textit{t} slíco  \edtext{spyrja}{\Afootnote{ad spirja  L679, L2404, L1572.}}, fyrr e\textit{n} þú f\textit{ær} a\textit{t} reyna, 
o\textit{c} munto e\textit{ck}i min\textit{n}i v\textit{er}þa, eþa hv\textit{aþ} \edtext{hefr þú t\textit{il} nafns}{\Afootnote{heitir þú  L679.}}?  
Kári nemdi s\textit{ic} o\textit{c} Ørnólf.
Hraungv\textit{iþr} m\textit{ælir} hefr þú n\textit{oc}ot  kallm\textit{an}l\textit{e}gt adhafst? 
Kári qv\textit{aþ}st e\textit{ck}i kun\textit{n}a þ\textit{ví} a\textit{t} hrósa. e\textit{n} í xxxiij  \edtext{v\textit{e}tr}{\Afootnote{ár  L2404.}} hef e\textit{c} í vícíngo v\textit{er}it, o\textit{c} m\textit{ar}gt sjeþ o\textit{c} heyrt,  eþa villto leggja  \edtext{t\textit{il}}{\Afootnote{í  L679.}}  atlỏgo   v\textit{it} m\textit{ic}?
Hr\textit{aung}v\textit{iþr} m\textit{ælir} á m\textit{or}gun skalto san\textit{n}a þ\textit{at}, a\textit{t} e\textit{c} sk\textit{al}
e\textit{ck}i  undanteljast. 
Kári m\textit{ælir} þat  þykist e\textit{c} sjá, a\textit{t} fáa muni  \edtext{á þin\textit{n} fund fýsa}{\Afootnote{fýsa á þinn fund  L679.}}, þ\textit{ví} e\textit{c} hefi  LX sin\textit{n}\textit{om}  einvíg  hald\textit{it}, o\textit{c}
afb\textit{or}it þ\textit{ei}m er móti vógo o\textit{c} engin\textit{n} þ\textit{eir}ra leitst mér  \edtext{sva}{\Afootnote{eins  L679.}} þussl\textit{e}gr  \edtext{s\textit{e}m þú ert}{\Afootnote{og þú  L679.}}, 
þ\textit{ví} munto v\textit{er}a hit  mesta fól, o\textit{c}
villda e\textit{c} \edtext{a\textit{t}}{\Afootnote{÷  L679.}} þú  þryfist  alldrei. 
Hr\textit{aung}v\textit{iþr} mæ\textit{lir} \edtext{þ\textit{ví}}{\Afootnote{þess  L679.}} heldr s\textit{em} þér rís hugr v\textit{it} mér fram\textit{ar} ỏþr\textit{om}, s\textit{em} þú segist  v\textit{it} h\textit{afa} b\textit{ar}ist, 
 \edtext{skalto fá a\textit{t} vita, a\textit{t}}{\Afootnote{skalltu vita  L679.}} þó mér blæþi or hv\textit{or}jo sári,  þá  \edtext{man}{\Afootnote{skal  L679.}} e\textit{c} e\textit{ck}i  \edtext{ren\textit{n}a undan þér}{\Afootnote{undan þier renna  L679.}}, o\textit{c} haf\textit{ir} þú hug s\textit{em} \edtext{dug\textit{an}di}{\Afootnote{dugandis  L679, L2404.}}\footnote{L679 and L2404 share the reading \textit{dugandis}, which is an archaic form of \textit{dugandi}, but both forms are registered in OH from the 16th to the 20th cent. In Blöndal's dictionary \textit{dugandis} is translated as hard-working (Da. \textit{dygtig}).} drengr, þá munto á m\textit{or}gun t\textit{il} reyna. 
Kári qv\textit{aþ}st  \edtext{vilja á hætta}{\Afootnote{á hætta vilja  L679.}}, 
o\textit{c} kỏstoþost þ\textit{ei}r lengi á orþ\textit{om}.
Síþ\textit{an} g\textit{en}gr Kári  heim aptr yf\textit{ir} Eyuna þ\textit{an}g\textit{at} s\textit{em} K\textit{óngr} lá á scip\textit{om}
sín\textit{om}. 
Ol\textit{afr} K\textit{óngr} sat yf\textit{ir} b\textit{or}þ\textit{om}, er  \edtext{þ\textit{ei}r}{\Afootnote{÷  L2404.}} b\textit{ræþ}r  \edtext{q\textit{vo}mo}{\Afootnote{komu aptr  L679.}}. 
K\textit{óngr} spyr tíþinda;  \edtext{þ\textit{ei}r}{\Afootnote{÷  L679.}} \edtext{b\textit{ræþ}r}{\Afootnote{÷  L1572.}} s\textit{egja} h\textit{it} san\textit{n}a, a\textit{t} Vícíng\textit{ar}
lægi und\textit{ir} Eyun\textit{n}i, o\textit{c} ásc\textit{or}uþo t\textit{il} b\textit{ar}daga  \edtext{þá}{\Afootnote{þegar  L679.}} m\textit{or}gn\textit{at}i. 
K\textit{ón}gr spratt upp o\textit{c} b\textit{at} m\textit{enn} búast t\textit{il} orrosto o\textit{c} sva g\textit{jỏr}þo þ\textit{ei}r,  \edtext{a\textit{t} þ\textit{ei}r sciptast}{\Afootnote{búast til  L679.}}, draga sum\textit{ir} fram\textit{m} f\textit{yrir} eyuna, e\textit{n} aþr\textit{ir} lỏgþo t\textit{il}
scóg\textit{ar} o\textit{c} t\textit{óc}o sér viþ\textit{ar} kylfr, oc 
||
\ledouternote{108r}
q\textit{vo}mo sva t\textit{il} sin\textit{n}a  \edtext{m\textit{ann}a}{\Afootnote{manna enn  L679.}} \edtext{er vícíng\textit{ar} af visso}{\Afootnote{þá vícingar verþa þess varir þa  L2404.}}, h\textit{er}klæþdost  þ\textit{ei}r o\textit{c} þrifo vapn
sín, o\textit{c} þurfti þ\textit{ar} e\textit{ck}i a\textit{t} \edtext{hỏgg\textit{om}}{\Afootnote{sỏkom  L2404.}} a\textit{t} spyrja,  g\textit{en}go  \edtext{þ\textit{ar}}{\Afootnote{þeir  L679, ÷  L2404.}} hv\textit{or}utveggjo í liþ an\textit{n}\textit{ar}s.  
Þeir Kári o\textit{c} Ørnólfr  \edtext{g\textit{an}ga nú fram\textit{m}}{\Afootnote{gengu hart framm  L679, ganga hart fram  L2404.}} o\textit{c} deyþa m\textit{ar}g\textit{an}  \edtext{Vícíng}{\Afootnote{mann  L679.}}; 
Þ\textit{ei}r Hraungv\textit{iþr} \edtext{o\textit{c} m\textit{enn} h\textit{an}s}{\Afootnote{÷  L679.}}, sæcja  \edtext{þ\textit{an}gat}{\Afootnote{þar ad  L679.}} s\textit{em} Kári er, e\textit{n} er Kári sjer þ\textit{at} t\textit{ec}r h\textit{an}n  st\textit{or}t a\textit{t} vega,  \edtext{sva}{\Afootnote{svo ad  L679.}} h\textit{ann} fell\textit{ir} \edtext{xij í hv\textit{or}jo hỏggi}{\Afootnote{xij i hóggi hvórju  L679, ótal manna  L2404.}}; 
e\textit{n} s\textit{em} Hraungv\textit{iþr} sjer þetta, kall\textit{ar} h\textit{ann} \edtext{ærit}{\Afootnote{afar  L1572.}} hátt o\textit{c} íllil\textit{e}ga o\textit{c} biþr m\textit{enn} sína v\textit{el} fram\textit{m} g\textit{an}ga, 
hleypr  \edtext{h\textit{ann}}{\Afootnote{÷  L679.}} nú uppá  sceiþina, er Ol\textit{afr} K\textit{óngr} v\textit{ar} á, o\textit{c} \edtext{hjó}{\Afootnote{fellir  L679.}} m\textit{ar}g\textit{an}\textit{n}\footnote{Double abbreviation in B11109, the superscript \enquote{n} has a macron above, therefore expanded with two \enquote{n}s. The same abbreviation used in the following word \enquote{menn}.} m\textit{an}\textit{n} \edtext{t\textit{il} b\textit{an}a}{\Afootnote{÷  L679.}}; 
Kári  \edtext{kémr}{\Afootnote{+ þá  L679.}} á móti, 
h\textit{afþi} h\textit{ann} þá  \edtext{hlaupit aptr}{\Afootnote{aptr hlaupid  L1572.}} á K\textit{ón}gs  \edtext{skeiþ}{\Afootnote{s[kip]  L679, sceiþina  L2404, skipid  L1572.}}. 
Hraungv\textit{iþr} m\textit{ælir} til  Kára: 
nú erto q\textit{vo}min\textit{n} híng\textit{at} aflóga hundurin\textit{n}, 
o\textit{c} scul\textit{om} v\textit{it} nú reyna, hv\textit{or}j\textit{om}
betr bíta vapnin. 
Kári reiþdist orþ\textit{om} h\textit{an}s,
o\textit{c} hjó t\textit{il} Hraungv\textit{iþs} 
e\textit{n} h\textit{ann}
skaut  \edtext{skyldi f\textit{yrir} s\textit{ic}}{\Afootnote{fyrir sig skildi  L1572.}}, o\textit{c} festist sv\textit{er}þ\textit{it} í, 
Hraungv\textit{iþr} sjer þ\textit{at}, o\textit{c} sn\textit{ar}ar scyldinom í lopt upp;
Kára v\textit{er}þr laust sv\textit{er}þ\textit{it} \edtext{óv\textit{or}om}{\Afootnote{÷  L1572.}},  \edtext{han\textit{n} hleypr}{\Afootnote{og hleipr  L679, hleypur hann  L1572.}} í lopt  \edtext{ept\textit{ir} þ\textit{ví}}{\Afootnote{upp  L679, upp eftir því  L1572.}}, 
o\textit{c} hæf\textit{ir}  m\textit{ed}alkablan\textit{n} o\textit{c} q\textit{vo}m fj\textit{ær}ri  niþr. 
Ørnólfr h\textit{afþi} staþ\textit{it} n\textit{ær}str Kára br\textit{oþr} s\textit{inum} \edtext{n\textit{ær} eþ han\textit{n}}{\Afootnote{þá hann  L679, L2404.}} ljóp ept\textit{ir} sv\textit{er}þíno; 
Hraungv\textit{iþr} ætla\textit{þ}i a\textit{t} leggja til  Kára, 
e\textit{n} \edtext{misti}{\Afootnote{hann misti  L679.}} h\textit{an}s, þ\textit{at} v\textit{ar} sva gott  \edtext{sv\textit{er}þ}{\Afootnote{+ er Hraungvidr átti  L1572.}}, a\textit{t} v\textit{ar}la  \edtext{f\textit{an}st}{\Afootnote{fékkst  L1572.}} betra\footnote{
	In L679 the entire phrase \enquote{þad var svo gött sverd ad valla fanst betra} is in parenthesis.}  sv\textit{er}þ,  
\edtext{þ\textit{at} er Hraungv\textit{iþr} átti}{\Afootnote{÷  L1572.}}, 
þ\textit{ví} h\textit{ann} kaus jafn\textit{an} m\textit{ann} f\textit{yrir} hỏggi  er h\textit{ann} brá þ\textit{ví}. 
Nú s\textit{em} Kári q\textit{vo}m fjærri niþr leggr Hr\textit{aung}v\textit{iþr} sv\textit{er}þ\textit{it} f\textit{yrir} brjóst Ørnólfi  \edtext{o\textit{c}}{\Afootnote{so L679.}} í gégn\textit{om} h\textit{ann},  \edtext{o\textit{c}}{\Afootnote{stód og  L679.}} kast\textit{ar} h\textit{ono}m xiv feta  l\textit{an}gt af sv\textit{er}þino,  
\edtext{o\textit{c}}{\Afootnote{÷  L1572.}} \edtext{fá þ\textit{ei}r b\textit{an}a er f\textit{yrir} urþo}{\Afootnote{feingu þeir bana er fijrir urdu  L679, oc feck hann bana  L2404, ÷  L1572.}}; 
Kári sjer  þetta o\textit{c} ryþst u\textit{m} fast, þ\textit{ar} t\textit{il} h\textit{ann} hỏggr t\textit{il} Hraungv\textit{iþs} o\textit{c} q\textit{vo}m  hỏgg\textit{it} í hjálmin\textit{n}, 
o\textit{c} v\textit{ar} \edtext{þ\textit{at}}{\Afootnote{÷  L679.}} s\textit{em} í vatn brygþi, 
e\textit{n} er q\textit{vo}m  \edtext{a\textit{t} beinino}{\Afootnote{á brinjuna  L679, ad Brynionni  L2404.}}, beit e\textit{ck}i á; 
Hraungv\textit{iþr} brosti o\textit{c} hjó t\textit{il} Kára; 
 \edtext{Kári}{\Afootnote{en Kári  L2404.}} ljóp í lopt upp ỏfugr aptr á bac yf\textit{ir}
þ\textit{an}\textit{n} er n\textit{ær}str stóþ, o\textit{c} fékk sá hỏgg\textit{it} af Hraungv\textit{iþi} o\textit{c} \edtext{þ\textit{ar}}{\Afootnote{þad  L679, ÷  L1572.}} t\textit{il} b\textit{an}a. 
Hraungv\textit{iþr} m\textit{ælir} 
||
\ledouternote{108v}
laung\textit{om} flýr þú Kári, o\textit{c} hefr þú \edtext{titlínga}{\Afootnote{titlings  L679.}} náttúro,
Kári leypr a\textit{t} Hraungv\textit{iþ} o\textit{c} hỏggr á ỏxlina af mesta  abli, 
sva sv\textit{er}þ\textit{it} gé\textit{ck} í sundr und\textit{ir} hjỏltun\textit{om}, 
e\textit{n} Hraungv\textit{iþr} féll á bæþi  \edtext{kné}{\Afootnote{knien  L679.}},  \edtext{o\textit{c}}{\Afootnote{hann  L679.}} spratt  \edtext{upp þeg\textit{ar}}{\Afootnote{þegar upp  L679.}} o\textit{c} m\textit{ælir} nú kýs ec Kára f\textit{yrir} oþdi
þ\textit{ér}, mín góþa kesja! o\textit{c} v\textit{er}tú þ\textit{ér} nú úti
um rauþ\textit{an} kyrtil, síþ\textit{an} lagþi h\textit{ann}
sv\textit{er}þino  \edtext{f\textit{yrir}}{\Afootnote{i  L679.}} qvid Kára, o\textit{c} \edtext{í}{\Afootnote{÷  L679.}} gégn\textit{om} h\textit{ann}, e\textit{n} Kári stóþ s\textit{em}
bundin\textit{n} \edtext{v\textit{ær}i}{\Afootnote{÷ L679.}} f\textit{yrir} lægino, o\textit{c} veik sér hv\textit{or}gi,
heldr féll m\textit{eþ} dýr\textit{om} drengscap, v\textit{ar}þ h\textit{ann}
m\textit{ỏr}gom h\textit{ar}mdauþi, þ\textit{ví} h\textit{ann} \edtext{hafþi}{\Afootnote{var  L679.}} hin\textit{n} \edtext{ágætasti}{\Afootnote{vænsti  L679.}} m\textit{aþr} \edtext{v\textit{er}it}{\Afootnote{÷  L679.}}   \edtext{m\textit{ed}an}{\Afootnote{+ hann L679, L1572.}} lifþi.
\pend \endnumbering  

\beginnumbering \pstart 
\vspace{5mm} \subsection*{III kap\textit{ítuli}} 
J  \edtext{þ\textit{ess}o bili}{\Afootnote{þann tijma  L679.}} q\textit{vo}mo nú K\textit{ón}gs m\textit{enn} af eyun\textit{n}i m\textit{eþ} sín\textit{ar} viþar  kylfr s\textit{em} fyrr er frá sagt, laupa þ\textit{ei}r nú á scip Ol\textit{afs} K\textit{óngs} v\textit{ar} þá Kári fallin\textit{n} o\textit{c} Ørnólfr, h\textit{ann} h\textit{afþi} b\textit{or}it m\textit{er}ki  \edtext{Ol\textit{afs}}{\Afootnote{÷  L679.}} K\textit{óngs} Hr\textit{aung}v\textit{iþr} b\textit{at} Ol\textit{af} K\textit{óng}
 \edtext{a\textit{t}}{\Afootnote{÷  L679.}} uppgéfast  \edtext{m\textit{eþ} liþ}{\Afootnote{+ sitt  L679, L2404, oc ganga med lidid  L1572.}} 
 \edtext{o\textit{c} g\textit{an}ga}{\Afootnote{÷ L1572.}} sér t\textit{il} h\textit{an}da; e\textit{n} K\textit{óngr} vænti liþs o\textit{c} tregþaþi andsv\textit{ar} a\textit{t} géfa. 
 Und\textit{an} Kóngs liþi \edtext{ljóp}{\Afootnote{+ einn L679, L1572.}} m\textit{aþr} heldr vascl\textit{e}gr, 
 hár o\textit{c} h\textit{er}þabreiþr, h\textit{ann} h\textit{afþi} síþan\textit{n} hatt á hỏfþi, s\textit{em} slapti of\textit{an} f\textit{yrir} augo h\textit{an}s, 
 geit\textit{ar} scégg h\textit{afþi} h\textit{ann} \edtext{hvítgrátt}{\Afootnote{kvijtt  L679.}} o\textit{c} \edtext{m\textit{ic}it síþt}{\Afootnote{sva micit  L2404.}}, 
 \edtext{sva þ\textit{at}}{\Afootnote{at  L2404.}} \edtext{tóc á bríngo}{\Afootnote{á bringo tók  L2404.}} niþr, h\textit{ann} ljóp  á K\textit{ón}gs scip,  \edtext{o\textit{c}}{\Afootnote{÷  L679.}} þ\textit{ar} t\textit{il} 
 \edtext{a\textit{t}}{\Afootnote{er  L1572.}} \edtext{Kári}{\Afootnote{÷ L679.}} v\textit{er}þr f\textit{yrir} 
 \edtext{fót\textit{om}}{\Afootnote{honum  L2404.}} \edtext{h\textit{an}s}{\Afootnote{+ Kári daudr  L679, ÷  L2404.}}, þá mæ\textit{lir} h\textit{ann}: hv\textit{or}t  erto fallin\textit{n}
fóstbr\textit{óþir}? o\textit{c} fin\textit{n}ast f\textit{ær}ri þín\textit{ir} líc\textit{ar},  
\edtext{o\textit{c}}{\Afootnote{+ er L679.}} 
 lícast \edtext{t\textit{il}, a\textit{t}}{\Afootnote{÷ L679,  L1572.}} 
 e\textit{c} leiti h\textit{an}s, er þ\textit{ér} hefr á þiljo kastat, 
 litast h\textit{ann} \edtext{þá}{\Afootnote{÷  L679.}}   
 \edtext{u\textit{m}}{\Afootnote{+ og L679, L2404, L1572.}},  
 \edtext{sjer}{\Afootnote{finnur  L679.}} Ørnólf dauþ\textit{an} liggja, 
 o\textit{c} m\textit{ar}gt an\textit{n}\textit{at} liþ, þykist 
 \edtext{nú þ\textit{ess}i m\textit{aþr}}{\Afootnote{hann nú  L2404.}} vita  
 \edtext{a\textit{t}}{\Afootnote{÷ L2404.}} m\textit{er}ki K\textit{ón}gs muni  \edtext{niþr }{\Afootnote{÷  L679.}} fall\textit{it}, 
þrýfr  \edtext{t\textit{il}}{\Afootnote{hann nú til þess  L679.}} o\textit{c} 
\edtext{reisti}{\Afootnote{reisir  L679.}}  \edtext{upp}{\Afootnote{þad upp  L679.}} aptr, o\textit{c} kall\textit{ar} á K\textit{óng} o\textit{c} biþr h\textit{ann} v\textit{er}a glaþ\textit{an} o\textit{c} eggja liþit t\textit{il} k\textit{ar}lm\textit{an}l\textit{e}gr\textit{ar}
fram\textit{m}gaungo.  Ol\textit{afr} K\textit{óngr} m\textit{ælir} hv\textit{or} ert þú  hin\textit{n} m\textit{ic}li m\textit{aþr}? h\textit{ann}
sv\textit{ar}ar: Hrómundr er  \edtext{nafn mitt}{\Afootnote{mitt mitt nafn  L679.}},\footnote{Dittography in L679: \enquote{mitt mitt nafn}.}  o\textit{c} er Greipss\textit{on} \edtext{Arm\textit{aþr}}{\Afootnote{eirn madr  L679, L2404, L1572.}} yþ\textit{ar}, o\textit{c} v\textit{or}om
v\textit{it} \edtext{Kári}{\Afootnote{bádir  L679.}} \edtext{lagsm\textit{enn}}{\Afootnote{+ vid Kári  L679.}} 
o\textit{c} fóst||\ledouternote{109r}bræþr, þá v\textit{it} heima vór\textit{om}. K\textit{óngr}
m\textit{ælir} ólícr sýnist o\textit{ss} þú  \edtext{a\textit{t} v\textit{er}a}{\Afootnote{vera ad þú siert  L679.}} \edtext{fóstbróþ\textit{ir}}{\Afootnote{lagsmadr  L679.}} Kára,  \edtext{þ\textit{ar}}{\Afootnote{þvj  L679.}} þú  \edtext{orþin\textit{n} ert}{\Afootnote{ert ordinn  L679, L2404, ert  L1572.}} gamall o\textit{c} grár af  hær\textit{om}, e\textit{n} ef þú
hefr h\textit{an}s  \edtext{lagsm\textit{aþr}}{\Afootnote{fostbrodir  L679.}} v\textit{er}it, þá manto h\textit{an}s hefna  á
þ\textit{ess}\textit{om} degi. 
Hróm\textit{undr} m\textit{ælir}\footnote{A colon following the letter \enquote{m} in B11109 is interpreted as an abbreviation mark, not as a punctuation mark.}  e\textit{ck}i v\textit{ar} Kári hyrþm\textit{aþr} min\textit{n} eþr  stafnbúi, e\textit{n} þó villda e\textit{c} \edtext{a\textit{t} mér f\textit{ær}i}{\Afootnote{miér færist  L679.}} ei verr e\textit{n} þ\textit{ei}m  \edtext{er}{\Afootnote{sem  L679, L2404.}} ept\textit{ir} \edtext{slíca m\textit{enn}}{\Afootnote{slijkann mann  L679, L2404.}} \edtext{ega}{\Afootnote{eiga  L679, L1572, áttir  L2404.}}\footnote{The form \enquote{ega} is an archaic spelling of verb \textit{eiga}, it can be found in OH in the 19th cent.}  a\textit{t} mæla s\textit{em} þ\textit{ei}r  \edtext{b\textit{ræþ}r}{\Afootnote{fostbræþur  L2404.}} v\textit{or}o; 
ljóp þá Hróm\textit{undr} \edtext{á}{\Afootnote{upp á  L679.}} dr\textit{ec}an Hraungviþs, 
h\textit{ann} h\textit{afþi} st\textit{or}a gaþdakylfo í an\textit{n}\textit{ar}i hendi,  e\textit{n} sv\textit{er}þ í an\textit{n}\textit{ar}i o\textit{c} \edtext{scjỏld}{\Afootnote{skjólldinn  L679.}}, 
e\textit{n} hjálm eng\textit{an}, nema stálhúfo, 
und\textit{ir} hattin\textit{om}, lemr h\textit{ann} nú o\textit{c}
b\textit{er} allt  \edtext{þ\textit{at}}{\Afootnote{hvad  L679, sem  L2404.}} f\textit{yrir} v\textit{er}þr m\textit{eþ} kylfun\textit{n}i,
e\textit{n} er Hraungv\textit{iþr} sj\textit{er} þetta, a\textit{t} liþ
h\textit{an}s fellr hrỏnn\textit{om},  
v\textit{er}þr h\textit{ann} ákafl\textit{e}ga reiþr, o\textit{c} spyr: hv\textit{or} erto \edtext{þin\textit{n}}{\Afootnote{hinn  L679, L2404, L1572.}} grái  \edtext{scégg}{\Afootnote{skeggadi  L679.}} brúsi, s\textit{em} sva g\textit{jỏr}ir a\textit{t} 
\edtext{drepa o\textit{c} deyþa}{\Afootnote{deida og drepa  L679.}} m\textit{eþ} \edtext{lurci o\textit{c} spít\textit{om}}{\Afootnote{spijtum og lurkum  L679.}}, 
þú mant v\textit{er}a  \edtext{þin\textit{n}}{\Afootnote{hinn  L679, ÷  L2404.}} \edtext{ólukko}{\Afootnote{vondur  L2404.}} fretk\textit{ar}l, faþ\textit{ir} Kára. 
Hrómundr m\textit{ælir} hv\textit{or}ki er e\textit{c} 
faþ\textit{ir} \edtext{h\textit{an}s}{\Afootnote{÷  L679.}}, \edtext{sun}{\Afootnote{sonr  L679, L2404, L1572.}} eþr  \edtext{frændi}{\Afootnote{+ hans  L679.}}, 
en\textit{n} þó  scalto  \edtext{fá a\textit{t}}{\Afootnote{þad  L679.}} frétta, a\textit{t} e\textit{c} sk\textit{al} hefna þ\textit{eir}ra bræþra, o\textit{c} reyna manto kylfo mína, áþr e\textit{n}
v\textit{it} scilj\textit{om}, e\textit{n} Hróm\textit{undr} h⟨ei⟩ti e\textit{c}, o\textit{c} \edtext{vita þykist}{\Afootnote{veit  L679, vita þykjumst  L2404.}} e\textit{c}, a\textit{t} þú ert h\textit{it} v\textit{er}sta trỏll,
a\textit{t} engin\textit{n} j\textit{ar}n bíta á þ\textit{ic}, o\textit{c} þ\textit{ar} f\textit{yrir} hæf\textit{ir} þér bỏþla refsíng.
Hraungv\textit{iþr} v\textit{ar}þ æf\textit{ar} reiþr
o\textit{c} leypr a\textit{t} Hróm\textit{undi}, o\textit{c} ætlaþi  a\textit{t} kjósa  \edtext{h\textit{ann}}{\Afootnote{÷  L679.}} 
f\textit{yrir}  oþdi  ept\textit{ir} \edtext{v\textit{an}a}{\Afootnote{+ sijnum  L679.}}, e\textit{n} Hróm\textit{undr} v\textit{ar}þ fljót\textit{ar}i, oc  sló kylfun\textit{n}i ut\textit{an} und\textit{ir} v\textit{an}g\textit{an}\textit{n}\footnote{An unclear reading in L679: \enquote{vanga} or \enquote{varga}.}  á Hraungv\textit{iþ} svo st\textit{or}t  hỏgg, a\textit{t} hausin\textit{n} brotnaþi í smá mola,  \edtext{e\textit{n}}{\Afootnote{svo  L679.}} heilinn  
\edtext{vall  út}{\Afootnote{vall úr L679, ÷  L1572.}}, 
\edtext{oc}{\Afootnote{÷  L1572.}} lac niþr
\edtext{s\textit{em} vatn}{\Afootnote{÷ L2404}}, 
féll þá Hraungv\textit{iþr} í  svíma,  sva h\textit{ann}  rauc  út af,  \edtext{a\textit{t}}{\Afootnote{svo  L679, og  L1572.}} \edtext{m\textit{ar}g\textit{ir}  þeinkto}{\Afootnote{þenktu margir  L1572.}} h\textit{ann} dauþ\textit{an}\textit{n}; nú  sprettr Hraungv\textit{iþr} á fætr aptr, o\textit{c} m\textit{ælir} t\textit{il}
Hróm\textit{undar}\footnote{In B11109, there is a colon, which can introduce the direct speech, but there is no abbreviation mark to denote the abbreviation of the name. In this and similar cases the colon is treated as an abbreviation mark.} 
e\textit{c} hefi  víþa v\textit{er}it o\textit{c}
f\textit{ar}it, o\textit{c} m\textit{ar}g\textit{an} kappa a\textit{t} velli lagt, o\textit{c} \edtext{hefi}{\Afootnote{hef  L679, L2404, L1572.}} ||
\ledouternote{109v} %16:36
e\textit{c} alldrei þ\textit{ví}líkt hỏgg fengit á allri
min\textit{n}i æfi, o\textit{c} alldrei  \edtext{hefi}{\Afootnote{hef  L679, L2404, L1572.}} e\textit{c} þin\textit{n} jafníngja  \edtext{fundit}{\Afootnote{+ á æfi minni  L679.}}, o\textit{c} v\textit{er}þr einosin\textit{n}i allt  fyrst,
féll þá niþr kesj\textit{an} or hỏnd\textit{om} h\textit{an}s, lamdi Hróm\textit{undr} h\textit{an}n 
m\textit{eþ} kylfun\textit{n}i t\textit{il}   \edtext{dauþs}{\Afootnote{+ og  L679.}}, 
lauc sva Hraungv\textit{iþr} æfi sin\textit{n}i. Ept\textit{ir} þetta v\textit{ar} nú  hróp\textit{at}  \edtext{sigurs  hróp}{\Afootnote{sigr óp  L679, sigr hróp  L1572, L2404.}}, o\textit{c} m\textit{ỏnnom}  griþ  géfin; er þ\textit{at}  þyggja  villdo, kallaþi
Hróm\textit{undr} t\textit{il} þ\textit{eir}ra er ept\textit{ir} stóþo,
o\textit{c} spyr:  hv\textit{or}t þ\textit{ei}r  \edtext{vilji heldr}{\Afootnote{vildu heldur  L679, heldur vilji  L2404.}} \edtext{kjósa}{\Afootnote{þiggja  L679.}} a\textit{t} g\textit{an}ga á hendr Ol\textit{afs} K\textit{óngs} eþr  \edtext{a\textit{t}}{\Afootnote{÷  L679.}} b\textit{er}jast  \edtext{[\textit{oc}]}{\Afootnote{÷  L2404, L1572.}} villdo all\textit{ir} \edtext{o\textit{c} v\textit{el} flest\textit{ir}}{\Afootnote{÷  L1572.}} \edtext{heldr}{\Afootnote{÷  L679.}} 
K\textit{óngi} t\textit{il} \edtext{h\textit{an}da o\textit{c}  hlýþnis}{\Afootnote{hlijdni og handa  L679, handa  L1572, L2404.}} g\textit{an}ga, 
lauc sva b\textit{ar}dag\textit{an}om, a\textit{t} m\textit{enn} sættost.  \pend \endnumbering  

\beginnumbering \pstart   
\vspace{5mm}\subsection*{IV kap\textit{ítuli}}
Um m\textit{or}gunin\textit{n} ept\textit{ir} f\textit{er}\footnote{
	The scribe of B11109 uses the same abbreviation mark for \enquote{eptir} and \enquote{fer}, both expanded following the written-out forms from the manuscript.}  
Hróm\textit{undr} t\textit{il} m\textit{eþ} liþi K\textit{ón}gs a\textit{t} kan\textit{n}a  scip Hraungv\textit{iþs} o\textit{c} binda  \edtext{of}{\Afootnote{um  L679, L2404, ÷  L1572.}}\footnote{
	The scribe's \enquote{a}'s and \enquote{o}'s resemble each other closely in B11109, so it could be \enquote{af} or \enquote{of}; the reading \enquote{of} is more probable in this context.}  
sár þ\textit{eir}ra er fall\textit{it} hỏfþo, 
e\textit{n} þó lífs v\textit{or}o, lítast h\textit{ann} nú u\textit{m},
o\textit{c} \edtext{of}{\Afootnote{um  L679, L2404, L1572.}} síþ\textit{ir}
fin\textit{n}r h\textit{ann} ein\textit{n} m\textit{ann} \edtext{f\textit{yrir} sér}{\Afootnote{÷  L679.}} á  drecan\textit{om}, h\textit{ann} sá a\textit{t} augo h\textit{an}s lifþo, e\textit{n}  lim\textit{ir}   hr\textit{ær}þost   valla; 
þ\textit{ess}i  m\textit{aþr} lá í  \edtext{lyptíngo}{\Afootnote{+ á drekanum  L679.}}.
Hróm\textit{undr} spyr: hv\textit{or}t þ\textit{ess}i megi mæla?
hin\textit{n}  ansaþi:  \edtext{lin\textit{n}\textit{ir} þ\textit{ví}}{\Afootnote{eij er þad  L679.}} sva a\textit{t} segja. Hróm\textit{undr} \edtext{m\textit{ælir}}{\Afootnote{spir  L679, L1572.}} hv\textit{or}t er nafn  þitt? h\textit{ann}
s\textit{varar} Helgi h\textit{ei}ti e\textit{c}, o\textit{c}
v\textit{ar} e\textit{c} kallaþr Helgi hin\textit{n} Fræcni,
o\textit{c} v\textit{ar} Hraungv\textit{iþ} br\textit{óþir} m\textit{inn}
sva b\textit{ar} t\textit{il}, a\textit{t} v\textit{it} b\textit{or}þomst
\edtext{v\textit{it} Scotta}{\Afootnote{÷ L679.}} 
\edtext{f\textit{yrir} full\textit{om} xiv dỏg\textit{om}}{\Afootnote{vid skotta L679.}}, a\textit{t} e\textit{c} s\textit{ær}þist sva e\textit{c}
v\textit{ar}þ óvígr, o\textit{c} hefi ec  leg\textit{it} í
re\textit{ck}jo síþ\textit{an}, e\textit{n} br\textit{óþir} m\textit{inn}
b\textit{ar} af flest\textit{om}, o\textit{c} héldt híngat, 
hefi e\textit{c} lát\textit{it} m\textit{ic} or re\textit{ck}jo f\textit{ær}a
m\textit{ér} t\textit{il} gam\textit{an}s a\textit{t} sjá á rỏscva  m\textit{enn} o\textit{c} kémpr, jafnv\textit{el} \edtext{þótt}{\Afootnote{þo  L679.}} e\textit{c} \edtext{eckért sjálfr gæti}{\Afootnote{gjæti eckert sjalfr  L679.}}, 
hefr þú  un\textit{n}\textit{it} h\textit{it}
\edtext{v\textit{er}sta o\textit{c}}{\Afootnote{÷ L679.}} óþ\textit{ar}fasta   v\textit{er}c æfi
þin\textit{n}\textit{ar}, a\textit{t} drepa Hraungviþ br\textit{óþr} m\textit{inn} s\textit{em} v\textit{ar} hin\textit{n} vascasti  m\textit{aþr}.
Hróm\textit{undr} m\textit{ælir} villto þyggja griþ af o\textit{ss} eþr ekki? 
Helgi m\textit{ælir} þ\textit{at} er nú í þin\textit{n}i hendi, e\textit{n} heldr er e\textit{c} n\textit{ær} a\textit{t} deya, e\textit{n} þ\textit{at} e\textit{c} \edtext{géti aldrei hefnt}{\Afootnote{hefni alldrei  L679.}} Hraungv\textit{iþ} 
þ\textit{ví} a\textit{t} v\textit{it} \edtext{hỏfþ\textit{om}}{\Afootnote{hófum  L679, L1572.}} h\textit{ei}t\textit{it} þ\textit{ví} hv\textit{or} ỏþr\textit{om}, a\textit{t} sá 
||
\ledouternote{110r} 
 scyldi ærulaus  
m\textit{ann}a á m\textit{ed}al s\textit{em} e\textit{ck}i hefndi,
o\textit{c} féll m\textit{ér} 
\edtext{nú t\textit{il} þ\textit{ess}}{\Afootnote{þad nú til  L679, L1572, nú til þad  L2404.}} s\textit{em} e\textit{c} gét ei af hỏnd\textit{om} leyst nema e\textit{c}
græþslo  fái, e\textit{n} þ\textit{at} læt e\textit{c} þ\textit{ér}  ódul\textit{it}, \edtext{a\textit{t}}{\Afootnote{÷  L679.}} ef e\textit{c} næ lífi a\textit{t} halda, þá sk\textit{al} e\textit{c} e\textit{ck}i láta b\textit{ræþ}r þína  \edtext{v\textit{er}þa elli dauþa}{\Afootnote{elli dauda verda  L679.}}. 
Hróm\textit{undr} t\textit{óc} a\textit{t}  undrast  orþ Helga,  
\edtext{e\textit{n}}{\Afootnote{er  L679, L2404, því L1572.}}\footnote{
	The word is abbreviated in L1572 and its expansion is uncertain, might be \enquote{því} or \enquote{þar}.} 
h\textit{ann}  flakti  
\edtext{allr}{\Afootnote{sundr  L679.}} af sár\textit{om} \edtext{í sundr}{\Afootnote{÷  L679.}}.  
þá m\textit{ælir} Hróm\textit{undr} \edtext{þin\textit{n}\textit{ar}  hreysti}{\Afootnote{hreisti þinnar  L679.}} scalto njóta, 
en\textit{n} e\textit{ck}i heitstrengíng\textit{ar}, o\textit{c}  ávin\textit{n}st  þ\textit{at}
s\textit{em}  auþn\textit{an} leyf\textit{ir}; 
v\textit{ar} Helgi
t\textit{ec}in\textit{n} o\textit{c} læcnaþr, o\textit{c} lét Hróm\textit{undr} h\textit{ann} \edtext{h\textit{it}}{\Afootnote{þad  L679, L1572.}} besta  viþrv\textit{ær}i  fá þ\textit{ar} t\textit{il} \edtext{a\textit{t}}{\Afootnote{÷  L679, L2404.}} h\textit{ann} v\textit{ar} fullgróin\textit{n}, 
lícr hér nú  \edtext{af}{\Afootnote{fijrst frá  L679.}} h\textit{ono}m a\textit{t} segja  \edtext{fyrst  a\textit{t} sin\textit{n}i}{\Afootnote{÷  L679, fyrst at segia  L2404.}}. 
Ept\textit{ir}  \edtext{allt}{\Afootnote{÷  L679.}} þetta 
lætr Ol\textit{afr} K\textit{óngr} leysa upp
scip sín, o\textit{c} \edtext{halda}{\Afootnote{+ sijdann  L679, heldr  L2404.}} austr í haf, o\textit{c} sva þ\textit{aþ}an f\textit{yrir} Svíþjóþ, t\textit{ac}a þ\textit{ei}r sér  \edtext{víþa strandhỏgg}{\Afootnote{strandhógg vijda  L679.}}, 
ræna fé, e\textit{n}  h\textit{er}ja  á m\textit{enn} oc  bren\textit{n}a bæi, ept\textit{ir} Vícínga  ⸌venjo⸍, \edtext{sva}{\Afootnote{og  L679.}} víþa spurþist  til  þ\textit{eir}ra st\textit{or}v\textit{er}ca. 
\pend \endnumbering  

\beginnumbering \pstart  
\vspace{5mm}\subsection*{V kap\textit{ítuli}} 
J þen\textit{n}\textit{an} tíma réþo f\textit{yrir} Svíaveldi ij  \edtext{Kóng\textit{ar}}{\Afootnote{kongasynir fyrir Svíjaveldi  L2404.}},  \edtext{sem}{\Afootnote{er  L679, oc  L2404.}} \edtext{Hálfdán\textit{ar} h\textit{ie}to}{\Afootnote{hietu Hálfdánar hverotveggio  L2404.}}, \edtext{er þetta g\textit{jỏr}þist}{\Afootnote{÷  L1572.}}, sigl\textit{ir} þá Ol\textit{afr} K\textit{óngr} t\textit{il} Eyar þ\textit{ei}rr\textit{ar} \edtext{s\textit{em}}{\Afootnote{er  L679, L2404, L1572.}} Mána Ey er kỏlluþ, 
hún liggr í suþreya  hafi, \edtext{þ\textit{an}g\textit{at}}{\Afootnote{+ til  L679.}} lét K\textit{óngr} halda scip\textit{om} sín\textit{om} á eino qvỏldi, þ\textit{ar} g\textit{en}go þ\textit{ei}r á l\textit{an}d,   \edtext{uppá eina fagra eyri}{\Afootnote{upp  L679.}}, 
o\textit{c} litast u\textit{m}, o\textit{c} \edtext{sjá}{\Afootnote{+ þar  L679, sáu  L1572.}} e\textit{ck}ert  \edtext{þ\textit{at}}{\Afootnote{÷  L679, hvad  L1572.}} m\textit{ar}cv\textit{er}þugt  
\edtext{kalla}{\Afootnote{÷  L679.}} \edtext{má}{\Afootnote{÷  L679, L2404, mætti  L1572.}}, 
g\textit{an}ga  \edtext{þá}{\Afootnote{þeir  L2404.}} K\textit{ón}gs m\textit{enn} lengra á Eyuna, o\textit{c} í þ\textit{ess}o
sjá þ\textit{ei}r  \edtext{ein\textit{n}}{\Afootnote{÷  L679.}} gamlan\textit{n} m\textit{ann} \edtext{q\textit{vo}ma}{\Afootnote{ganga  L679.}} á veg f\textit{yrir} þá, heylsa þ\textit{ei}r  \edtext{á h\textit{ann}}{\Afootnote{honum  L2404.}}, o\textit{c} spyrja  a\textit{t} nafni,  \edtext{e\textit{n}}{\Afootnote{÷  L2404.}} h\textit{ann} \edtext{sagþist}{\Afootnote{kvadst  L679, qvad  L2404.}} Máni  \edtext{h\textit{ei}ta}{\Afootnote{heiti  L679.}}, e\textit{n} hv\textit{or}j\textit{ir} erot þér? 
þ\textit{ei}r qv\textit{aþ}ost v\textit{er}a m\textit{enn} Ol\textit{afs} \edtext{K\textit{óngs}}{\Afootnote{kongi  L679.}} K\textit{ar}l m\textit{ælir} \edtext{hv\textit{ar} t\textit{il}}{\Afootnote{til kvórs  L679, L2404, L1572.}} er  \edtext{Ol\textit{afr}}{\Afootnote{÷  L679.}} K\textit{óngr} híng\textit{at} q\textit{vo}min\textit{n}? þ\textit{ei}r
s\textit{egja} h\textit{ann} er hin\textit{n} ágætasti  \edtext{K\textit{óngr} af Sækóng\textit{om}}{\Afootnote{sækongr  L679, kongr  L2404.}}, o\textit{c} lætr abla sér  \edtext{o\textit{c} m\textit{ỏnnom} sín\textit{om}}{\Afootnote{÷  L679.}} fjár o\textit{c} \edtext{frægþar}{\Afootnote{frægþa  L2404.}}. K\textit{ar}l m\textit{ælir} lítil frægþ er þ\textit{at} slíc\textit{om} m\textit{ann}i s\textit{em} Ol\textit{afr} K\textit{óngr} er  sagþr, a\textit{t} h\textit{ann} sculi  \edtext{láta}{\Afootnote{÷  L679.}}  féfletta    vesælínga   o\textit{c} \edtext{fátæca}{\Afootnote{+ menn  L679.}} 
||
\ledouternote{110v}
 \edtext{o\textit{c}}{\Afootnote{÷  L679, enn  L1572.}} hitt er  \edtext{l\textit{an}gt\textit{om}}{\Afootnote{÷  L679.}} n\textit{ær} líkind\textit{om} a\textit{t} f\textit{ar}a á  \edtext{mót}{\Afootnote{móti  L2404, mots  L1572.}} Vícíng\textit{om} s\textit{em} hér liggja hino megin\textit{n}
ey\textit{ar}in\textit{n}\textit{ar}, o\textit{c} reyna v\textit{it} þá,
heldur e\textit{n} a\textit{t} g\textit{jỏr}a mér   tjón   \edtext{eþr}{\Afootnote{og  L679.}}  scaþa; þ\textit{ess}ir  \edtext{vita}{\Afootnote{hafa vitad  L679.}} allt af yþr,  o\textit{c} h\textit{afa} í ráþi, a\textit{t}
t\textit{ac}a af yþr scipin o\textit{c} all\textit{an} far\textit{ar} beina,  \edtext{þ\textit{ví} þ\textit{ei}r e\textit{ro}}{\Afootnote{ero þeir  L1572.}} hin\textit{ir} v\textit{er}sto m\textit{enn} o\textit{c}
h\textit{er}vícíng\textit{ar},  \edtext{o\textit{c}}{\Afootnote{÷  L679.}} 
h\textit{afa} þ\textit{ei}r  \edtext{fal\textit{it} hér}{\Afootnote{falid L679, herfalid  L1572.}} fé á  l\textit{an}di, 
t\textit{il} hv\textit{or}s e\textit{c} \edtext{kyn\textit{n}i a\textit{t}}{\Afootnote{kann  L679.}} segja, ef vin\textit{ir} mín\textit{ir} v\textit{ær}ot. Hróm\textit{undr} stendr  \edtext{nú}{\Afootnote{÷  L679.}} upp v\textit{it} \edtext{scipin}{\Afootnote{skipid  L679.}}, þ\textit{ví} h\textit{ann} h\textit{afþi} e\textit{ck}i á l\textit{an}d
g\textit{en}g\textit{it}, þykist  h\textit{ann} heyra
m\textit{ann}a mál, o\textit{c} hinomegin\textit{n}
Ey\textit{ar}in\textit{n}\textit{ar} \edtext{v\textit{ær}o}{\Afootnote{lego  L2404, væri  L1572.}} Vícíng\textit{ar}, g\textit{en}gr h\textit{ann} á
l\textit{an}d t\textit{il} K\textit{ón}gs m\textit{ann}a, o\textit{c} spyr
a\textit{t}: hv\textit{or} er sá   kratt\textit{an}s   k\textit{ar}l,
s\textit{em} \edtext{a\textit{t}}{\Afootnote{svo  L679.}} lýgr o\textit{c}  n\textit{ar}r\textit{ar}  \edtext{o\textit{ss}}{\Afootnote{ockr  L679.}}? þ\textit{ei}r s\textit{egja} h\textit{ono}m slíct  \edtext{er}{\Afootnote{+ þeir  L679.}} visso. 
Hróm\textit{undr} \edtext{s\textit{egir} hvaþa}{\Afootnote{÷  L2404.}}  scỏm\textit{m}  er  \edtext{þetta}{\Afootnote{þad ad  L679, þad  L2404.}},  
\edtext{þér}{\Afootnote{þid  L679.}} látit  \edtext{þen\textit{n}an\textit{n}}{\Afootnote{hann  L679.}}  ganta   \edtext{yþr}{\Afootnote{ockr  L679.}}. Máni sv\textit{arar} þ\textit{at} er l\textit{an}gt\textit{om} betra
f\textit{yrir} góþa m\textit{enn} a\textit{t} hjálpa vesæl\textit{om} m\textit{ỏnnom}, e\textit{n}
a\textit{t}  hrakyrþa   þá,  \edtext{o\textit{c}  níþa}{\Afootnote{÷  L679, L2404.}} niþr í  orþ\textit{om}, o\textit{c} \edtext{betra}{\Afootnote{litit  L2404.}} þyc\textit{ir} þ\textit{ei}m sum\textit{om} hv\textit{or}j\textit{om}
a\textit{t} \edtext{ræna}{\Afootnote{fiefletta  L679.}} m\textit{ic}; sjá\textit{it} þ\textit{at} e\textit{ck}i  \edtext{a\textit{t}}{\Afootnote{hvar  L1572.}} þ\textit{ei}r f\textit{ar}a þ\textit{ar} m\textit{eþ} kýr mín\textit{ar},
sva e\textit{c} má bj\textit{ar}g\textit{ar}laus  \edtext{ept\textit{ir}}{\Afootnote{÷  L2404.}} v\textit{er}a,  í þ\textit{ess}\textit{om} kotbæ, er
e\textit{c} \edtext{uppbyggt hefi}{\Afootnote{hefi uppbigt  L679.}} á eyo þ\textit{ess}ari, þykist ec  nú vita, a\textit{t} þér
munduþ  \edtext{vilja}{\Afootnote{÷  L679, L2404, L1572.}} slíct eyrindi ega, at  ræna m\textit{ic} grip\textit{om} \edtext{o\textit{c} gagnssemd\textit{om}}{\Afootnote{÷  L2404.}}, e\textit{n} e\textit{c}  hygg   þ\textit{at}  sæmil\textit{e}gra   f\textit{yrir} þ\textit{ic} Hróm\textit{undr} o\textit{c} þína líca, a\textit{t} fást
v\textit{it} nokkr   st\textit{or}vyrci,  \edtext{sva}{\Afootnote{sem nockur  L679.}} 
frægþ o\textit{c}  f\textit{or}prís megi   afg\textit{an}ga. 
Hróm\textit{undr} m\textit{ælir} t\textit{il} hv\textit{or}ra st\textit{or} ræþa   vís\textit{ar}  þú o\textit{ss}? 
Máni sv\textit{arar} Þráin\textit{n} h\textit{ie}t m\textit{aþr}, h\textit{ann} v\textit{ar} h\textit{it} allra \edtext{v\textit{er}sta}{\Afootnote{mesta  L679.}} galdra trỏll, fullr fjỏlcíngis o\textit{c} fj\textit{an}dscap\textit{ar}, h\textit{ann}  h\textit{er}jaþi   á  \edtext{Saxl\textit{an}d}{\Afootnote{Valland  L679.}}\footnote{Note that L679 agrees here with the \textit{rímur}, which have \enquote{Valland}.}  o\textit{c} v\textit{an}\textit{n} þ\textit{at} m\textit{eþ}
gỏldr\textit{om}, o\textit{c}  \edtext{v\textit{ar}þ}{\Afootnote{var L679.}} þ\textit{ar} síþ\textit{an} \edtext{K\textit{ón}gr}{\Afootnote{ + yfir  L679.}}; 
h\textit{ann} átti  \edtext{ofrnægþ}{\Afootnote{ofur gnægd  L679, L2404, L1572.}}\footnote{
	Note that a better reading is in L679, L2404, and L1572 than in B11109.} af gulli o\textit{c} sylfri o\textit{c} ỏþr\textit{om}
dýrgrip\textit{om}; h\textit{ann} héldt  m\textit{ar}ga  \edtext{b\textit{er}s\textit{er}ci o\textit{c}
	blám\textit{enn}}{\Afootnote{blámenn og berserki  L1572.}}, s\textit{em} þ\textit{at} hỏfþo t\textit{il} scémtun\textit{ar} sér, at
glíma v\textit{it} \edtext{scóg\textit{ar}eykr}{\Afootnote{eikr  L679.}} o\textit{c} Fíla, o\textit{c}  jamn\textit{an}  v\textit{or}o  \edtext{þ\textit{ei}r}{\Afootnote{+ brædur  L2404.}} m\textit{eþ} K\textit{on}gi,  \edtext{þ\textit{ar} t\textit{il}}{\Afootnote{þá  L1572.}} h\textit{ann} \edtext{deyþi}{\Afootnote{dó  L2404.}}, drógo þ\textit{ei}r \edtext{h\textit{ann} þá á}{\Afootnote{á hann  L1572.}} \edtext{scóg}{\Afootnote{skogien  L679.}}, o\textit{c} urpo haug \edtext{yfir}{\Afootnote{+ hann  L679.}},  o\textit{c} léto h\textit{ann} í haugin\textit{n},
o\textit{c} gull  \edtext{þ\textit{at}}{\Afootnote{÷  L2404.}} er h\textit{ann} átti; þ\textit{ei}r þỏkto h\textit{ann} \edtext{gullhríngum}{\Afootnote{+ og  L679, L2404.}}, spen\textit{n}to fagort ⸌men⸍  a\textit{t} hálsi  \edtext{h\textit{an}s}{\Afootnote{honum  L679, L2404.}},  \edtext{e\textit{n} sv\textit{er}þ lỏgþo}{\Afootnote{oc lögþo sverþ  L2404.}} \edtext{þ\textit{ei}r í h\textit{an}s}{\Afootnote{i  L2404.}} 
||
\ledouternote{111r} 
 \edtext{kné}{\Afootnote{+ hans  L2404.}}; m\textit{ar}g\textit{ir} m\textit{enn} o\textit{c} \edtext{kémpr}{\Afootnote{kappar L2404.}} h\textit{afa} reynt s\textit{ic} \edtext{v\textit{it}}{\Afootnote{til  L679, ÷  L1572.}} a\textit{t} g\textit{an}ga í haugin\textit{n}, e\textit{n}
Þráin\textit{n} er ramr draugr, a\textit{t} h\textit{ann} rífr af mỏn\textit{n}\textit{om} hold  o\textit{c} klæþi, o\textit{c}
engv\textit{ir} q\textit{vo}most þaþ\textit{an} lífs er þ\textit{an}gat  \edtext{f\textit{or}o}{\Afootnote{komu  L679.}}. Hróm\textit{undr} sp\textit{yr} \edtext{hv\textit{or}t í}{\Afootnote{i hvórja  L679.}} átt vís\textit{ar} þú o\textit{ss} t\textit{il} þ\textit{ess}a haugs?
K\textit{ar}l s\textit{egir} \edtext{þér}{\Afootnote{þid  L679.}} sculot 
sigla héþ\textit{an} í suþr  \edtext{í}{\Afootnote{÷  L679, L2404.}} vj dægr, o\textit{c} þá manot  \edtext{þ\textit{it}}{\Afootnote{þér  L2404.}} fin\textit{n}a   þyccv\textit{an}\textit{n}  scóg o\textit{c}
dym\textit{m}\textit{an}\textit{n}; v\textit{it} þ\textit{an}\textit{n} scóg
er  \edtext{v\textit{an}t a\textit{t} v\textit{er}a}{\Afootnote{optast  L679.}} kyrt o\textit{c} \edtext{brimleysa}{\Afootnote{brimleisi  L1572.}}, þ\textit{ar} stendr haugr Þráins. Nú f\textit{yrir} þ\textit{at}
s\textit{em} e\textit{c} hefi  sagt þér, þá legg t\textit{il} \edtext{mín}{\Afootnote{þijn  L679, L2404, L1572.}}  góþ orþ, a\textit{t} e\textit{c} nái aptr kýr tetr\textit{om} mín\textit{om}. 
Hróm\textit{undr} g\textit{jỏr}ir sva, a\textit{t} h\textit{ann} kall\textit{ar} t\textit{il} þ\textit{eir}ra  \edtext{er}{\Afootnote{+ med  L1572.}} kýrn\textit{ar} \edtext{teymdo}{\Afootnote{leiddu  L679, fóru  L1572.}}, o\textit{c} seg\textit{ir} þ\textit{ei}m a\textit{t} láta þ\textit{ær}
af hendi þ\textit{ei}m er  ætti,  \edtext{e\textit{n}}{\Afootnote{og  L2404.}} vin\textit{n}a an\textit{n}at   m\textit{ann}dóms   sticci
m\textit{er}cil\textit{e}gra. Þ\textit{ei}r g\textit{jỏr}þo  \edtext{sva}{\Afootnote{÷  L2404.}} s\textit{em} Hróm\textit{undr} villdi v\textit{er}a láta, g\textit{en}gu
nú hv\textit{or}j\textit{ir} \edtext{leiþ\textit{ir} sínar}{\Afootnote{leid sijna  L679, L1572.}}.  Máni k\textit{ar}l  \edtext{q\textit{vo}m}{\Afootnote{for  L1572.}} heim, o\textit{c} þócti vænt om, a\textit{t} friþr v\textit{ar} fengin\textit{n}, e\textit{n} Hróm\textit{undr} o\textit{c}
K\textit{ón}gs m\textit{enn} g\textit{en}go t\textit{il} scipa sin\textit{n}a,
o\textit{c} sỏgþo frá  Mána k\textit{ar}li. 
\pend \endnumbering  

\beginnumbering \pstart  
\vspace{5mm}\subsection*{VI kap\textit{ítuli}} 
Nú undo m\textit{enn} á segl, oc  \edtext{héldo t\textit{il}}{\Afootnote{sigldu i  L1572.}} \edtext{suþors fulla}{\Afootnote{sudr fjalla  L679, sudr full  L1572.}} 6  \edtext{daga}{\Afootnote{dægr  L1572.}} ept\textit{ir} þ\textit{ví} s\textit{em} Máni k\textit{ar}l  \edtext{h\textit{afþi}}{\Afootnote{+ fyrir  L679.}} s\textit{agt} \edtext{þ\textit{ei}m}{\Afootnote{÷  L679.}},  
\edtext{a\textit{t} þ\textit{ei}r sáo}{\Afootnote{sáu þeir  L1572.}} \edtext{m\textit{ỏr}g  an\textit{n}es, o\textit{c} þar ept\textit{ir}}{\Afootnote{÷  L679.}} \edtext{slétt\textit{an}}{\Afootnote{÷  L1572.}} scóg, dym\textit{m}\textit{an} o\textit{c} þyccv\textit{an}\textit{n},
þ\textit{ar} sjá þ\textit{ei}r m\textit{ic}it  rjóþr, o\textit{c} \edtext{n\textit{or}þ\textit{an}}{\Afootnote{nordr  L2404, L1572.}} und\textit{ir} þ\textit{ví} m\textit{ic}in\textit{n} \edtext{o\textit{c}}{\Afootnote{÷  L679, L2404.}} \edtext{ramg\textit{jỏr}þ\textit{an}}{\Afootnote{ramgjörfann  L679, ÷  L2404.}} hól eþr haug, s\textit{em} v\textit{ar} bæþi  hár o\textit{c} fagr. 
Hró\-m\textit{undr} m\textit{ælir} t\textit{il} K\textit{ón}gs m\textit{ann}a: 
hér scul\textit{om} \edtext{vér}{\Afootnote{vid  L679, L2404.}} u\textit{m}búast o\textit{c} sjá o\textit{ss} om, o\textit{c} leggja scip v\textit{or} t\textit{il} \edtext{lægis}{\Afootnote{lags L679, lagis L1572.}}, 
e\textit{n} gaungum á
l\textit{an}d, o\textit{c} leit\textit{om} \edtext{a\textit{t} Þráin\textit{n}}{\Afootnote{Þráins  L679, L2404.}}, þ\textit{ar} h\textit{ann} hefr n\textit{oc}ro a\textit{t}  miþla;  þ\textit{ei}r
qváþo já v\textit{it}, o\textit{c} s\textit{ỏgþo} Hróm\textit{undr} ráþa scyldi.
Þ\textit{ei}r qveþja  K\textit{óng} leyfis o\textit{c}
g\textit{an}ga á l\textit{an}d, þ\textit{ar} t\textit{il} haugurin\textit{n}
nálgast þ\textit{ei}m,  lyst\textit{ir} þá á móti þ\textit{ei}m  \edtext{sva}{\Afootnote{÷  L679, L2404.}} \edtext{m\textit{ic}illi  ólyct}{\Afootnote{fijlu  L679.}},  \edtext{a\textit{t}}{\Afootnote{svo  L679.}} þ\textit{ei}m réþi aptr  a\textit{t} snúa; lét þá Hróm\textit{undr} \edtext{fólc sitt draga}{\Afootnote{draga liþ sitt  L2404.}} \edtext{grímr á sic}{\Afootnote{á sig grijmr  L679.}},  o\textit{c} m\textit{eþ} svodd\textit{an} hætti
q\textit{vo}most þ\textit{ei}r t\textit{il} haugsins, sýndist þ\textit{ei}m 
h\textit{ann} \edtext{þá}{\Afootnote{÷  L679}} allr í ein\textit{om} loga. 
Hróm\textit{undr} b\textit{at} þá e\textit{ck}i géfa  s\textit{ic} 
a\textit{t}  gl\textit{ær}íng\textit{om}  \edtext{þ\textit{ess}\textit{om}}{\Afootnote{þeim  L679.}}\footnote{The word is abbreviated in L679 and the same abbreviation is used three lines above for \enquote{þeim}.}; tóco  \edtext{þá m\textit{enn}}{\Afootnote{menn þá  L1572.}} a\textit{t} \edtext{róta u\textit{m}}{\Afootnote{brióta  L2404.}} \edtext{haugin\textit{n}}{\Afootnote{haugum  L679.}}\footnote{The word is abbreviated in L679 and the expansion \enquote{haugu\textit{m}} is uncertain.}, oc 
||
\ledouternote{111v}%11:43
 fengo ei  \edtext{á leiþis}{\Afootnote{alldeilis  L679.}} q\textit{vo}m\textit{it}, fyrr e\textit{n} á hin\textit{om} \edtext{jv}{\Afootnote{sjótta  L679.}} degi, þá  q\textit{vo}mo þ\textit{ei}r ein\textit{om} glugga á
h\textit{ann}, sáo þ\textit{ei}r  \edtext{þá}{\Afootnote{÷  L2404.}} ein\textit{n}  f\textit{or}kun\textit{n}ar  \edtext{stór\textit{an}}{\Afootnote{vænan  L679.}} B\textit{er}s\textit{er}c, 
h\textit{ann} v\textit{ar} bæþi blár o\textit{c} sv\textit{ar}tr, digr o\textit{c} dramblætisfullr,
h\textit{ann} sat á stóli af rauþa gulli, o\textit{c} v\textit{ar} klæþdor  \edtext{gyllt\textit{om} búnaþi}{\Afootnote{raudum búnadi gilltum  L679.}}; 
f\textit{yrir} nỏs\textit{om} h\textit{an}s stóþ eitt  \edtext{st\textit{or}t}{\Afootnote{fullt  L679, ÷  L2404, L1572.}} glóþ\textit{ar} kér  af
\edtext{log\textit{an}di}{\Afootnote{glöandi L679.}}  eldi, í
þ\textit{an}\textit{n} eld blés Þráin\textit{n} af megni, sva  \edtext{at}{\Afootnote{÷  L679.}} \edtext{logan\textit{n}}{\Afootnote{loga  L679.}} lagþi t\textit{il}  ræfors  om haugin\textit{n}; sv\textit{er}þ  \edtext{hékk}{\Afootnote{fieck  L679.}}\footnote{Dittography in L2404, \enquote{heck} repeated twice.}  scamt  frá
h\textit{ono}m, þ\textit{ví}líkt hỏfþo þ\textit{ei}r  \edtext{alldrei}{\Afootnote{+ fijrr  L679.}} sjeþ, 
þ\textit{at} v\textit{ar} bæþi breiþt  o\textit{c}
l\textit{an}gt, o\textit{c} allt  \edtext{gullin}{\Afootnote{gull  L1572.}} scépt, skálpr  
þ\textit{ess} v\textit{ar} af sylfri, blaþit  \edtext{v\textit{ar}}{\Afootnote{÷  L679.}} af   gresj\textit{ar}ni,
o\textit{c} m\textit{ar}g\textit{ir} \edtext{góþ\textit{ir}}{\Afootnote{÷  L1572.}} gimstein\textit{ar} í smeldir,  \edtext{þ\textit{ei}r}{\Afootnote{+ sem  L1572.}} h\textit{ar}þast\textit{ir} fengist hỏfþo, sva s\textit{em}
dem\textit{an}t o\textit{c} m\textit{ar}m\textit{ar}i, ásamt  \edtext{fleyri}{\Afootnote{fleirum  L679.}},  \edtext{þó}{\Afootnote{+ at  L2404.}} hér séo e\textit{ck}i nefnd\textit{ir}. Þráin\textit{n} h\textit{afþi}  fásjeþ  \edtext{gullmen}{\Afootnote{gull  L679.}} á hálsi sér o\textit{c} b\textit{or}þ m\textit{eþ} ỏll\textit{om} búnaþi
stóþ  f\textit{yrir} knjám hỏn\textit{om} m\textit{eþ}
kræst\textit{om} rétt\textit{om}; h\textit{ann} \edtext{r\textit{ac}}{\Afootnote{+ þá  L679.}} up\textit{p} st\textit{or}t hljóþ  \edtext{sva}{\Afootnote{og  L679.}} m\textit{ic}it  \edtext{o\textit{c}}{\Afootnote{÷  L679.}}  digurt, 
a\textit{t} Hróm\textit{undi}
o\textit{c} \edtext{m\textit{ỏnnom} h\textit{an}s}{\Afootnote{hans mónnum  L679.}}  blỏskraþi, þeg\textit{ar} þ\textit{ei}r  \edtext{bruto}{\Afootnote{rifu  L1572.}} \edtext{gatit   á}{\Afootnote{upp  L679.}} haugin\textit{n}, o\textit{c} byrtuna af degin\textit{om} lagþí  \edtext{í haugin\textit{n}}{\Afootnote{inn  L1572.}}; h\textit{ann} kall\textit{ar} út t\textit{il} K\textit{ón}gs
man\textit{n}a o\textit{c} spyr:  hv\textit{or}j\textit{ir}
e\textit{ro} þ\textit{ei}r   fúlo   f\textit{an}t\textit{ar} 
s\textit{em} g\textit{an}ga  \edtext{á}{\Afootnote{÷  L2404.}} hús  \edtext{mitt}{\Afootnote{min  L1572.}}, o\textit{c} scém\textit{m}a  \edtext{þ\textit{at}}{\Afootnote{÷  L1572.}} f\textit{yrir} mér saclaus\textit{om}, gỏml\textit{om} o\textit{c}
gráh\textit{ær}þ\textit{om}, o\textit{c} lítit hafit
þ\textit{it} t\textit{il} scémtun\textit{ar} \edtext{veslíng\textit{ar}}{\Afootnote{÷  L1572.}} a\textit{t} sæcja á min\textit{n} fund,  eþa hv\textit{or} er
sá þ\textit{ar} úti stendr,  \edtext{s\textit{em}}{\Afootnote{er  L2404.}} \edtext{vill}{\Afootnote{+ þad  L2404.}} t\textit{il} vin\textit{n}a, a\textit{t} q\textit{vo}ma 
in\textit{n} t\textit{il} mín, o\textit{c} njóta þ\textit{eir}ra þriggja
dýrgripa er  \edtext{til}{\Afootnote{eg  L679, L2404.}} \edtext{hefi}{\Afootnote{til  L2404.}} \edtext{betsta}{\Afootnote{best  L679, L2404.}} a\textit{t} bjóþa, s\textit{em} a\textit{t} \edtext{e\textit{ro}}{\Afootnote{er  L679.}}: sv\textit{er}þ\textit{it} Mistiltein\textit{n} o\textit{c} Menit á hálsi mín\textit{om}, ásamt hin\textit{n} \edtext{góþi}{\Afootnote{goda  L679.}} 
\edtext{Gullhríngr}{\Afootnote{gúllhring L679.}}, 
er  e\textit{c} lengst æfi
min\textit{n}\textit{ar} átt hefi, þyci mér  \edtext{lícl\textit{e}gast}{\Afootnote{liklegt  L1572.}}, at  Olafr K\textit{óngr} egi sva á m\textit{ỏnnom} vỏl,
a\textit{t} einhv\textit{or} t\textit{il} reyni. Hróm\textit{undr} m\textit{ælir} þ\textit{at} máttú f\textit{yrir} satt halda,
a\textit{t} ef Kári hefþi lifat,  
mundi  \edtext{e\textit{c}}{\Afootnote{hann  hafa L679, hann L2404, L1572.}}\footnote{Note that better readings are in L679, L2404, and L1572 than in B11109.}  \edtext{t\textit{il}}{\Afootnote{÷  L1572.}}  vogat   \edtext{h\textit{afa}}{\Afootnote{÷  L679.}}, a\textit{t} fin\textit{n}a þ\textit{ic}, þ\textit{ar} s\textit{em} þú
hefr  \edtext{gỏfugr hỏfþíngi}{\Afootnote{göfigr maþr  L2404.}} haldin\textit{n} v\textit{er}it, o\textit{c} \edtext{sýna ⸌fáir⸍}{\Afootnote{fáir sijna  L679, L2404, L1572.}} svoþdan  \edtext{hýbýla}{\Afootnote{ + prídi og  L1572.}}
||
\ledouternote{112r} 
rausn   \edtext{o\textit{c}  príþiheit}{\Afootnote{÷  L2404, L1572.}}, sva s\textit{em} hér er a\textit{t} scoþa, o\textit{c}  sac\textit{ir}   tignar   þin\textit{n}\textit{ar},
manto  \edtext{leyfa}{\Afootnote{lofa  L679, L2404.}} mér a\textit{t} q\textit{vo}ma f\textit{yrir} kné  \edtext{yþ\textit{ar}}{\Afootnote{þijn  L679, L1572.}}, sva  e\textit{c} megi ei  ofsỏg\textit{om} \edtext{af segja}{\Afootnote{segja af  L679.}} þin\textit{n}i géstrisni.
Þráin\textit{n} m\textit{ælir} fleyri sóctu m\textit{ic} heim f\textit{or}þ\textit{om}, þeg\textit{ar} Lxij m\textit{enn}  veitto   mér   ásteytíng 
m\textit{eþ} \edtext{ráni o\textit{c} róti}{\Afootnote{röti og ráni  L679.}} á \edtext{hús kofa}{\Afootnote{húsum  L679.}}  mín\textit{om},  \edtext{o\textit{c}}{\Afootnote{en  L2404.}} fengo  \edtext{þ\textit{ei}r}{\Afootnote{þo  L679, L2404.}} lít\textit{it} t\textit{il} ábata. V\textit{er}a má s\textit{egir}
Hróm\textit{undr} a\textit{t} nú  t\textit{ac}ist betr
t\textit{il}, o\textit{c} vyrþist  \edtext{rífl\textit{e}gra}{\Afootnote{riflega  L679.}} \edtext{a\textit{t} g\textit{jỏr}a v\textit{el} v\textit{it} ein\textit{n} gést}{\Afootnote{vid eirn giest ad gjóra  L679.}},  
heldr e\textit{n} \edtext{greiþa}{\Afootnote{giora  L1572.}} kostnaþ sva m\textit{ỏr}g\textit{om}, o\textit{c} skal e\textit{c} nú  fyrstr t\textit{il} v\textit{er}þa; lét þá Hróm\textit{undr} binda
s\textit{ic} í festi,  \edtext{o\textit{c}}{\Afootnote{÷  L679.}} sje  sva niþr í haugin\textit{n}, o\textit{c} b\textit{at} \edtext{áþr}{\Afootnote{÷  L679.}} m\textit{enn} sína  \edtext{a\textit{t}}{\Afootnote{ádr  L679.}} gjæta v\textit{el} fest\textit{ar}in\textit{n}\textit{ar},
o\textit{c} e\textit{ck}i frá ren\textit{n}a,  \edtext{þó}{\Afootnote{+ at  L2404.}} þ\textit{ei}r  \edtext{kyn\textit{n}o a\textit{t} heyra}{\Afootnote{heirdi  L679.}} noccot  vofeifl\textit{e}gt,  \edtext{a\textit{t} þ\textit{ei}r}{\Afootnote{og  L679.}} léto sér þ\textit{at} ei blỏskra. 
En\textit{n} s\textit{em} Hróm\textit{undr} kémr í haugin\textit{n}, v\textit{er}þr h\textit{ann} þá sva  \edtext{dym\textit{m}r}{\Afootnote{+ og fullr  L679, L2404, L1572.}} af \edtext{ógérþ\textit{ar}}{\Afootnote{÷  L1572.}} fýlo  \edtext{o\textit{c} reyc}{\Afootnote{÷  L1572.}}, a\textit{t} \edtext{Hróm\textit{undr}}{\Afootnote{hann  L1572.}} v\textit{er}þr a\textit{t}  þreifa   f\textit{yrir} sér,
fin\textit{n}r h\textit{ann} þá m\textit{ar}ga  \edtext{góþa dýrgripi og}{\Afootnote{dijrgripi og goda  L679, L2404, L1572.}} gullhrínga. Hróm\textit{undr} batt  \edtext{þ\textit{at}}{\Afootnote{fed  L1572.}} í festina, o\textit{c} \edtext{lét}{\Afootnote{+ þad  L679.}} uppdraga, f\textit{an}\textit{n} h\textit{ann} \edtext{þ\textit{ar}}{\Afootnote{+ var  L2404.}}  ofr  gnægþ   
m\textit{ar}gra  \edtext{fásjeþra}{\Afootnote{fásiedna  L679, fáséþnra  L2404.}} \edtext{hluta}{\Afootnote{gripa  L2404.}}. 
\pend 
\endnumbering 

\beginnumbering 
\pstart  
\vspace{5mm}\subsection*{VII kap\textit{ítuli}} 
Nú s\textit{em} Hróm\textit{undr} v\textit{ar} a\textit{t} láta gripina í festina, tóc  Þráin\textit{n} t\textit{il} a\textit{t} mæla o\textit{c} s\textit{egir} st\textit{or} hugaþr erto  \edtext{Hróm\textit{undr}}{\Afootnote{÷  L679.}} at  grípa  \edtext{f\textit{rá}}{\Afootnote{fyrir  L679, L2404.}} mér þ\textit{at} e\textit{c} á, v\textit{er}þ  \edtext{e\textit{c}}{\Afootnote{÷  L2404.}} nú a\textit{t}  gjalda   elli min\textit{n}\textit{ar}, e\textit{n} þeg\textit{ar} e\textit{c} v\textit{ar} úngr, hefþ\textit{ir} þú e\textit{ck}i rænt m\textit{ic} \edtext{sva}{\Afootnote{÷  L679, L1572.}} stór\textit{om} þegar e\textit{c} van\textit{n} allt Vall\textit{an}d m\textit{eþ} \edtext{fjỏlcíngi o\textit{c} gỏldr\textit{om}}{\Afootnote{gólldrum og fiólkingi  L679.}}, o\textit{c} m\textit{ar}ga  K\textit{on}ga o\textit{c}
K\textit{ónga}syni  \edtext{lagþi e\textit{c} a\textit{t} velli}{\Afootnote{hef eg ad velli lagt  L679.}}, e\textit{n} þeg\textit{ar} á m\textit{ic} f\textit{ær}þist  \edtext{alldurin\textit{n}}{\Afootnote{ellinn  L679.}} létst e\textit{c} deya, v\textit{ar} e\textit{c} lifandi í þen\textit{n}\textit{an}\textit{n} haug látin\textit{n},  o\textit{c} síþ\textit{an} hef e\textit{c} \edtext{hér}{\Afootnote{÷  L679.}} lifat v\textit{it}  sjón á grip\textit{om} \edtext{þ\textit{ess}\textit{om}}{\Afootnote{mijnum  L679, L2404, minum þessum  L1572.}} s\textit{em} þú \edtext{hripsar}{\Afootnote{+ nu  L679.}} frá mér, s\textit{em} an\textit{n}\textit{ar} þjófr. Hróm\textit{undr} m\textit{ælir} fleyr⸌a⸍ v\textit{er}þr a\textit{t} \edtext{vin\textit{n}a}{\Afootnote{gióra  L679, L1572.}} e\textit{n} gott  þyc\textit{ir},  
\edtext{o\textit{c}}{\Afootnote{÷  L679, L2404, L1572.}} sva  \edtext{f\textit{er}}{\Afootnote{verþr  L2404.}} mér, þyc\textit{ir} sá laung\textit{om} \edtext{no\textit{ck}ro}{\Afootnote{÷  L679.}}  nýtor 
||
\ledouternote{112v} 
 \edtext{er}{\Afootnote{sem  L2404, L1572.}} vin\textit{n}r t\textit{il} fæþis. Þá m\textit{ælir} Þráin\textit{n}: haf
þú e\textit{ck}i spott  \edtext{í}{\Afootnote{÷  L679.}} fram\textit{m}i v\textit{it} m\textit{ic} gamlan\textit{n}, þó  \edtext{a\textit{t}}{\Afootnote{÷  L679, L1572.}} þú sért úngr á bein\textit{om}, þá  þ\textit{or}i e\textit{c}
samt a\textit{t} ⸌kalla þig þjóf a\textit{t}⸍ gótsi míno. 
Þesso reiþdist Hróm\textit{undr} oc ljóp a\textit{t} Þráin\textit{n}, e\textit{n} \edtext{h\textit{ann}}{\Afootnote{Þráinn  L679.}} t\textit{óc} \edtext{á}{\Afootnote{i  L679, L1572.}} móti af megni. Þá m\textit{ælir} Hróm\textit{undr} nú þ\textit{ar}f e\textit{ck}i f\textit{yrir} a\textit{t} þreifast, o\textit{c} má sér  \edtext{e\textit{n} þá}{\Afootnote{÷  L679, L2404.}} no\textit{ck}ot hússbóndin\textit{n}. Já, s\textit{egir} Þráin\textit{n}, þó a\textit{t} e\textit{c} sé orþin\textit{n} gamall, þá scalto  e\textit{ck}i,  \edtext{búo}{\Afootnote{bú  L679, búe  L2404, ÷  L1572.}} \edtext{hundurin\textit{n} þin\textit{n}}{\Afootnote{hundr  L679, L2404, ÷  L1572.}}! 
b\textit{er}a fé frá mér o\textit{c} ræna m\textit{ic} \edtext{at  gamni þíno}{\Afootnote{÷  L2404.}}, o\textit{c} v\textit{el} f\textit{er} \edtext{þ\textit{at}}{\Afootnote{ad  L679, þat at  L2404, L1572.}} þú kaup\textit{ir} fullo, a\textit{t} þú hefr  \edtext{mic heimsóct}{\Afootnote{fúndid mig  L679.}}, o\textit{c}  ómaca\textit{t}  m\textit{ic}
or sæti míno, o\textit{c} víst scalto mega  sjá v\textit{it} mér
dauþ\textit{om}, þó a\textit{t} þú sért lif\textit{an}di, o\textit{c}
t\textit{il} no\textit{ck}ors vilda e\textit{c} þér kjæmi fjỏldi
gullhrínga min\textit{n}a, s\textit{em} þér meþ  sv\textit{er}þi
míno mest í augo g\textit{an}ga. Hróm\textit{undr} \edtext{s\textit{egir}}{\Afootnote{mælir  L679, L2404, L1572.}} v\textit{it} scul\textit{om} þessi   f\textit{an}gbrỏgþ  leica
f\textit{yrir} ut\textit{an} sv\textit{er}þ o\textit{c} scjỏldo. Ec  \edtext{ætli}{\Afootnote{ætla  L679, L2404, L1572.}} sva mani  betst f\textit{ar}a s\textit{egir} Þráin\textit{n},
a\textit{t} \edtext{þér}{\Afootnote{ei  L1572.}} mani lítil fremd  \edtext{í}{\Afootnote{÷  L679, L2404.}} a\textit{t} vega  a\textit{t} mér; t\textit{óc} þá
Þráin\textit{n} a\textit{t} blása eldi or kétil kérino s\textit{em} stóþ f\textit{yrir} h\textit{ono}m,  \edtext{tindraþist}{\Afootnote{tindradi  L679.}} þá  \edtext{login\textit{n}}{\Afootnote{logann  L679.}} \edtext{á}{\Afootnote{i  L2404.}} \edtext{mill\textit{om}}{\Afootnote{milli  L679.}} þ\textit{eir}ra  \edtext{o\textit{c}}{\Afootnote{÷  L679.}} 
om  haugin\textit{n}, sva byrti,
þá sá Hróm\textit{undr} hv\textit{ar} st\textit{or} kétill stóþ  \edtext{f\textit{yrir} fót\textit{om} sér, o\textit{c} v\textit{ar}}{\Afootnote{÷  L2404.}} fullr af  búk\textit{om}  m\textit{ann}a; þat þóctist  \edtext{h\textit{an}n}{\Afootnote{Hromundr  L679.}} vita a\textit{t} \edtext{Þráin\textit{n}}{\Afootnote{hann  L679.}} m\textit{un}di vilja kippa sér í  \edtext{þen\textit{n}an\textit{n} kétil}{\Afootnote{hann  L679.}},  héldost  \edtext{þ\textit{ei}r}{\Afootnote{+ þá  L679.}} í henþr,  \edtext{o\textit{c} toguþost}{\Afootnote{og togast  L1572, ÷  L679.}} yf\textit{ir} \edtext{om}{\Afootnote{÷  L679.}} kétilin\textit{n}, 
þá  fan\textit{n} Hróm\textit{undr}
a\textit{t} Þráin\textit{n} h\textit{afþi} \edtext{e\textit{ck}i}{\Afootnote{÷  L2404.}} \edtext{nýl\textit{e}ga}{\Afootnote{÷  L679, L2404.}} \edtext{sc\textit{or}it af sér}{\Afootnote{af ser skorid  L1572, ÷  L2404.}} \edtext{nỏglurn\textit{ar}}{\Afootnote{+ þvj  L679, neglur  L2404.}}, þ\textit{ær} krepptost of\textit{an} f\textit{yrir} gómana. Þá
m\textit{ælir} Þráin\textit{n}:  lofa manto mér a\textit{t}
t\textit{ac}a mér mun\textit{n}bita  \edtext{áþr}{\Afootnote{+ enn  L679, L2404, L1572.}} v\textit{it} glímom. V\textit{er}a má þ\textit{at}
s\textit{egir} \edtext{Hróm\textit{undr}}{\Afootnote{+ og  L679.}} J þ\textit{ví} greip Þráin\textit{n} kétilin\textit{n}
o\textit{c}  fleygþi   a\textit{t} Hróm\textit{undi}
e\textit{n} h\textit{ann} brá  \edtext{v\textit{it} o\textit{c}}{\Afootnote{sér  L679.}} \edtext{und\textit{an}; í þ\textit{ví} ljóp}{\Afootnote{liop i þvi undan þá veþr  L2404, undan oc i þvi hliop  L1572.}} Þr\textit{áinn} a\textit{t} Hróm\textit{undi} o\textit{c} \edtext{m\textit{ælir}}{\Afootnote{÷  L2404.}} nú man mál a\textit{t} géfa upp a\textit{t} b\textit{or}þa, o\textit{c} g\textit{an}ga t\textit{il} vin\textit{n}o, þá dym\textit{m}di  \edtext{sva}{\Afootnote{+ ad  L679.}} \edtext{koldỏct}{\Afootnote{koldimt  L679.}} v\textit{ar}þ í haugn\textit{om}; þá  fleygþi
Hróm\textit{undr} sv\textit{er}þino, oc vildi   liþugr   v\textit{er}a. Þetta sá 
||
\ledouternote{113r} 
 Þráin\textit{n} o\textit{c}  kippti  kỏmp\textit{om}
o\textit{c} ljóp a\textit{t} Hróm\textit{undi}, o\textit{c} f\textit{ær}þi  \edtext{han\textit{n}}{\Afootnote{÷  L2404.}} í  lopt upp, e\textit{n} h\textit{ann} q\textit{vo}m 
\edtext{fót\textit{om} kjænl\textit{e}ga}{\Afootnote{kjænlega fötum  L679.}} und\textit{ir} s\textit{ic}, 
sva h\textit{ann} féll  eigi,
e\textit{n} t\textit{óc} þó allst\textit{er}cl\textit{e}ga á móti.  
\edtext{Þá m\textit{ælir} Þráin\textit{n}}{\Afootnote{Þráinn mælir  L679.}}: e\textit{ck}i  erto  \edtext{m\textit{enn}scr m\textit{aþr}}{\Afootnote{menskum manni lijkr  L679.}} a\textit{t} þú stendr. 
Þúsund  \edtext{m\textit{enn}ri}{\Afootnote{meiri  L679, L1572.}} ertú,  s\textit{egir} Hróm\textit{undr} o\textit{c} hefr þú
frægr m\textit{aþr} v\textit{er}it á  \edtext{úng\textit{om}}{\Afootnote{unga  L679.}} alldri. Litl\textit{ar} sjást\footnote{An uncertain reading in L1572: \enquote{tiást} or \enquote{fiást}.}  þ\textit{ess} menj\textit{ar}, s\textit{egir}
Þráin\textit{n}; 
\edtext{t\textit{óc} þá}{\Afootnote{tok hann þá L679, toku þeir  L2404.}} 
\edtext{Þráin\textit{n}}{\Afootnote{÷ L679.}}
 at  f\textit{ær}ast í auca ablssíns, sva Hróm\textit{undr} þóktist þurfa  a\textit{t} neyta orco sin\textit{n}\textit{ar}, gé\textit{ck} þá  \edtext{upp allt hv\textit{aþ}}{\Afootnote{upp þat  L2404, allt upp er  L1572.}} \edtext{f\textit{yrir}}{\Afootnote{+ þeim  L679.}} v\textit{ar}d bæþi  tré o\textit{c} grjót,  \edtext{ásamt aþr\textit{ir} lut\textit{ir}, hv\textit{or}jo nafni s\textit{em} h\textit{ie}to}{\Afootnote{÷  L2404, L1572.}}, v\textit{or}o  þá h\textit{ar}þ\textit{ar} \edtext{sveiflr}{\Afootnote{sviptingar  L1572.}},  \edtext{e\textit{n}}{\Afootnote{og  L679, L2404.}} \edtext{stund\textit{om}}{\Afootnote{÷  L2404.}} 
hriggspen\textit{n}r, ýmist  blés Þráin\textit{n} eldi
eþr eitraþri spýu úr kjapti sér, 
o\textit{c} stund\textit{om} héldt h\textit{ann} tỏn\textit{n}\textit{om} \edtext{í}{\Afootnote{yfir  L679.}} axl\textit{ir} Hróm\textit{undar} 
en\textit{n} stund\textit{om} reif h\textit{ann} \edtext{m\textit{eþ} nỏgl\textit{om} s\textit{em} kỏttr}{\Afootnote{sem kóttr med nóglunum  L679.}}; 
v\textit{it} þetta  
h\textit{er}dti  Hróm\textit{undr} sva  \edtext{h\textit{ono}m}{\Afootnote{sic enn  L2404.}} ⸌tóc⸍ lítt  \edtext{a\textit{t} fin\textit{n}ast t\textit{il}}{\Afootnote{til at finnaz  L2404.}}  átekta   \edtext{Þráins}{\Afootnote{þeira  L679.}},  lét h\textit{ann} þá draugin\textit{n} sæcja, e\textit{n}  v\textit{ar}þist   sjálfr   fỏll\textit{om}. 
Þat  fan\textit{n} Þráin\textit{n}, o\textit{c} m\textit{ælir}
betr mátto  \edtext{adhafast}{\Afootnote{hafast  L679.}}, ef þú  \edtext{scalt}{\Afootnote{villt  L1572.}} fá gull mitt. Já, s\textit{egir} Hróm\textit{undr} o\textit{c} brá
v\textit{it} h\textit{ar}dt,  \edtext{sva}{\Afootnote{+ at  L2404.}} Þráin\textit{n} féll á  \edtext{an\textit{n}at knéþ}{\Afootnote{bædi knien  L679.}}. 
St\textit{er}cr m\textit{aþr} erto s\textit{egir} Þráin\textit{n}, at
fella m\textit{ic} \edtext{á}{\Afootnote{÷  L679.}} ỏþr\textit{om} fæti, 
\edtext{gaml\textit{an}\textit{n} o\textit{c} styrþ\textit{an}}{\Afootnote{stirþan oc gamlan  L2404.}} frá  \edtext{eldin\textit{om}}{\Afootnote{elli órum  L1572.}}\footnote{
	An error in L1572, \enquote{ór\textit{um}} probably fo  \enquote{árum}, rather than \enquote{örum}. The scribe normally uses \enquote{ỏ} for \enquote{ö}.}, 
o\textit{c} man láta a\textit{t} líc\textit{om} a\textit{t} qvỏlda taci. 
Statto upp aptr  s\textit{egir} Hróm\textit{undr} o\textit{c}
tvisv\textit{ar} v\textit{er}þr gamall m\textit{aþr} b\textit{ar}n,
o\textit{c} sva fer  þér nú, 
o\textit{c} aldrei  \edtext{hefþi}{\Afootnote{hef  L679.}} e\textit{c} m\textit{ic} \edtext{híngat}{\Afootnote{÷  L679.}} ómacat v\textit{it} \edtext{b\textit{ar}n}{\Afootnote{bórn  L679, L2404, L1572.}} a\textit{t} glíma, hefþi e\textit{c} þetta vitat, o\textit{c} hefr
Máni k\textit{ar}l ílla  \edtext{a\textit{t} mér logit}{\Afootnote{logied ad mier  L679.}}, a\textit{t} þú v\textit{ær}ir h\textit{it} v\textit{er}sta galdra trỏll.
V\textit{er}a  má a\textit{t} \edtext{ei}{\Afootnote{÷  L679.}} sé allt log\textit{it}, s\textit{egir} Þráin\textit{n}, e\textit{n} heyri
e\textit{c} a\textit{t} \edtext{fry\textit{ar} þú}{\Afootnote{þú frijar  L679, L2404, L1572.}} \edtext{mér}{\Afootnote{+ nu  L1572.}}, o\textit{c} skal nú uppst\textit{an}da,  \edtext{m\textit{eþ} þ\textit{ví}}{\Afootnote{medann  L1572.}} \edtext{á}{\Afootnote{÷  L679, L1572.}} \edtext{dagin\textit{n}}{\Afootnote{dagr  L679.}} er  \edtext{liþit}{\Afootnote{lidinn  L679, L1572.}},\footnote{A grammatical error in B11109 and probably also in L2404. Note that the word is abbreviated in L2404 as \enquote{liþ\textit{it}}, while the abbreviations in L679 and L1572 are straightforward with a nasal stroke for \enquote{lidin\textit{n}}.}  \edtext{o\textit{c} dym\textit{m}a t\textit{ec}or í haugi mín\textit{om}}{\Afootnote{÷  L679.}}. 
Af þ\textit{ess}o þóttist Hróm\textit{undr} 
||
\ledouternote{113v}
 ráþa, a\textit{t} Þráin\textit{n} m\textit{un}di  \edtext{v\textit{er}ri v\textit{er}a}{\Afootnote{vera verri  L679.}} u\textit{m} nætr e\textit{n} \edtext{u\textit{m}}{\Afootnote{÷  L2404.}} daga, o\textit{c} t\textit{il} þ\textit{ess} m\textit{un}di h\textit{ann} tím\textit{an}\textit{n}
dreg\textit{it} h\textit{afa},  \edtext{t\textit{ac}ast}{\Afootnote{taka  L679.}} þ\textit{ei}r nú  \edtext{á aptr}{\Afootnote{aptr til glijmú  L679.}} engo  síþr e\textit{n} fyrri, sva allt   hlaut   upp a\textit{t}
g\textit{an}ga  \edtext{þ\textit{at}}{\Afootnote{sem  L2404.}} f\textit{yrir} v\textit{ar}þ,   gnísti   þá Þráin\textit{n} 
\edtext{tỏn\textit{n}\textit{om}}{\Afootnote{+ og L679.}}, 
beit o\textit{c} \edtext{reif og}{\Afootnote{÷  L679.}}  beljaþi   s\textit{em} graþúngor,  
sva ỏskr h\textit{an}s heyrþo þ\textit{ei}r s\textit{em} úti hjá staþd\textit{ir} v\textit{or}o,  
skálf þá allr haugorin\textit{n}, s\textit{em} á h\textit{or}þræþi léci, yf\textit{ir}gáfo
þ\textit{ei}r þá festina  af ótta, o\textit{c} villdo   f\textit{or}þa   \edtext{sjálf\textit{om}}{\Afootnote{÷  L679, L2404.}} sér,  
\edtext{t\textit{óc}}{\Afootnote{+ þá  L679, L2404.}} a\textit{t} q\textit{vo}ma  \edtext{a\textit{t}}{\Afootnote{÷  L2404.}} þ\textit{ví}, a\textit{t} \edtext{eptir v\textit{ar} no\textit{ck}ot}{\Afootnote{nockúd var eptir  L679.}} hjá \edtext{Þráin\textit{n}}{\Afootnote{Þráini  L2404.}},  \edtext{t\textit{óc} h\textit{ann} þá a\textit{t} géfa}{\Afootnote{gaf hann þá  L2404.}} Hróm\textit{undi} st\textit{or} hỏgg oc pústra, 
setti h\textit{ann} þá  nỏglr sín\textit{ar} á   kaf   \edtext{v\textit{it}}{\Afootnote{i  L2404.}} háls h\textit{an}s o\textit{c} reif   hold   frá  \edtext{bein\textit{om}}{\Afootnote{beini  L679.}} allt  \edtext{á lend\textit{ar}}{\Afootnote{ad lendum  L679.}} niþr. Þá m\textit{ælir} Hróm\textit{undr} þ\textit{at} finn e\textit{c} a\textit{t} \edtext{klóra kan\textit{n} piltkindin}{\Afootnote{pilltkindinn kann ad klora  L679, klóra kant þú  L2404, L1572.}}\footnote{The feminine noun \textit{piltkind} is registered in OH from the 19th to the 20th cent.}  \edtext{m\textit{ỏnnom}}{\Afootnote{mier  L679.}} á b\textit{ac}ino o\textit{c} e\textit{ck}i  sljófgast  \edtext{hỏnom}{\Afootnote{þér  L2404, L1572.}} fíngurn\textit{ar}. 
Segþo t\textit{il} s\textit{egir} Þráin\textit{n}, ef
e\textit{c} skal betr g\textit{jỏr}a, þ\textit{ví} e\textit{ck}i 
man  \edtext{blána né blæþa}{\Afootnote{blæda nie blána  L679.}}, o\textit{c} ósópaþ\textit{ar} e\textit{ro} \edtext{þín\textit{ar} síþr báþar}{\Afootnote{badar sijdr þijnar  L679, sidr þinar badar  L1572.}},  \edtext{elleg\textit{ar}}{\Afootnote{eda  L679.}} kinok\textit{ar}  þú þér,  \edtext{a\textit{t} þú}{\Afootnote{og  L679.}} f\textit{ær}ist  \edtext{und\textit{an}, oc hnickist}{\Afootnote{÷  L679.}} or f\textit{an}gi  \edtext{mér}{\Afootnote{mijnu  L679.}}, o\textit{c} þ\textit{at} segi e\textit{c} þér, a\textit{t} e\textit{ck}i
mátto  krimpast  \edtext{v\textit{it} þ\textit{at}}{\Afootnote{÷  L679.}}, þó  e\textit{c} kun\textit{n}i a\textit{t} hafa  einstỏko nỏgl á fíngrom  \edtext{mín\textit{om}}{\Afootnote{÷  L679.}}; en\textit{n} ei  a\textit{t} síþr skalto nú  \edtext{e\textit{ck}i}{\Afootnote{÷  L679, L1572.}} héþ\textit{an} laus v\textit{er}þa, þ\textit{ví} nú skal e\textit{c}
þ\textit{ic}  qvican\textit{n} í sundr rífa  taug   frá taug. 
Hróm\textit{undr}  þeinkti  \edtext{m\textit{eþ} sér}{\Afootnote{÷  L679.}}, a\textit{t} \edtext{þetta eina m\textit{un}di Þráin\textit{n}}{\Afootnote{Þráinn mundi þetta  L679.}} satt s\textit{egja}  h\textit{er}þtist  \edtext{h\textit{ann}}{\Afootnote{+ þá  L679.}} af  \edtext{stórri}{\Afootnote{mikillri  L679.}} bræþi, o\textit{c} m\textit{ælir}
t\textit{il} Þráins: v\textit{ar}la b\textit{er} e\textit{c} vit  \edtext{t\textit{il} o\textit{c} skynsemd}{\Afootnote{til  L2404, oc skinsemd til  L1572.}},  a\textit{t} þú, s\textit{em} K\textit{óngr} hefr heit\textit{it},
skul\textit{ir} \edtext{orþin\textit{n} v\textit{er}a}{\Afootnote{vera ordinn  L679, L2404.}} a\textit{t} v\textit{er}sta kétti, oc  \edtext{er}{\Afootnote{+ þier  L679.}} skỏm\textit{m} a\textit{t} \edtext{þér}{\Afootnote{þuj  L679.}} gráh\textit{ær}þ\textit{om} k\textit{ar}li, a\textit{t} g\textit{jỏr}ast
slíc  \edtext{fj\textit{an}d\textit{an}s}{\Afootnote{÷  L2404.}} \edtext{fordæþa}{\Afootnote{fordædi  L679.}}, o\textit{c} mantú sonr trỏllkono  \edtext{s\textit{em}}{\Afootnote{er  L679, L1572.}} Gun\textit{n}lỏþ h\textit{ie}t, 
hún átti  \edtext{b\textit{ỏr}n m\textit{ỏr}g}{\Afootnote{mórg bórn  L679.}}, o\textit{c} v\textit{or}o þau  \edtext{ỏll}{\Afootnote{allt  L679, L1572.}}  blauþ\textit{ir}  kétt\textit{ir},
o\textit{c} fin\textit{n} e\textit{c} a\textit{t} þú ert  ein
kétta, þ\textit{ví} þ\textit{at} sá e\textit{c} áþ\textit{an}\textit{n},
a\textit{t} þú  v\textit{er}md\textit{ir} 
þ\textit{ic} á milli fót\textit{an}a,  þ\textit{ví} þú satst
m\textit{eþ} kétilin\textit{n} í klofino, e\textit{n} ert orþin  \edtext{hundgỏmul}{\Afootnote{hundgamall  L679.}}. 
V\textit{it} \edtext{orþ þessi}{\Afootnote{þessi ord  L679.}} reiþdist Þráin\textit{n}, o\textit{c} t\textit{óc} a\textit{t} seilast m\textit{eþ} krócana út á  \edtext{síþurn\textit{ar}}{\Afootnote{+ og  L679.}}; í þ\textit{ví} brá Hróm\textit{undr} hỏn\textit{om} hælkróc, sva  h\textit{ann} \edtext{datt}{\Afootnote{fiell  L679.}} aptr á bac, þá blés Þráin\textit{n} mæþilega, e\textit{n} haugin\textit{n} 
||
\ledouternote{114r} 
\edtext{fyllti}{\Afootnote{filltist  L679.}} af ódaun  \edtext{o\textit{c} reyc}{\Afootnote{÷  L2404.}}.  \edtext{Fellr}{\Afootnote{so fellr  L1572.}} hv\textit{or} þó fræcin\textit{n} sé, s\textit{egir} Þráin\textit{n}. Skjald\textit{an} heyrist þ\textit{at}, s\textit{egir}
Hróm\textit{undr} a\textit{t} kỏtt\textit{om} v\textit{er}þi hálk\textit{an}
a\textit{t} bilto meini, o\textit{c} v\textit{er}þr einosin\textit{n}i allt
fyrst; t\textit{óc} þá Þráinn  a\textit{t} brjótast om o\textit{c}
blása eldi or  \edtext{gini}{\Afootnote{kjapti  L679.}} sér, sva no\textit{ck}ot glampaþi om haugin\textit{n}; miþt í
þ\textit{ess}o sá Hróm\textit{undr} a\textit{t} þ\textit{ar} s\textit{em}
Þráin\textit{n} lá, a\textit{t} súlo  stoþin  v\textit{ar} \edtext{hjá}{\Afootnote{+ þar  L679, + þeim  L2404, þar hjá  L1572.}} \edtext{s\textit{em}}{\Afootnote{er  L1572.}} Mistiltein\textit{n} \edtext{á he\textit{ck}}{\Afootnote{hieck á L679.}}, h\textit{ann} \edtext{bregþst}{\Afootnote{brá  L2404.}} \edtext{scjótt   v\textit{it}}{\Afootnote{vid skiott  L1572.}}, o\textit{c}  þrífr   sv\textit{er}þ\textit{it} á  \edtext{lopt}{\Afootnote{lopti  L679.}} af mesta fimleika, o\textit{c} \edtext{hendti}{\Afootnote{+ sverdid  L1572.}} á lopti, e\textit{n} Þráin\textit{n} v\textit{ar} sein\textit{n}
a\textit{t} st\textit{an}da á fætr, en\textit{n} hafþi  \edtext{þó læst\textit{ar}}{\Afootnote{þá læst  L2404.}} hendr í klæþa  slitri  \edtext{og}{\Afootnote{á  L2404.}} síþ\textit{om} Hróm\textit{undar} \edtext{Hróm\textit{undr}}{\Afootnote{og  L679.}} hjó þá af Þráin\textit{n} báþ\textit{ar} hendurn\textit{ar}, o\textit{c}
héngo  þ\textit{ær} \edtext{báþ\textit{ar}}{\Afootnote{÷  L679.}} í síþ\textit{om} Hróm\textit{undar} Þá m\textit{ælir} Þráin\textit{n}:  gjæfo munr   \edtext{v\textit{ar}þ nú}{\Afootnote{var  L679.}} m\textit{eþ} \edtext{o\textit{ck}or}{\Afootnote{yckr  L679.}}, a\textit{t} þú náþ\textit{ir} sv\textit{er}þi  \edtext{míno}{\Afootnote{+ eij  L679.}}, 
o\textit{c} ætlaþi e\textit{c} þ\textit{at} \edtext{aldrei, a\textit{t} þú}{\Afootnote{÷  L679.}}, min\textit{n} góþi  \edtext{Mistiltein\textit{n}}{\Afootnote{+ at þu  L679.}}! mund\textit{ir} mér a\textit{t} \edtext{meini}{\Afootnote{bana  L679.}} v\textit{er}þa, o\textit{c} er þ\textit{ví} alldrei gott a\textit{t}
treysta á  \edtext{gripi}{\Afootnote{+ sijna  L679.}} o\textit{c} g\textit{er}sem\textit{ar} \edtext{sín\textit{ar}}{\Afootnote{÷  L679.}}, 
o\textit{c} san\textit{n}ast þ\textit{at} nú á mér. Þá sp\textit{urdi}
Hróm\textit{undr} hv\textit{aþ} m\textit{ar}ga hefr þú í \edtext{einvígi}{\Afootnote{einvigum  L1572.}}  sigraþ   m\textit{eþ} þ\textit{ess}o
sv\textit{er}þi? Þr\textit{áinn} m\textit{ælir} J  C  \edtext{hólmgaungr}{\Afootnote{holmgaungum  L1572.}} hefi e\textit{c} þ\textit{at} bor\textit{it}, o\textit{c} \edtext{fé\textit{ck}}{\Afootnote{+ eg  L679, fengit  L1572.}} aldrei eitt sar á  líkama min\textit{n}; xxiv K\textit{on}ga
hjó e\textit{c} t\textit{il} b\textit{an}a m\textit{eþ} þ\textit{ví},
o\textit{c} þ\textit{at} \edtext{sagþi}{\Afootnote{+ mer  L2404.}} Semíngr  \edtext{K\textit{óngr} mér}{\Afootnote{Bersercr (Kongr)  L2404.}} \edtext{v\textit{it} Svíþjóþ}{\Afootnote{÷  L679.}}, þá e\textit{c} v\textit{an}\textit{n} \edtext{af hỏn\textit{om} þetta 	sv\textit{er}þ}{\Afootnote{þetta sverd af honum vid Sviþjod  L679.}}, a\textit{t} e\textit{c} m\textit{un}di  \edtext{jafn\textit{an}}{\Afootnote{÷  L2404.}}  sigr   \edtext{fá}{\Afootnote{+ á  L679.}}, m\textit{ed}an e\textit{c} á þ\textit{ví} héldi. Hr\textit{ómundr} m\textit{ælir} e\textit{c} held þ\textit{at} \edtext{betsta}{\Afootnote{mesta  L679, L2404.}} lukko v\textit{er}c a\textit{t} drepa þ\textit{ic}, m\textit{eþ} þ\textit{ess}o sv\textit{er}þi  o\textit{c} ept\textit{ir} þetta
hjó Hróm\textit{undr} hỏfuþ\textit{it} af  \edtext{Þráin\textit{n}}{\Afootnote{Þráni  L2404.}} o\textit{c} brendi h\textit{ann} síþ\textit{an} \edtext{á báli upp t\textit{il}}{\Afootnote{ad  L679.}} ỏsko,  \edtext{v\textit{ar}þ}{\Afootnote{÷  L679, + nu  L2404, + þá  L1572.}} \edtext{bj\textit{ar}t í haugn\textit{om}}{\Afootnote{÷  L679.}},  \edtext{þ\textit{ar}}{\Afootnote{÷  L679, þá  L1572.}} \edtext{t\textit{óc}}{\Afootnote{+ þá  L679.}} Hr\textit{ómundr} \edtext{sv\textit{er}þ\textit{it}}{\Afootnote{÷  L679.}} Mistiltein, Men\textit{it} o\textit{c} Hríngin\textit{n}, o\textit{c} \edtext{m\textit{ar}ga}{\Afootnote{+ þar  L2404.}} aþra góþa gripi, gé\textit{ck} síþ\textit{an} t\textit{il}
fest\textit{ar}in\textit{n}\textit{ar} o\textit{c} las s\textit{ic} upp,
v\textit{or}o þá allir K\textit{ón}gs m\textit{enn} \edtext{frá}{\Afootnote{burt  L1572.}} h\textit{or}fn\textit{ir}; gé\textit{ck} Hróm\textit{undr} þ\textit{ar}
t\textit{il} \edtext{a\textit{t}}{\Afootnote{÷  L679, L2404, L1572.}} h\textit{ann} \edtext{hitti}{\Afootnote{finnr  L679.}} Olaf K\textit{óng} o\textit{c} m\textit{enn} h\textit{an}s.
K\textit{óngr}  fagnaþi   
\edtext{Hróm\textit{undi}}{\Afootnote{hónum L679.}}  v\textit{el},
o\textit{c} m\textit{ælir} hv\textit{or}nig skildot þiþ
Þráin\textit{n}? Hróm\textit{undr} m\textit{ælir} allt ge\textit{ck} at  ósk\textit{om}, m\textit{eþ} h\textit{an}s egin sv\textit{er}þi hjó
e\textit{c} \edtext{af hỏn\textit{om} hỏfuþ\textit{it}}{\Afootnote{hófúd af honum  L679.}}, 
o\textit{c} \edtext{brendi}{\Afootnote{+ hann  L679.}} síþ\textit{an} upp t\textit{il} ỏsko.
 Slíkt  \edtext{v\textit{ar}}{\Afootnote{+ hid  L679, er  L1572.}} mesta happa v\textit{er}k  \edtext{s\textit{egir}}{\Afootnote{÷  L679.}}
 || 
 \ledouternote{114v}
 \edtext{K\textit{óngr}}{\Afootnote{÷  L679.}} o\textit{c} erto 
 \edtext{frægr}{\Afootnote{frægdar L1572}}\footnote{In L1572 the reading is abbreviated with a superscript \enquote{r}, which is usually used for \enquote{ar}, hence the expansion \enquote{frægdar}.} \edtext{m\textit{aþr}}{\Afootnote{+ af þessu segir Kongr  L679.}}.  \edtext{Toko}{\Afootnote{taka  L1572.}} þá K\textit{ón}gs m\textit{enn} ofr gnægþ af gulli
o\textit{c} \edtext{sylfri}{\Afootnote{giersemum  L679.}}, 
o\textit{c} \edtext{héldo}{\Afootnote{+ svo  L679.}} \edtext{brott}{\Afootnote{búrt L679, L2404, L1572.}} frá Svíþjóþ  
\edtext{o\textit{c}}{\Afootnote{÷  L679.}} t\textit{il} Norvegs  
þ\textit{ar} s\textit{em} í Bjỏrg\textit{om} \edtext{h\textit{ie}t}{\Afootnote{heitir  L679.}}, 
þ\textit{ar} h\textit{afþi} K\textit{óngr} vetrseto m\textit{eþ} \edtext{hyrþ}{\Afootnote{÷  L2404, mönnum  L1572.}}  \edtext{sína}{\Afootnote{sinni  L679, ÷  L2404, sinum  L1572.}},  \edtext{þ\textit{ann} vetr}{\Afootnote{÷  L2404.}},  \edtext{hjá}{\Afootnote{÷  L2404, med  L1572.}} Burgeis no\textit{ck}rom,  
\edtext{h\textit{ann} v\textit{ar}}{\Afootnote{÷  L2404.}} \edtext{Gnúdi}{\Afootnote{+ var  L2404.}} kallaþr, 
h\textit{ann} gaf K\textit{óngi} m\textit{ar}ga  \edtext{góþa}{\Afootnote{fásiena  L679.}} gripi o\textit{c} sva m\textit{ỏnnom} h\textit{an}s; h\textit{ann} gaf
Hróm\textit{undi} ein\textit{n}  ra\textit{ck}a,  \edtext{s\textit{em} v\textit{ar} sva v\textit{el}}{\Afootnote{er vel var  L679.}} viti b\textit{or}in\textit{n}, s\textit{em} m\textit{aþr}, o\textit{c}
skjótr s\textit{em}  ỏr, o\textit{c} hit  mesta
g\textit{er}semi  \edtext{v\textit{ar}}{\Afootnote{÷  L1572.}} \edtext{h\textit{ann}}{\Afootnote{sa  L679, ÷  L1572.}},  \edtext{sá}{\Afootnote{hann  L679.}} v\textit{ar} \edtext{Hrókr kallaþr}{\Afootnote{kallaþr Hrókur  L2404.}}. Hróm\textit{undr} gaf Gnúþa  \edtext{aptr ágæt\textit{an}\textit{n}}{\Afootnote{÷  L679.}} hríng af betsta gulli, oc  v\textit{ar} talat  \edtext{a\textit{t}}{\Afootnote{÷  L1572.}} h\textit{ann} \edtext{v\textit{ær}i}{\Afootnote{+ einn  L1572.}} or haugi Þráins, o\textit{c}  vóg   v\textit{it} ij m\textit{er}cr  sylfors; 
þ\textit{ess}a gjỏf ỏfundaþi Vole hyrþm\textit{aþr} K\textit{ón}gs,  \edtext{h\textit{ann}}{\Afootnote{og  L679.}} kémor a\textit{t} máli v\textit{it} Bíld félaga
sin\textit{n}, o\textit{c} segist vilja drepa  hund
þ\textit{an}\textit{n} er Hróm\textit{undr} h\textit{afi} eignast,
o\textit{c} s\textit{egir} a\textit{t} K\textit{óngr}  vyrþi  han\textit{n} l\textit{an}gt yf\textit{ir} aþra m\textit{enn},  
 sỏk\textit{om}  þ\textit{ess} \edtext{a\textit{t}}{\Afootnote{÷  L679.}} h\textit{ann} hafi í haugin\textit{n} g\textit{en}g\textit{it}, 
 oc  abl\textit{at}  góþra gripa. 
 Bíldr m\textit{ælir} \edtext{þetta}{\Afootnote{þat  L679.}} er dagsan\textit{n}a, o\textit{c} h\textit{it} betsta ráþ  \edtext{s\textit{em} þú s\textit{egir}}{\Afootnote{÷  L679.}} o\textit{c} skúl\textit{om} vit í nótt  \edtext{t\textit{il} reyna}{\Afootnote{reina til þess  L679.}}. 
 \pend 
 \endnumbering  

\beginnumbering
\pstart  
\vspace{5mm} \subsection*{VIII kap\textit{ítuli}} 
 Ein\textit{n} dag v\textit{ar} þ\textit{at} u\textit{m} v\textit{e}torin\textit{n}, 
 a\textit{t} K\textit{óngr} \edtext{befalaþi}{\Afootnote{kallar L2404, sagdi  L1572.}} \edtext{m\textit{ỏnnom} sín\textit{om}, 
 	a\textit{t} f\textit{ar}a}{\Afootnote{hyrd sína  L2404.}} \edtext{m\textit{eþ} sér á jagt}{\Afootnote{á jakt med sier  L679, á jagt  L2404.}},  \edtext{a\textit{t}}{\Afootnote{og  L679.}} \edtext{f\textit{an}ga}{\Afootnote{veida  L1572.}} \edtext{veiþ\textit{ir} fugla o\textit{c} dýra}{\Afootnote{dijr og fugla  L679, L1572.}}; þ\textit{ei}r g\textit{jỏr}a  s\textit{em} K\textit{óngr} \edtext{vildi}{\Afootnote{baud  L1572.}},  \edtext{ríþa}{\Afootnote{+ á  L679, L2404, ridu á  L1572.}} \edtext{m\textit{er}kr}{\Afootnote{+ og skoga  L679.}}, o\textit{c} \edtext{skjóta}{\Afootnote{skutu  L1572.}} fugla,  \edtext{e\textit{n}}{\Afootnote{og  L679.}} \edtext{jaga}{\Afootnote{jøgudu  L1572.}} dýr; 
 þ\textit{ei}m  sóktist ílla dýra jagtin\textit{n}, þ\textit{ví} dýrin\textit{n} v\textit{or}o  allstygg.  
 Hróm\textit{undr} sigaþi  \edtext{ra\textit{ck}a}{\Afootnote{seppa  L679.}} sín\textit{om} ept\textit{ir} dýrun\textit{om},  \edtext{e\textit{n}}{\Afootnote{og  L679.}} \edtext{h\textit{ann} dró}{\Afootnote{dro hann þaug óll  L679.}} upp  \edtext{hv\textit{or}t eitt}{\Afootnote{÷  L679.}}, 
 o\textit{c} u\textit{m} síþ\textit{ir} 
 fengo  \edtext{m\textit{enn}}{\Afootnote{þeir L1572.}}
 sva  \edtext{m\textit{ỏr}g}{\Afootnote{mikid  L679.}}  t\textit{il}  \edtext{bráþs}{\Afootnote{brada L679, L1572.}} 
 s\textit{em} þá gyrnti, o\textit{c} héldo  \edtext{nú}{\Afootnote{so  L1572.}} h\textit{ei}m m\textit{eþ} st\textit{or}ri  \edtext{skémtun}{\Afootnote{skiemtan  L679, L2404, L1572.}} \edtext{o\textit{c}}{\Afootnote{÷  L2404.}} \edtext{fỏgnuþi}{\Afootnote{gledi  L679, ÷  L2404.}}, o\textit{c} sỏgþo frá þ\textit{ví}, a\textit{t} Hrókr hefþi betst  \edtext{dregit}{\Afootnote{haft  L679.}} upp  dýrin\textit{n}, o\textit{c} eins hefþi h\textit{ann} fugl\textit{an}a sigraþ, lét þá K\textit{ón}gr  \edtext{und\textit{ir}}{\Afootnote{÷  L1572.}} \edtext{b\textit{or}þ}{\Afootnote{drickju bord  L679, ÷  L1572.}} setjast  \edtext{t\textit{il} matv\textit{er}þa o\textit{c}	dry\textit{ck}jo}{\Afootnote{÷  L679.}}, o\textit{c} v\textit{or}o allkát\textit{ir} e\textit{n} þeg\textit{ar} hætt v\textit{ar} \edtext{dry\textit{ck}jo}{\Afootnote{÷  L679.}}, g\textit{en}gjo m\textit{enn} t\textit{il} hvílo, o\textit{c}
t\textit{óc}o náþ\textit{ir}. 
Þ\textit{ei}r  \edtext{Vole o\textit{c} Bíldr}{\Afootnote{Billdr og Voli  L679.}}, hỏfþo gát á  \edtext{hundin\textit{om}}{\Afootnote{÷  L2404.}} Hrók, o\textit{c} þá 
||
\ledouternote{115r} 
m\textit{enn} v\textit{or}o sof\textit{an}di, g\textit{en}go
þ\textit{ei}r  \edtext{a\textit{t}}{\Afootnote{÷  L2404.}} hỏn\textit{om} o\textit{c} \edtext{stúngo}{\Afootnote{+ hann  L679, L1572.}} t\textit{il} b\textit{an}a, þ\textit{ví} h\textit{ann} \edtext{lá}{\Afootnote{var  L1572.}} sof\textit{an}di o\textit{c} \edtext{v\textit{ar}}{\Afootnote{÷  L1572.}} þreyttr; h\textit{ann} spratt upp v\textit{it} lag\textit{it} o\textit{c}
náþi til  Bílds, o\textit{c} beit í gégn\textit{om} h\textit{an}dlegg h\textit{an}s, e\textit{n} Vole hjó af Hróci   trín\textit{it}.
 \edtext{V\textit{it} þetta}{\Afootnote{i þvj  L679.}} vaknaþi Hróm\textit{undr} o\textit{c} sp\textit{yr} hv\textit{aþ}
u\textit{m} v\textit{ær}i; h\textit{an}n  setst upp o\textit{c}
sér hundin\textit{n} dauþ\textit{an}, e\textit{n} þ\textit{ei}r  \edtext{Vole o\textit{c} Bíldr}{\Afootnote{Bildr og Voli  L679.}} lupo  \edtext{á gátt}{\Afootnote{i burt  L1572.}}, o\textit{c} misti h\textit{ann} þ\textit{eir}ra í þ\textit{at}
sin\textit{n}. Hróm\textit{undi} líkaþi ílla  a\textit{t}
h\textit{ann} misti ra\textit{ck}an\textit{n}, e\textit{n} lét þó eigi á sér
heyra. Ann\textit{an}\textit{n} dag  \edtext{ept\textit{ir}}{\Afootnote{÷  L679.}} héldt Ol\textit{afr} K\textit{óngr} á skóg  \edtext{a\textit{t} veiþa}{\Afootnote{og  L679.}}, 
 \edtext{hỏn\textit{om} tócst}{\Afootnote{tokst hónum  L679.}} ílla  \edtext{a\textit{t} veiþa}{\Afootnote{veidinn  L679.}}; þ\textit{an}\textit{n} dag v\textit{ar} Hróm\textit{undr} heima
o\textit{c} f\textit{or} hv\textit{or}gi  K\textit{óngr}
sp\textit{yr} hv\textit{ar} Hróm\textit{undr} v\textit{ær}i? m\textit{enn}
qváþo h\textit{ann} sjúk\textit{an} ept\textit{ir} heima  \edtext{í m\textit{or}gun}{\Afootnote{÷  L1572.}}.  Þetta heyrdi Vole, o\textit{c} brosti. Ekki d\textit{au}þsjúkr s\textit{egir} h\textit{ann}, h\textit{ann} man  \edtext{ept\textit{ir} venjo}{\Afootnote{÷  L679.}} v\textit{er}a  \edtext{á tali}{\Afootnote{ad tala  L679.}} \edtext{hjá}{\Afootnote{vid  L679, L2404, L1572.}} Svanhvít syst\textit{ir} K\textit{ón}gs, þ\textit{ví} \edtext{hún elsk\textit{ar} h\textit{ann}}{\Afootnote{hann elskar hana  L679.}} meirst allra m\textit{anna}. Þ\textit{at} hefi e\textit{c} hugsat s\textit{egir} K\textit{óngr} a\textit{t}
tign\textit{ir} m\textit{enn} \edtext{skyldo}{\Afootnote{+ eiga  L679.}} Svanhv\textit{ít} oc  Dagnýo  \edtext{eignast}{\Afootnote{÷  L679.}}, e\textit{n} e\textit{ck}i kotk\textit{ar}la syn\textit{ir}. 
Allt sannast \edtext{síþ\textit{ar} u\textit{m} þ\textit{et}ta}{\Afootnote{þetta um sijdir  L679.}} s\textit{egir} V\textit{ole} o\textit{c} lauk sva tali,  \edtext{a\textit{t}}{\Afootnote{÷  L1572.}} K\textit{óngr} héldt heim  u\textit{m} qvỏld\textit{it},
o\textit{c} g\textit{en}go þá all\textit{ir} t\textit{il} dry\textit{ck}jo.
K\textit{óngr} sp\textit{yr} er Hróm\textit{undr} \edtext{sjúkr}{\Afootnote{veikr  L679.}} e\textit{n} nú? Ei er þ\textit{at} s\textit{egir} Hróm\textit{undr} heldr
bý e\textit{c} a\textit{t} kumlu\textit{m} Þráin\textit{n}s. Jlla
tókst o\textit{ss} veiþi í dag s\textit{egir} K\textit{óngr} þ\textit{ví} bæþi
v\textit{an}taþi þ\textit{ic} o\textit{c} \edtext{seppa}{\Afootnote{racka  L1572.}} þin\textit{n}. 
V\textit{er}a  \edtext{má}{\Afootnote{+ ad  L679.}} \edtext{sva}{\Afootnote{þad  L1572.}} \edtext{sé}{\Afootnote{hafi verid  L679, ÷  L1572.}} a\textit{t} h\textit{ann} hafi v\textit{an}taþ s\textit{egir}
Hróm\textit{undr} o\textit{c} þa\textit{ck}ir m\textit{eþ}
t\textit{il}hlýþil\textit{e}g\textit{om} laun\textit{om}, vildi ec  þ\textit{ei}m  \edtext{kun\textit{n}a}{\Afootnote{kunnad hafa  L679.}} s\textit{em} \edtext{h\textit{ann} drápo}{\Afootnote{drapu hann  L679, L2404.}} í fyrri nótt, o\textit{c} léto h\textit{ann} saclausan\textit{n} gjalda mín. Øll von er þ\textit{ess}, s\textit{egir}
K\textit{óngr} a\textit{t} slíkt vilj\textit{ir} þú, 
þ\textit{ví} h\textit{ann} v\textit{ar} afbragþ an\textit{n}\textit{ar}a
ra\textit{ck}a; liþo sva tím\textit{ar}, a\textit{t} ekk\textit{er}t
b\textit{ar} t\textit{il} tíþinda. 
Nú víkr sỏgun\textit{n}i t\textit{il} Greips F\textit{ỏþr}\footnote{Expansion follows the single written-out example of \enquote{fỏþr} on f. 117v.} þ\textit{ei}rra  bræþra o\textit{c}
Hróm\textit{undar} a\textit{t} h\textit{ann} \edtext{g\textit{jỏr}ir}{\Afootnote{giórdi  L679.}} \edtext{orþ Hróm\textit{undi}}{\Afootnote{hónum ord  L679.}} 
o\textit{c} biþr  \edtext{han\textit{n}}{\Afootnote{honum  L679.}} \edtext{þiggja heim boþ sitt, 
	o\textit{c} q\textit{vo}ma á sin\textit{n} fund}{\Afootnote{a sinn fund ad þiggja heim bod sitt  L679.}}. 
Þetta boþ 
||
\ledouternote{115v}
 þáþi Hróm\textit{undr} \edtext{a\textit{t} q\textit{vo}ma t\textit{il} fund\textit{ar}	 
 	v\textit{it} h\textit{ann}}{\Afootnote{og ferdadist af stad  L679, oc fer til fundar vid hann  L1572.}}. 
Greipr  \edtext{fagn\textit{ar}}{\Afootnote{fagnadi  L1572.}} hỏn\textit{om} v\textit{el}, o\textit{c} g\textit{jỏr}ir vegl\textit{e}ga v\textit{ei}tslo móti hỏn\textit{om}, o\textit{c}
sp\textit{yr} m\textit{ar}gra  luta frá Ol\textit{afi}
K\textit{óngi} Hróm\textit{undr} s\textit{agdi} slíkt er villdi, o\textit{c}
qv\textit{aþ} K\textit{óngr} g\textit{jỏr}a v\textit{el} t\textit{il} sín o\textit{c}
b\textit{ræþ}ra sin\textit{n}a. En\textit{n} hv\textit{or}nig géþjast þér at
Bíld o\textit{c} Vola s\textit{agdi} Gr\textit{eipr} allt læt
e\textit{c} þ\textit{at} ósagt s\textit{agdi} Hróm\textit{undr} o\textit{c}
man  þ\textit{at} sein\textit{n}a heyrast. Draum min\textit{n} hefi
e\textit{c} a\textit{t} s\textit{egja} þér Fr\textit{ændi}\footnote{The expansion based on the reading in L697.}  s\textit{egir} Gr\textit{eipr} m\textit{ic} dreymdi Ol\textit{af} K\textit{óng}
o\textit{c} þ\textit{ic}, o\textit{c} þókti mér  \edtext{a\textit{t}}{\Afootnote{÷  L2404, L1572.}} þú hafa  géf\textit{it} hỏn\textit{om} gullhrínga  \edtext{á báþ\textit{ar} hendurn\textit{ar}}{\Afootnote{tvö  L679.}}, o\textit{c} þókti mér  \edtext{a\textit{t}}{\Afootnote{÷  L679, L2404.}} hríng\textit{ar}n\textit{ir} \edtext{v\textit{ær}o}{\Afootnote{vera  L679, L2404.}} af  \edtext{eign}{\Afootnote{ætt  L679.}} \edtext{Hróks}{\Afootnote{+ hins  L679, L2404.}} sv\textit{ar}ta Mágs m\textit{ins}\footnote{The expansion based on the readings in L679 and L1572.}  \edtext{o\textit{c} þóktist e\textit{c}}{\Afootnote{eg þottist  L679, oc þóttiz  L2404.}} \edtext{sjá}{\Afootnote{÷  L2404.}} ept\textit{ir} hríngun\textit{om}, þ\textit{ví} \edtext{a\textit{t}}{\Afootnote{÷  L679, L2404.}} þ\textit{ei}r v\textit{or}o  \edtext{bæþi m\textit{ar}g\textit{ir}}{\Afootnote{bádir vænir  L679.}} o\textit{c} fagr\textit{ir};  
\edtext{þókti ⸌mér⸍}{\Afootnote{mier þotti  L679, L2404, L1572.}} \edtext{K\textit{óngr}}{\Afootnote{hann  L679.}} skeinkja  \edtext{þér}{\Afootnote{+ á  L2404.}} mót hríngun\textit{om}, sva st\textit{or}a yfirhỏfn, 
a\textit{t} hún huldi  þ\textit{ic} all\textit{an}\textit{n} \edtext{á}{\Afootnote{÷  L1572.}} milli   hvirfils o\textit{c}  ilja,  o\textit{c}
þ\textit{ar} á ept\textit{ir} sá e\textit{c} þ\textit{ic} í gỏmlo rúmi,
o\textit{c} \edtext{hafþ\textit{ir} þú}{\Afootnote{höfþo þeir  L2404.}} þá m\textit{ic}il svefn  læti; mér þókti Svanhv\textit{ít}
g\textit{an}ga a\textit{t} þér  \edtext{sof\textit{an}di}{\Afootnote{÷  L679.}}, o\textit{c} búa u\textit{m} \edtext{þ\textit{ic}}{\Afootnote{+ sofandi  L679.}},  e\textit{n} Dagný hjálpaþi hen\textit{n}i t\textit{il}
þ\textit{ess}a st\textit{ar}fa, e\textit{n} \edtext{þá}{\Afootnote{þegar  L679.}} þ\textit{ær} skildo v\textit{it} þ\textit{ic},  laut Svanhvít
a\textit{t} þér o\textit{c} kysti þ\textit{ic}, o\textit{c} t\textit{óc} hnoþagrasfag\textit{ur}t o\textit{c} batt þ\textit{ess} enda u\textit{m} hỏnd þér
o\textit{c} s\textit{agdi} þ\textit{ess}o skalto e\textit{ck}i sleppa, á
m\textit{ed}an  e\textit{c} sleppi e\textit{ck}i, síþ\textit{an}
rakti hún niþr hnoþ a\textit{t} ept\textit{ir} sér o\textit{c} gé\textit{ck} í
burt, e\textit{n} þú svafst ept\textit{ir} s\textit{em} áþr. Nú
hef e\textit{c} sagt þér draum  min\textit{n}, e\textit{n} þú mant
ráþa  \edtext{v\textit{er}þa m\textit{eþ}}{\Afootnote{÷  L679.}} \edtext{tíþin\textit{n}i}{\Afootnote{tijdindi  L679, ÷  L2404.}}. 
Hróm\textit{undr} \edtext{svar\textit{ar}}{\Afootnote{mælir  L1572.}}:  bæþi er þ\textit{at}, a\textit{t} \edtext{draumurin\textit{n} þyk\textit{ir} e\textit{ck}i sva}{\Afootnote{mier þiki draumurinn  L679, draumurinn þikir mer eckie so  L1572.}} m\textit{er}kil\textit{e}gr, enda erto  \edtext{o\textit{c} m\textit{aþr} orþin\textit{n}}{\Afootnote{ordinn madr  L679, og ordin madr  L1572.}} gamall. Satt er þ\textit{at} s\textit{egir} Gr\textit{eipr} \edtext{a\textit{t} e\textit{c} er gamall
	v\textit{or}þin\textit{n}}{\Afootnote{÷  L2404, ad eg er ordinn gamall  L1572.}},  e\textit{n} víst man draumurin\textit{n} eitthvaþ
m\textit{er}kja, þ\textit{ví} mér féllst  hugr u\textit{m}
hríng\textit{an}a, o\textit{c} þ\textit{at} villda e\textit{c},
Fr\textit{ændi}! a\textit{t} þú  \edtext{talaþ\textit{ir}}{\Afootnote{+ eij  L679.}} v\textit{it} þ\textit{ær} systr  Ol\textit{afs}
K\textit{ón}gs, o\textit{c} manto þ\textit{eir}ra njóta, 
e\textit{n} Bíld
o\textit{c} Vola skalto v\textit{ar}ast  þ\textit{ví}
þ\textit{ei}r e\textit{ro} ótrú\textit{ir} fleyrst\textit{om} o\textit{c}
s\textit{egja} K\textit{óngi} satt o\textit{c} log\textit{it} af ỏþr\textit{om} m\textit{ỏnnom}.  
Sat nú Hróm\textit{undr} \edtext{hjá F\textit{ỏþr} sín\textit{om} no\textit{ck}ra stund}{\Afootnote{um stund hjá fódr sijnum L679.}} o\textit{c} ept\textit{ir} þ\textit{at} býst 
||
\ledouternote{116r} 
h\textit{ann} 
\edtext{á burt}{\Afootnote{i burt L679, á brott L2404.}}, 
gé\textit{ck} þá Gr\textit{eipr} á veg m\textit{eþ} hỏn\textit{om},
o\textit{c} \edtext{b\textit{at}}{\Afootnote{+ hónum  L679.}} allra vyrta, at  heil\textit{ir} aptr findost;  \edtext{f\textit{er}þaþist}{\Afootnote{letti ei  L2404.}} Hróm\textit{undr} \edtext{t\textit{il} þ\textit{ess} a\textit{t} heim kom}{\Afootnote{þá heim  L679, fyrr enn hann kom heim  L2404, til þess hann heim kom  L1572.}},  o\textit{c} lét fátt yf\textit{ir} f\textit{er}þ
sin\textit{n}i 
\pend \endnumbering  

\beginnumbering 
\pstart  
\vspace{5mm}\subsection*{IX kap\textit{ítuli}} 
Ekki l\textit{an}gt ept\textit{ir} þetta, 
gé\textit{ck} Hróm\-\textit{undr} a\textit{t} fin\textit{n}a þ\textit{ær} \edtext{systr}{\Afootnote{+ þeir  L2404.}}; b\textit{ræþ}r  h\textit{an}s g\textit{en}go m\textit{eþ} hỏn\textit{om}, þ\textit{ær} t\textit{óc}o v\textit{el} \edtext{q\textit{vo}mo}{\Afootnote{kvedju  L679.}} þ\textit{eir}ra, 
o\textit{c} sp\textit{urþo} tíþinda, e\textit{n} þ\textit{ei}r  \edtext{sỏgþost}{\Afootnote{qvadu  L2404.}} engin vita. 
Hró\-m\textit{undr} m\textit{ælir} þ\textit{ess} eyrindis q\textit{vo}m e\textit{c} híngat, 
a\textit{t} e\textit{c} villdi ega gott v\textit{it} yþr\footnote{L679 has a reading \enquote{þig} corrected to \enquote{ydr}.}  \edtext{a\textit{t}}{\Afootnote{og  L1572.}} \edtext{skoþa}{\Afootnote{rada  L679.}} draum hylli F\textit{ỏþr} míns, þ\textit{ví} e\textit{c} þenki yþr bæþi   lítilát\textit{ar}  o\textit{c} \edtext{spak\textit{ar}}{\Afootnote{spakastar  L1572.}}.  \edtext{Hún s\textit{egir}}{\Afootnote{hún svarar  L2404, Þá segir Svanhvit  L1572.}} 
Jlla f\textit{ær}i mér a\textit{t} synja góþ\textit{om} m\textit{ỏnnom} \edtext{þ\textit{ess}}{\Afootnote{slijks  L679.}} er e\textit{c} \edtext{v\textit{ei}ta má}{\Afootnote{má veita  L2404.}}, oc  slíkt  sóm\textit{ir}  e\textit{ck}i, eþa hv\textit{or}s villto beiþast?
Hróm\textit{undr} s\textit{egir} hen\textit{n}i  draumin\textit{n} ept\textit{ir} þ\textit{ví} s\textit{em} v\textit{er}it hafþi. Hún  \edtext{m\textit{ælir}}{\Afootnote{segir  L679.}} Þ\textit{at} ræþ e\textit{c} þér,  a\textit{t} þú sért hjá
F\textit{ỏþr} þín\textit{om}, o\textit{c} \edtext{þ\textit{ei}r}{\Afootnote{÷  L1572.}} b\textit{ræþr} þín\textit{ir} m\textit{eþ} þ\textit{ér}, 
o\textit{c} skiptit yþr 
\edtext{e\textit{ck}ert}{\Afootnote{ecki L679.}} af br\textit{óþr} mín\textit{om},
þ\textit{ví} h\textit{ann} heldr ij falska m\textit{enn}, s\textit{e}m  stunda ept\textit{ir} a\textit{t}  rógb\textit{er}a   \edtext{þ\textit{ic}}{\Afootnote{ydúr  L679.}}, frá mér.  
Vænt er þ\textit{at} a\textit{t} v\textit{er}a  drengr, o\textit{c} vin\textit{n}a t\textit{il} s\textit{egir} Hróm\textit{undr} o\textit{c} skal e\textit{c} þ\textit{ví} heldr híngat  q\textit{vo}ma s\textit{em} þ\textit{ei}r m\textit{ei}ra af skjala, á m\textit{ed}an  \edtext{a\textit{t}}{\Afootnote{÷  L679.}} þér mér þ\textit{at} leyf\textit{it} \edtext{oc}{\Afootnote{enn  L679, ÷  L2404.}} \edtext{e\textit{ck}i ban\textit{n}\textit{it}}{\Afootnote{÷  L2404.}}. 
Ei man þ\textit{at} sva fljótt v\textit{er}þa s\textit{egir} hún, a\textit{t} e\textit{c} v\textit{ar}ni þér  \edtext{viþrmælis}{\Afootnote{vidmælis  L679.}}, o\textit{c} þ\textit{at} hefi e\textit{c} lengi þeinkt, a\textit{t} v\textit{it} m\textit{un}d\textit{om} \edtext{einhv\textit{or}n  tíma samtal ega}{\Afootnote{samtal eiga einkvórn tijma  L679.}}, 
o\textit{c} þ\textit{at} dreymdi m\textit{ic} eittsin\textit{n}, a\textit{t} e\textit{c} þóktist búa u\textit{m} þ\textit{ic} í eino rúmi, o\textit{c} v\textit{ar}sto lítt haldin\textit{n}, e\textit{n} þó mælt\textit{ir} þú t\textit{il} mín, e\textit{n} b\textit{ræþ}r\textit{om} þín\textit{om} bjó e\textit{c} an\textit{n}\textit{at} rúm, o\textit{c} tỏluþo þ\textit{ei}r e\textit{ck}i  \edtext{neitt}{\Afootnote{÷  L679.}} t\textit{il} mín.  
Máské\footnote{Note the use of \enquote{máské}, a Danish loanword from \textit{måske}, examples in OH dated to the 16th-20th cent.} sva v\textit{er}þi s\textit{egir} Hróm\textit{undr} o\textit{c} t\textit{óc} fíngorgull o\textit{c} gaf hen\textit{n}i, o\textit{c} gé\textit{ck} \edtext{burt síþ\textit{an}}{\Afootnote{sijdann heim  L679.}}, e\textit{n} hún b\textit{at} h\textit{ann} v\textit{el} f\textit{ar}a.
Litlo ept\textit{ir} þetta q\textit{vo}ma þ\textit{ei}r  \edtext{Vole o\textit{c} B\textit{íldr}}{\Afootnote{Bjldur og Voli  L679.}} \edtext{á tal}{\Afootnote{ad tali  L679.}} v\textit{it} K\textit{óng} o\textit{c} s\textit{egja} hỏn\textit{om} a\textit{t} Hróm\textit{undr} m\textit{eþ} b\textit{ræþ}r\textit{om} sín\textit{om} hafi g\textit{en}g\textit{it} t\textit{il} \edtext{þ\textit{eir}ra systra}{\Afootnote{tals vid þær sistr  L679.}}, eptir  v\textit{an}a sín\textit{om},  \edtext{s\textit{egja} þ\textit{ei}r}{\Afootnote{÷  L1572.}}. o\textit{c} má þ\textit{at} þykja und\textit{ar}l\textit{e}gt
o\textit{c} óhỏfþínglegt, a\textit{t} þ\textit{ar} sé  \edtext{e\textit{ck}ert}{\Afootnote{ecki  L679.}} a\textit{t} fund\textit{it},  \edtext{o\textit{c} af tala\textit{t}}{\Afootnote{÷  L2404.}}, a\textit{t} ein\textit{n} kotk\textit{ar}lsson hafi  gistíng hjá Kóngb\textit{or}n\textit{om} Jómfrúm  \edtext{u\textit{m}}{\Afootnote{og hafi þar gisting  L679.}} nætr o\textit{c} daga. K\textit{óngr} 
||
\ledouternote{116v}
m\textit{ælir} þetta hefi e\textit{c} ei fyrr vit\textit{at},
o\textit{c} skal sva e\textit{ck}i lengi t\textit{il} g\textit{an}ga  ef e\textit{c} má  \edtext{a\textit{t}}{\Afootnote{÷ L679, L2404.}} \edtext{no\textit{ck}ro}{\Afootnote{nockud  L679, ÷  L2404.}} \edtext{aktast}{\Afootnote{ráþa  L2404.}}, o\textit{c} lát\textit{it} m\textit{ic} vita n\textit{ær} þetta
t\textit{il} fellr  o\textit{c} \edtext{þ\textit{at}}{\Afootnote{þid  L679, L2404, L1572.}}  fregn\textit{it}. 
Þ\textit{ei}r játto þ\textit{ess}o, o\textit{c} skildo tal\textit{it};  \edtext{leituþost}{\Afootnote{leitast  L679.}} þ\textit{ei}r nú  \edtext{ept\textit{ir}}{\Afootnote{vid  L1572.}} á all\textit{ar} lund\textit{ir} a\textit{t} sitja u\textit{m}
Hróm\textit{und} n\textit{ær} h\textit{ann} \edtext{g\textit{an}gi}{\Afootnote{færi  L679.}} a\textit{t} fin\textit{n}a  Svanhv\textit{ít} \edtext{o\textit{c}}{\Afootnote{÷  L679.}} eittsin\textit{n} er Hróm\textit{undr} v\textit{ar} \edtext{á tali}{\Afootnote{ad tala  L679.}} v\textit{it} þ\textit{ær} \edtext{systr}{\Afootnote{+ þá  L679.}}, f\textit{or}o  þ\textit{ei}r  \edtext{á K\textit{ón}gs fund}{\Afootnote{nú til Kongs  L679.}} 
o\textit{c} s\textit{ỏgþu} han\textit{om}, a\textit{t} \edtext{nú v\textit{ær}i Hróm\textit{undr}}{\Afootnote{Hromundr væri  L679.}} q\textit{vo}min\textit{n} t\textit{il} \edtext{þ\textit{eir}ra systra}{\Afootnote{tals vid Svanhvit  L679, þeirra  L2404.}}. 
K\textit{óngr} bregþr  \edtext{fljótt}{\Afootnote{skiótt  L1572.}} v\textit{it} o\textit{c} \edtext{kall\textit{ar}}{\Afootnote{+ þá  L679.}} m\textit{enn} o\textit{c} biþr þá  t\textit{ac}a vapn sín,
o\textit{c} fylgja sér ept\textit{ir}; sn\textit{er}o þ\textit{ei}r þá
all\textit{ir} a\textit{t} h\textit{er}b\textit{er}gj\textit{om}
Svanh\textit{vítar} v\textit{ar} þá Dagný a\textit{t}  lauga  líntrefla þ\textit{eir}ra systra, e\textit{n} Svanh\textit{vít} v\textit{ar} a\textit{t} \edtext{máli}{\Afootnote{tala  L679, mæla  L1572.}} v\textit{it} Hróm\textit{und} Þá t\textit{óc} t\textit{il} orþa  þ\textit{er}na  Dagný\textit{ar}, o\textit{c} \edtext{s\textit{agdi}}{\Afootnote{mælti  L1572.}} lít þú upp Jómfr\textit{ú} o\textit{c} \edtext{bú þ\textit{ic}}{\Afootnote{búst þú  L679, L1572, búst  L2404.}} v\textit{it} br\textit{óþir} þín\textit{om} oc  hirþ
h\textit{an}s. Þá sá Dagný m\textit{ic}in\textit{n} fjỏlda a\textit{t}
drýfa; hún  ran\textit{n} á skunda  \edtext{t\textit{il}}{\Afootnote{og fann  L679.}} systr  \edtext{sin\textit{n}\textit{ar}}{\Afootnote{sijna  L679.}}, o\textit{c} \edtext{s\textit{egir}}{\Afootnote{÷  L2404.}} hen\textit{n}i tíþindin a\textit{t} Ol\textit{afr} K\textit{óngr} Br\textit{óþir} hen\textit{n}\textit{ar}
q\textit{vo}mi. Svanhv\textit{ít} gé\textit{ck} a\textit{t}
Hróm\textit{undi} o\textit{c} m\textit{ælir} nú mátto
e\textit{ck}i  tefja, þ\textit{ví} Br\textit{óþir} m\textit{inn} kémr in\textit{n}\textit{an} \edtext{lítils tíma}{\Afootnote{stundar  L1572.}} o\textit{c} vill  f\textit{an}ga þ\textit{ic}, skalto
t\textit{ac}a v\textit{it} \edtext{qven\textit{n}-skrúþa}{\Afootnote{skrúda  L679.}} v\textit{or}um, o\textit{c} \edtext{mæta}{\Afootnote{mæt  L679, L2404, L1572.}} K\textit{óngi} v\textit{it} gjætt\textit{ir}, o\textit{c}
b\textit{er} sængr  \edtext{fỏt}{\Afootnote{vorar  L679.}} o\textit{c} hvílo \edtext{voþ\textit{ir}}{\Afootnote{fót  L679.}} í f\textit{an}gi þér, en\textit{n} frétti K\textit{óngr} a\textit{t} Hróm\textit{undi} skalto
s\textit{egja} \edtext{a\textit{t}}{\Afootnote{÷  L679, L1572.}} h\textit{ann} q\textit{vo}min\textit{n} \edtext{sé}{\Afootnote{÷  L679.}} á  \edtext{fund  v\textit{or}n}{\Afootnote{vorn fúnd  L679.}}, 
o\textit{c} \edtext{sé nú}{\Afootnote{÷  L2404.}} \edtext{a\textit{t} tala}{\Afootnote{á tali  L679, L1572, tali  L2404.}} v\textit{it} m\textit{ic},  
\edtext{manot}{\Afootnote{munú  L679, L2404, L1572.}} þ\textit{it} \edtext{þá}{\Afootnote{÷  L679.}} skílja, e\textit{n} þú  skalt  \edtext{þá}{\Afootnote{÷  L679.}} hitta F\textit{ỏþr} þin\textit{n}, o\textit{c} \edtext{t\textit{ac}}{\Afootnote{takto  L2404.}} b\textit{ræþ}r þína m\textit{eþ} þér; síþ\textit{an} t\textit{óc} Svanh\textit{vít} ljósgul\textit{an}\textit{n} stein, o\textit{c}
dró u\textit{m} andlit h\textit{an}s, fé\textit{ck} h\textit{ann} þá  yf\textit{ir}lit hen\textit{n}\textit{ar} þjónosto mey\textit{ar}; qvaþdi
Hróm\textit{undr} \edtext{nu}{\Afootnote{÷  L679, L1572.}} h\textit{an}a,  e\textit{n} hún  \edtext{kysti h\textit{ann} o\textit{c}}{\Afootnote{÷  L679.}} \edtext{b\textit{at}}{\Afootnote{+ hann  L1572.}} heil\textit{an} f\textit{ar}a o\textit{c} aptr q\textit{vo}ma;
f\textit{or} þá Hróm\textit{undr} í  \edtext{Mey\textit{ar}skrúþ\textit{an}}{\Afootnote{meyarskruþunom  L2404.}}, o\textit{c} stritaþist út m\textit{eþ} sængrfata bagg\textit{an} á  ỏxlin\textit{n}i, 
o\textit{c} q\textit{vo}m þá K\textit{óng}r á móti  \edtext{stúlko þ\textit{ess}ari}{\Afootnote{þessari stúlkú  L679.}}, 
o\textit{c} sp\textit{yr} tíþinda,  \edtext{o\textit{c}}{\Afootnote{+ spir  L679, ÷  L2404, eda  L1572.}} hv\textit{or}t  \edtext{a\textit{t}}{\Afootnote{÷  L679, L1572.}} Hróm\textit{undr} sé þ\textit{ar} \edtext{eþa}{\Afootnote{÷  L2404.}} ei? 
Hún s\textit{egir} a\textit{t} sva muni  v\textit{er}a
a\textit{t} þau  \edtext{mælist}{\Afootnote{talist  L679.}} v\textit{it}, Svanh\textit{vít} o\textit{c} h\textit{ann}. K\textit{óngr}
sp\textit{yr} hv\textit{ar} \edtext{e\textit{ro} þau}{\Afootnote{þaúg sieú  L679.}} nú? 
stúlk\textit{an} s\textit{egir} í
nátth\textit{er}b\textit{er}gi  \edtext{hen\textit{n}\textit{ar}}{\Afootnote{÷  L679.}} \edtext{v\textit{ar} h\textit{ann} nú}{\Afootnote{voru þaúg  L679, var hann  L1572.}} f\textit{yrir} stutto,  
o\textit{c} má e\textit{c}
e\textit{ck}i tefja  \edtext{v\textit{er}k}{\Afootnote{+ mytt  L679, L2404, L1572.}}, a\textit{t} \edtext{h\textit{er}ma}{\Afootnote{giora  L1572.}} \edtext{fremr}{\Afootnote{framar  L679, L2404, ÷  L1572.}}.  \edtext{Far þú s\textit{egir} K\textit{óngr}}{\Afootnote{kongr segir far þú  L1572.}} o\textit{c} 
||
\ledouternote{117r}
 vin\textit{n} þ\textit{at} s\textit{em} \edtext{þér}{\Afootnote{fijrir þig  L679.}} v\textit{ar} \edtext{skip\textit{at}}{\Afootnote{lagt  L679.}}; gé\textit{ck} \edtext{þá stúlk\textit{an}}{\Afootnote{hun þá  L1572.}} sin\textit{n} veg,  \edtext{þan\textit{n}}{\Afootnote{÷  L2404.}} s\textit{em} hen\textit{n}i  v\textit{ar} f\textit{yrir}
lagþr, e\textit{n} K\textit{óngr} reiþ t\textit{il} p\textit{or}tdyra  \edtext{Jómfrúna}{\Afootnote{jómfruarinnar  L2404.}}, o\textit{c} v\textit{or}o þá allar  dyr lokaþ\textit{ar}.
K\textit{óngr} s\textit{agþi} \edtext{m\textit{ỏnnom}}{\Afootnote{+ sinom  L2404.}} upp a\textit{t} brjóta, e\textit{n} Dagný  \edtext{lét}{\Afootnote{dró  L1572.}} frá  lokr, o\textit{c} qv\textit{aþ} Hróm\textit{undr}
þ\textit{ar} dval\textit{it} hafa hjá systr s\textit{inni} o\textit{c} hér
hygg  e\textit{c} h\textit{ann} \edtext{e\textit{n}}{\Afootnote{÷  L679.}} v\textit{er}a þó felast kun\textit{n}i, o\textit{c} \edtext{meg\textit{it}}{\Afootnote{megi  L679, L2404, L1572.}} þér v\textit{an}dl\textit{e}ga leita  \edtext{láta}{\Afootnote{÷  L679.}} ept\textit{ir} han\textit{om}. K\textit{óngr} g\textit{en}gr nú
u\textit{m} h\textit{er}b\textit{er}gin o\textit{c} f\textit{yrir}
Svanh\textit{vít} hún fagn\textit{ar} K\textit{óngi}
v\textit{el}, o\textit{c} bauþ t\textit{il} v\textit{ei}tslo, e\textit{n}
K\textit{óngr} neitaþi, o\textit{c} \edtext{b\textit{at} h\textit{an}a}{\Afootnote{baud henni ad  L679.}} láta  \edtext{Hróm\textit{und} í fram\textit{m}i}{\Afootnote{fram Rómundr  L2404, Hromund framm  L1572.}}. 
Svanh\textit{vít} m\textit{ælir} hv\textit{aþ} vill
K\textit{ón}gurin\textit{n} leita  ept\textit{ir} \edtext{Hróm\textit{undi}}{\Afootnote{hónum  L679.}} 
þ\textit{ar} \edtext{s\textit{em}}{\Afootnote{÷  L679.}} \edtext{a\textit{t}}{\Afootnote{÷  L679, L2404.}} h\textit{ann} er h\textit{an}s betsti vin, o\textit{c} hefi e\textit{c}
hanom  \edtext{hér þ\textit{ví}}{\Afootnote{÷  L679, því  L2404, L1572.}} v\textit{el} tek\textit{it}, a\textit{t} mér vyrþist h\textit{ann} \edtext{ebla}{\Afootnote{afla  L679.}} o\textit{c} styrkja yþar makt o\textit{c} r\textit{ic}i,
e\textit{n} nú man h\textit{ann} bráþl\textit{e}ga á braut \edtext{snúin\textit{n}}{\Afootnote{snúaz  L2404.}} frá yþr,  \edtext{a\textit{t}}{\Afootnote{er  L1572.}} þér misgrun\textit{it} h\textit{ann} u\textit{m} þá luti er h\textit{ann}
er ei valdor af, o\textit{c} v\textit{ær}i þér heiþr h\textit{ær}ri,
a\textit{t} láta heingja  \edtext{þá lygara}{\Afootnote{á gálga  L2404.}}, s\textit{em} h\textit{ann} hafa  \edtext{v\textit{it} yþr rógbor\textit{it}}{\Afootnote{rogborid vid ydr  L679.}}, e\textit{n} e\textit{ro} þó hin\textit{ar} mesto  raggeitr,
nema a\textit{t} fremja  \edtext{galdra}{\Afootnote{galdr  L1572.}} o\textit{c} fjỏlkíngi  s\textit{em} engin\textit{n} er dáþ  \edtext{né}{\Afootnote{eda  L1572.}} drengskapr í. K\textit{óngr} m\textit{ælir} f\textit{yrir} m\textit{ic}
er  \edtext{q\textit{vo}m\textit{it}}{\Afootnote{+ þat  L679.}}, s\textit{em} satt man, 
a\textit{t} Hróm\textit{undr} \edtext{fipli}{\Afootnote{múni fipla  L679.}} þ\textit{ic}, 
o\textit{c} þ\textit{ví} held e\textit{c} þ\textit{at}
líkl\textit{e}gast, a\textit{t} þú  \edtext{hald\textit{ir} sv\textit{ar}i}{\Afootnote{takir svörum  L1572.}} h\textit{an}s. Svanh\textit{vít} m\textit{ælir} mansto e\textit{ck}i  \edtext{þ\textit{at}}{\Afootnote{÷  L679.}}, a\textit{t} hv\textit{or}ki þú sjálfr,  \edtext{eþr}{\Afootnote{ne  L2404.}} no\textit{ck}or  \edtext{an\textit{n}\textit{ar}}{\Afootnote{÷  L679.}} þin\textit{n}a m\textit{ann}a þ\textit{or}þi  a\textit{t} \edtext{f\textit{ar}a}{\Afootnote{ganga  L679, L2404.}} í haug Þráins nema  \edtext{Hróm\textit{undr}}{\Afootnote{hann  L679.}} ein\textit{n}. 
K\textit{óngr} m\textit{ælir} þ\textit{ar}   \edtext{af mektast}{\Afootnote{ofmetnast  L679.}} h\textit{ann}, o\textit{c} gyrnist a\textit{t}  brúka   þ\textit{ic} ept\textit{ir} \edtext{lystisemi}{\Afootnote{list  L1572.}} \edtext{sin\textit{n}i}{\Afootnote{þinni  L679.}},  \edtext{a\textit{t} h\textit{ann}}{\Afootnote{enn  L1572.}} leyn\textit{ir} m\textit{ic} þ\textit{ess}o, o\textit{c} \edtext{hér f\textit{yrir}}{\Afootnote{fijrir þad  L679.}} skal e\textit{c} láta festa  h\textit{ann} á  gálga, 
o\textit{c} strax mátto mér
h\textit{ann} í hendr selja. Svanhv\textit{ít} \edtext{s\textit{egir}}{\Afootnote{svarar  L2404.}} 
lát\textit{it} \edtext{þér}{\Afootnote{÷  L679.}} t\textit{ac}a h\textit{ann} \edtext{sjálf\textit{ir}}{\Afootnote{sjálfr þvj  L679, siálfr  L1572.}}, 
e\textit{c} hefi  \edtext{e\textit{ck}i han\textit{om}}{\Afootnote{hónum ecki  L679.}} \edtext{a\textit{t} fram\textit{m}visa}{\Afootnote{framm ad vijsa  L679.}}, o\textit{c} e\textit{ro} yþr heimil hús v\textit{or} t\textit{il}
\edtext{r\textit{an}sỏkun\textit{ar}}{\Afootnote{rannsoknar L679, L2404}}. 
K\textit{óngr} \edtext{skip\textit{ar}}{\Afootnote{baud  L1572.}} m\textit{ỏnnom} sín\textit{om} a\textit{t} leita
v\textit{an}dl\textit{e}ga, o\textit{c} sva g\textit{jỏr}þo þ\textit{ei}r,
o\textit{c} fundo e\textit{ck}i  a\textit{t} heldr.
K\textit{óngr} skip\textit{ar} \edtext{nú}{\Afootnote{÷  L679.}} a\textit{t} 
brjóta upp  \edtext{st\textit{or}ar hyrþslr}{\Afootnote{storhirdslur  L679, L2404, L1572.}},  
o\textit{c} leita ept\textit{ir} Hróm\textit{undi}
e\textit{n} h\textit{ann} fanst  \edtext{e\textit{ck}i}{\Afootnote{+ ad helldr  L679.}}. 
Þá reiþdist  \edtext{Ol\textit{afr}}{\Afootnote{÷ L679.}} K\textit{óngr} 
o\textit{c} sló Svanhv\textit{ít} \edtext{sva}{\Afootnote{÷  L679.}} st\textit{or}t andlits blak, a\textit{t} blóþ\textit{it} féll 
||
\ledouternote{117v}
 \edtext{of}{\Afootnote{nidr um  L679, om  L2404, L1572.}} h\textit{an}a o\textit{c} m\textit{ælir} haf þetta f\textit{yrir}
Hróm\textit{und} þín  armadubba\footnote{The expression \enquote{arma dubba} is an offensive term for women; examples in OH from the 17th and 18th cent. In Blöndal's dictionary  \textit{dubba}  (Da. \textit{kvinde, laptaske}) is illustrated with an example  \enquote{sagði þær örmu dubbur [...] skyldu}, while elsewhere in the dictionary we can find a similar expression: \enquote{þegi þú, þín arma gardranorn}, which is translated as \enquote{ti stille, din forbandede Heks}.}
o\textit{c} skalto jafn\textit{an} hafa reiþi mína. Svanhv\textit{ít}
m\textit{ælir} allt \edtext{of}{\Afootnote{um  L679, L2404, L1572.}} þ\textit{at} o\textit{c} ei a\textit{t} síþr skal e\textit{c} \edtext{þá}{\Afootnote{÷  L679, þó  L1572.}} Hróm\textit{und} v\textit{it} v\textit{ar}a o\textit{c} betst
un\textit{n}a,  e\textit{n} þ\textit{ar} á móti t\textit{il}
leggja a\textit{t} vin\textit{ir} þín\textit{ir} \edtext{Vole o\textit{c} Bíldr}{\Afootnote{Bijldr og Voli  L679.}} v\textit{er}þi  hengþ\textit{ir}. K\textit{óngr}
m\textit{ælir} lít\textit{it} man e\textit{c}  akta orþ þín, 
o\textit{c} man e\textit{c} ráþa þ\textit{ví} e\textit{c} vil, 
e\textit{n} Hróm\textit{undi}
hefr þú víst und\textit{an} q\textit{vo}m\textit{it}, oc  skal þér
\edtext{þ\textit{at}}{\Afootnote{÷  L679.}} e\textit{ck}i duga, 
a\textit{t} g\textit{jỏr}a m\textit{ic}
f\textit{or}n spurþ\textit{an} a\textit{t} tali ykkar
Hróm\textit{undar} þ\textit{ar} t\textit{il} g\textit{jỏr}ir þú
f\textit{or}smán tign þin\textit{n}i, a\textit{t} leggja lag  \edtext{þitt}{\Afootnote{÷  L2404.}} v\textit{it} slétt\textit{an} \edtext{Búrason}{\Afootnote{Bóndasun  L2404.}}. Svanh\textit{vít} \edtext{m\textit{ælir}}{\Afootnote{segir  L1572.}} \edtext{þ\textit{ar}}{\Afootnote{þad  L679, L2404.}} er  raun s\textit{em} \edtext{hún}{\Afootnote{÷  L2404.}} hrỏkkr, o\textit{c} hỏndl\textit{it} \edtext{þér ei}{\Afootnote{þid ecki  L679.}} Hróm\textit{und} at heldr, 
e\textit{n} \edtext{þ\textit{ar}}{\Afootnote{þér  L1572.}} miss\textit{it} \edtext{þér}{\Afootnote{þid  L679, L2404, ÷  L1572.}} hins vaskasta m\textit{ann}s,  \edtext{er}{\Afootnote{ef  L679.}} þér  lát\textit{it} h\textit{ann} burt flæmast sva
óhỏfþíngl\textit{e}ga. Ol\textit{afr} K\textit{óngr} m\textit{ælir} þ\textit{at} v\textit{ei}t e\textit{c} f\textit{yrir} víst
a\textit{t} þú hefr Hróm\textit{und} und\textit{an}
q\textit{vo}m\textit{it}, 
o\textit{c} skal  \edtext{þér}{\Afootnote{+ ei  L2404, L1572.}} slíkt  \edtext{duga}{\Afootnote{+ ecki  L679.}} a\textit{t} hylla h\textit{ann} t\textit{il} skam\textit{m}a m\textit{eþ}
þér? Þar  er raun s\textit{em} á skérst, s\textit{egir}
Svanh\textit{vít} v\textit{er}þr þá K\textit{óngr} af\textit{ar} reiþr, oc
snýr  \edtext{út}{\Afootnote{búrt  L679.}} af h\textit{er}b\textit{er}gino, 
o\textit{c} biþr m\textit{enn} sína  \edtext{a\textit{t}}{\Afootnote{÷  L679, L2404.}} skygnast eptir  Hróm\textit{undi} e\textit{n} h\textit{ann}
f\textit{an}st hv\textit{or}gi, snýr þá K\textit{óngr} t\textit{il}
Bæ\textit{ar} heim aptr  o\textit{c}  átelr   þá Bíld o\textit{c} Vola, a\textit{t}
þ\textit{ei}r hafi s\textit{ic} á st\textit{at} log\textit{it}, e\textit{n}
þ\textit{ei}r  \edtext{s\textit{egja} s\textit{ic}}{\Afootnote{seigjast  L679.}} satt sagt hafa, 
o\textit{c} t\textit{il} \edtext{bevísíng\textit{ar} hafi þ\textit{ær}}{\Afootnote{sanns þar um hefi  L2404, bevisingar hafi  L1572.}} \edtext{Jómfrúr}{\Afootnote{jomfrúrnar  L679, L1572, meyarnar  L2404.}} m\textit{eþ} kéndt  \edtext{h\textit{it}}{\Afootnote{þad  L1572.}} sama, 
v\textit{ar}þ nú K\textit{óngr} þetta  \edtext{sva}{\Afootnote{+ búid  L679, L1572.}} a\textit{t} h\textit{afa} \edtext{o\textit{c} u\textit{m}b\textit{er}a}{\Afootnote{÷  L1572.}}.  
Nú er a\textit{t} s\textit{egja} frá stúlkun\textit{n}i,
s\textit{em} út b\textit{ar} hvílo  voþ\textit{ir}n\textit{ar},
 a\textit{t} hún  \edtext{skipti}{\Afootnote{skiptir  L679.}}  skjótt   klæþ\textit{om}, o\textit{c} \edtext{gé\textit{ck}}{\Afootnote{for  L679.}} a\textit{t} fin\textit{n}a þá  \edtext{s\textit{yni} Greips}{\Afootnote{Greips syni  L1572.}}, o\textit{c} b\textit{at} þá fylgja sér; e\textit{n} \edtext{s\textit{em}}{\Afootnote{er  L1572.}} þ\textit{ei}r q\textit{vo}mo á skógin\textit{n}, snýr þá Hróm\textit{undr} út  \edtext{alv\textit{or}un\textit{n}i}{\Afootnote{af götunni  L2404.}},\footnote{The reading found in L2404 is a better reading in the context of the story.}  o\textit{c} s\textit{egir} þ\textit{ei}m a\textit{t}
fin\textit{n}a Fỏþr sin\textit{n}, o\textit{c} þ\textit{at} \edtext{g\textit{jỏr}þo}{\Afootnote{gióra  L1572.}} þeir.  Gr\textit{eipr} fagn\textit{ar} þ\textit{ei}m
ỏll\textit{om}, o\textit{c} b\textit{at} þá þ\textit{ar} \edtext{a\textit{t}}{\Afootnote{÷  L679.}} dveljast m\textit{ed}an villdo;  e\textit{n} þ\textit{ess} á  \edtext{mill\textit{om}}{\Afootnote{milli  L679, L2404.}} 
s\textit{em} þ\textit{ei}r þ\textit{ar} dvỏldo,  
\edtext{f\textit{or}}{\Afootnote{gékk  L679.}} Hróm\textit{undr} opt a\textit{t} fin\textit{n}a  vinkonr
sínar,  \edtext{o\textit{c}}{\Afootnote{÷  L2404.}} t\textit{óc}o þ\textit{ær} \edtext{han\textit{om}}{\Afootnote{þeim  L679.}} jafn\textit{an} v\textit{el}, o\textit{c} léto ei á  sér
festa reiþi Ol\textit{afs} Kóngs. \pend \endnumbering  

\beginnumbering \pstart  
\vspace{5mm}\subsection*{X kap\textit{ítuli}}
||
\ledouternote{118r}
Nú\footnote{From this point onwards, B11109 and L1572 number the chapters incorrectly, using \enquote{xi} for \enquote{x} and so on. In L2404 (using Roman numerals) and L670 (using Arabic numerals) the sequence of chapters is correctly numbered. This variation is not registered for the remaining chapters.} víkr Sỏgun\textit{n}i austr til Svíþjóþ\textit{ar}, 
a\textit{t} þ\textit{ei}r K\textit{ón}gar  Hálfdán\textit{ar} \edtext{sam\textit{an} draga}{\Afootnote{draga saman  L679.}} ógrin\textit{n}i h\textit{er}s 
o\textit{c} senda xij sína 
besto  \edtext{sveina}{\Afootnote{menn  L2404.}} t\textit{il} Ol\textit{afs} K\textit{óngs} m\textit{eþ} þau eyrindi,
a\textit{t} þ\textit{ei}r  \edtext{bjóþa}{\Afootnote{bjódi  L679.}} h\textit{ono}m  t\textit{il} b\textit{ar}daga  \edtext{o\textit{c} mótstỏþo sér a\textit{t} mæta}{\Afootnote{÷  L679, og ad mæta  L1572.}} v\textit{it} Væn\textit{ir}s Vatn o\textit{c} skal  \edtext{sú orrusta}{\Afootnote{orustan  L2404.}} haldast á h\textit{ar}þa Jsi f\textit{yrir} \edtext{þá skuld}{\Afootnote{þad  L679.}} a\textit{t} Ol\textit{afr} K\textit{óngr} hefr  \edtext{áþr fyrr}{\Afootnote{÷  L679.}} g\textit{jỏr}t ágáng á r\textit{ic}i þ\textit{eir}ra  \edtext{þ\textit{ei}m}{\Afootnote{÷  L679.}} a\textit{t} saklauso, 
 \edtext{o\textit{c} rænt þ\textit{ar} h\textit{er}f\textit{an}gi}{\Afootnote{÷  L679.}}. 
Sendim\textit{enn} Kóng\textit{an}a q\textit{vo}mo ein\textit{n} dag t\textit{il} Ol\textit{afs} K\textit{óngs} o\textit{c} g\textit{en}go í hỏllina 
o\textit{c} qvỏþdo h\textit{ann},  \edtext{þá}{\Afootnote{þat  L2404, ÷  L1572.}} er h\textit{ann} sat yf\textit{ir} b\textit{or}þum.  
\edtext{K\textit{óngr}}{\Afootnote{hann  L679.}} t\textit{óc} qveþjo þ\textit{eir}ra o\textit{c} sp\textit{yr} hv\textit{or}j\textit{ir} v\textit{ær}o? 
 \edtext{e\textit{n}}{\Afootnote{÷  L679.}} 
 þ\textit{ei}r  \edtext{sỏgþost}{\Afootnote{segjast vera  L679.}} send\textit{ir} 
 frá  \edtext{fræg\textit{om} K\textit{on}g\textit{om} í Svíþjóþ}{\Afootnote{Svia kongum  L1572.}} þ\textit{ess} eyrindis 
 a\textit{t} \edtext{þ\textit{ei}r}{\Afootnote{÷  L679.}} 
 \edtext{boþa}{\Afootnote{bioda L1572.}}
  yþr t\textit{il} b\textit{ar}daga  \edtext{s\textit{egja} þ\textit{ei}r}{\Afootnote{÷  L679, L1572.}}, 
  o\textit{c} skal yþ\textit{ar} \edtext{o\textit{c} þ\textit{eir}ra}{\Afootnote{÷  L679, L1572.}} samfundr 
  \edtext{v\textit{er}þa}{\Afootnote{vera L1572.}} á  vatnino Væn\textit{ir} \edtext{þ\textit{ar}}{\Afootnote{÷  L679, L1572.}} 
  á ísn\textit{om}; e\textit{n} ef þú þor\textit{ir} e\textit{ck}i, þá s\textit{egja} þ\textit{ei}r yþr  n\textit{or}ska   steingeit  \edtext{s\textit{em} haf\textit{ir} h\textit{or}n}{\Afootnote{÷  L2404.}} \edtext{e\textit{n} eng\textit{an}  m\textit{er}g}{\Afootnote{÷  L1572.}}. 
 K\textit{óngr} reiþdist þ\textit{ess}o, e\textit{n} lét þó e\textit{ck}i a\textit{t} g\textit{jỏr}a, 
 a\textit{t} leita á\footnote{Dittography in L679, \enquote{leita á} repeated.}  sendim\textit{enn}ina. 
 Han\textit{n} m\textit{ælti} þ\textit{or}a man e\textit{c} a\textit{t} q\textit{vo}ma, 
 þó a\textit{t} orka sé e\textit{ck}i \edtext{ỏll}{\Afootnote{÷  L1572.}} í h\textit{or}n\textit{om}, héldo  \edtext{sva sendim\textit{enn} aptr}{\Afootnote{þeir so  L679.}} t\textit{il} baka; 
 e\textit{n} Ol\textit{afr} K\textit{óngr} lætr sam\textit{an} 
 blása  \edtext{síno liþi}{\Afootnote{lidi sijnu  L679, L2404, L1572.}}, v\textit{an}t\textit{ar} þá s\textit{onu} Gr\textit{eips} o\textit{c} \edtext{þyk\textit{ir}}{\Afootnote{þotti  L679.}} nú  \edtext{st\textit{or}t}{\Afootnote{÷  L2404.}} sk\textit{ar}þí orþ\textit{it}, 
 a\textit{t} Hróm\textit{undr} v\textit{ar} \edtext{á burt}{\Afootnote{i burtú  L679, í brott  L2404, i burt  L1572.}} m\textit{eþ} b\textit{ræþ}r\textit{om} sín\textit{om}; 
 dreif þá a\textit{t} Ol\textit{afr} K\textit{óngr} \edtext{liþ h\textit{an}s þ\textit{at}}{\Afootnote{þad lid er  L679, lid hans er  L1572.}} t\textit{il} v\textit{ar}, s\textit{egir} þá K\textit{óngr}
m\textit{ỏnnom} sín\textit{om}, a\textit{t} þ\textit{ess}i orrusta skuli  \edtext{um mids vetr\textit{ar} bil haldast}{\Afootnote{halldast um mis vetrar bil  L679, haldast  L1572.}} á Væn\textit{ir}s Jsi. 
Nú  \edtext{fréttir}{\Afootnote{fretti  L2404.}} Hróm\textit{undr} a\textit{t} Ol\textit{afi} K\textit{óngi} er stefnt
t\textit{il} b\textit{ar}daga, h\textit{ann} f\textit{er} eptir 
venjo sin\textit{n}i a\textit{t} fin\textit{n}a Svanh\textit{vít} hún
fagn\textit{ar} han\textit{om} v\textit{el}, 
o\textit{c} \edtext{býr}{\Afootnote{bio  L1572.}} t\textit{il} v\textit{ei}tslo heimugl\textit{e}ga, 
o\textit{c} biþr Hróm\textit{undi} þiggja;  
\edtext{h\textit{ann}}{\Afootnote{og  L679.}} segist  sva g\textit{jỏr}a m\textit{un}di;  
\edtext{h\textit{ann}}{\Afootnote{og  L679.}} sp\textit{yr} m\textit{ar}gra luta  \edtext{of}{\Afootnote{um  L679, L2404, L1572.}} Ol\textit{af} K\textit{óng} 
hún s\textit{egir} \edtext{a\textit{t}}{\Afootnote{÷  L679, L2404, L1572.}} h\textit{ann} sé 
\edtext{nú}{\Afootnote{÷ L679.}} í liþs sam\textit{an} drætti, 
o\textit{c} haldi sn\textit{ar}t  \edtext{burt}{\Afootnote{÷  L679.}} t\textit{il} b\textit{ar}daga  \edtext{á}{\Afootnote{÷  L679.}} móti  \edtext{þ\textit{ei}m}{\Afootnote{÷  L679.}} Sví\textit{om}, eþa, s\textit{egir} hún, ætl\textit{ar} þú a\textit{t}
fylgja Br\textit{óþr} mín\textit{om} í þ\textit{ess}a f\textit{er}d.  
\edtext{Ogj\textit{ỏr}la}{\Afootnote{ögjór  L679.}} v\textit{ei}t e\textit{c} þ\textit{at}, s\textit{egir} Hróm\textit{undr},
e\textit{n} óm\textit{er}kil\textit{e}gt  \edtext{er}{\Afootnote{+ þad  L679.}} a\textit{t} sitja a\textit{t} gj\textit{ỏr}þalaus f\textit{yrir} góþa m\textit{enn} n\textit{ær} gran\textit{n}||\ledouternote{118v}ar þ\textit{eir}ra gjæta st\textit{ar}fa sin\textit{n}a. 
Enga bón nen\textit{n}i e\textit{c} \edtext{a\textit{t} fram\textit{m}b\textit{er}a}{\Afootnote{fram at bera  L2404.}} f\textit{yrir} K\textit{óng} s\textit{egir} hún, a\textit{t} þú mátt
sjálfráþr v\textit{er}a. 
Vel er þ\textit{at} s\textit{egir} Hróm\textit{undr} \edtext{þó}{\Afootnote{÷  L2404, L1572.}} \edtext{a\textit{t}}{\Afootnote{÷  L679.}} 
þ\textit{ei}r Bíldr o\textit{c} Vole  \edtext{prófi}{\Afootnote{reýni  L2404, prófa  L1572.}} sína  \edtext{frægþ}{\Afootnote{list frægd  L679.}} 
o\textit{c} hreysti \edtext{núna einosin\textit{n}i}{\Afootnote{nú einú sinni L679, ÷  L2404, L1572.}} 
 o\textit{c} v\textit{ei}ti K\textit{óng}i hollusto í þ\textit{ví} \edtext{s\textit{em}}{\Afootnote{÷  L679.}} \edtext{h\textit{ỏnom}}{\Afootnote{÷  L2404, þeim L1572.}} meirst áliggr. 
Ekki hrygg\textit{ir} m\textit{ic} þ\textit{at} s\textit{egir} hún, þó þ\textit{ei}r  \edtext{mætto}{\Afootnote{megi  L1572.}} neyta h\textit{an}da sin\textit{n}a, en\textit{n} s\textit{or}ga e\textit{c} \edtext{meira, a\textit{t}}{\Afootnote{÷  L679, meira  L1572.}} ef þú,  \edtext{eþr}{\Afootnote{og  L679, L2404, L1572.}} b\textit{ræþ}r þín\textit{ir} bíþa kaldt af þ\textit{ess}\textit{om} b\textit{ar}daga, þ\textit{ví} mér rís hugr  \edtext{u\textit{m}}{\Afootnote{vit  L2404, L1572.}} \edtext{þ\textit{at}}{\Afootnote{því  L1572.}} laung\textit{om}, e\textit{n} þó ef  \edtext{a\textit{t}}{\Afootnote{÷  L679, L1572.}} Br\textit{óþr} m\textit{inn} v\textit{er}þr ofr liþa b\textit{or}in\textit{n}, þá kan\textit{n}
sæmd  \edtext{v\textit{or}ri a\textit{t} v\textit{er}þa}{\Afootnote{vor uerda  L679.}} a\textit{t} ósóma.  
Hróm\textit{undr} brosti  \edtext{a\textit{t}}{\Afootnote{÷  L679, L2404.}}, 
o\textit{c} gé\textit{ck} a\textit{t} dre\textit{ck}a,  \edtext{o\textit{c}}{\Afootnote{enn  L679, ÷  L1572.}} \edtext{ept\textit{ir} dry\textit{ck}jona}{\Afootnote{siþan  L2404, eptir þad  L1572.}} \edtext{reiþ}{\Afootnote{for  L679.}} 
h\textit{ann} heim  \edtext{aptr}{\Afootnote{÷  L679.}} t\textit{il} F\textit{ỏþr} síns.  
\pend 
\endnumbering 
\beginnumbering 
\pstart  
\vspace{5mm}
\subsection*{XI kap\textit{ítuli}} 
Ept\textit{ir} þetta, \edtext{þeg\textit{ar}}{\Afootnote{+ ad  L679, L2404, L1572.}} sendim\textit{enn} Svía K\textit{ón}ga q\textit{vo}mo heim aptr  o\textit{c} s\textit{ỏgþo} slíkt er skéþ hafþi,  
\edtext{héldo þ\textit{ei}r Hálfdán\textit{ar} K\textit{ón}g\textit{ar} af  st\textit{ad} m\textit{eþ} m\textit{ic}in\textit{n} fjỏlda skipa o\textit{c} m\textit{anna}}{\Afootnote{÷  L679.}}, þ\textit{ei}r snúa skip\textit{om} sín\textit{om} vestor
m\textit{eþ} l\textit{an}di, þ\textit{ar} t\textit{il} þ\textit{ei}r  \edtext{q\textit{vo}ma}{\Afootnote{komu  L679, L1572.}} a\textit{t} Væn\textit{ir}, þ\textit{ar} \edtext{g\textit{an}ga}{\Afootnote{geingu  L679, L1572.}} þ\textit{ei}r á l\textit{an}d, oc  setja
h\textit{er}búþ\textit{ir} sín\textit{ar} o\textit{c} búast v\textit{el}
u\textit{m}; þ\textit{ei}r hỏfþo einvala liþ,  o\textit{c}
h\textit{it}  h\textit{ar}dfengasta; 
m\textit{eþ}
þ\textit{ei}m v\textit{ar} Helgi  \edtext{hin\textit{n}}{\Afootnote{÷  L679.}} Frækni, h\textit{ann} hafþi hin betsto h\textit{er}klæþi, o\textit{c} sv\textit{er}þ
þ\textit{at} er Hraungv\textit{iþr} \edtext{b\textit{or}it  hafþi}{\Afootnote{hafdi att  L679.}}; 
Frillo átti Helgi sér  \edtext{s\textit{em}}{\Afootnote{er  L2404, L1572.}} Kára h\textit{ie}t, 
hún v\textit{ar} \edtext{fjỏlkun\textit{n}ug}{\Afootnote{+ miöc  L2404.}} o\textit{c} \edtext{hamf\textit{ar}ar}{\Afootnote{hamfara  L679.}} giþja, 
hún fló optast í  álptar  ham,  \edtext{o\textit{c}}{\Afootnote{÷  L679.}} stund\textit{om} dreka,  \edtext{o\textit{c}}{\Afootnote{÷  L679, L1572.}} stund\textit{om} v\textit{ar} hún  \edtext{qvendi}{\Afootnote{kona  L679, L1572.}} eþr  \edtext{k\textit{ar}l}{\Afootnote{karlmadur  L1572.}}, 
stund\textit{om} hj\textit{or}tr eþr hind, o\textit{c} v\textit{ar}
Helga  \edtext{h\textit{it}}{\Afootnote{÷  L1572.}} meirsta  traust a\textit{t} hen\textit{n}i, hv\textit{or}jo
s\textit{em} v\textit{ar} a\textit{t} sæta, fylgþi hún h\textit{ỏnom}
jafnan, þá  \edtext{h\textit{ann}}{\Afootnote{÷  L679.}} 
t\textit{il} \edtext{no\textit{ck}urra}{\Afootnote{einhvórra  L679.}} stórræþa f\textit{er}þa\-þist. 
Nú er at  \edtext{greina}{\Afootnote{segja  L1572.}} frá Ol\textit{afi} K\textit{óngi} 
a\textit{t} h\textit{ann} heldr  \edtext{m\textit{eþ} h\textit{er} sin\textit{n} af st\textit{ad}}{\Afootnote{af stad med her sinn  L679.}}, 
o\textit{c} hafþi  þá Bíld og Vola m\textit{eþ} sér, stóþ nú
allt  \edtext{h\textit{an}s traust}{\Afootnote{traust hans  L1572.}} á  þeim  \edtext{báþ\textit{om}}{\Afootnote{÷  L679.}}, láta þ\textit{ei}r m\textit{ic}it af sér, o\textit{c} heita
K\textit{on}gi sínu 
||
\ledouternote{119r}
 fulltíngi. Sva v\textit{ar} vetr\textit{ar}lag, a\textit{t}
e\textit{ck}i hafþi sn\textit{ær}drif\textit{it},  e\textit{n}
g\textit{en}g\textit{it} \edtext{l\textit{an}g}{\Afootnote{langt  L2404, langsamt  L1572.}} frost m\textit{eþ} bj\textit{ar}t  \edtext{vind\textit{om}}{\Afootnote{vidri  L679, vidrum  L1572.}}. 
K\textit{óngi} greiþdist  \edtext{v\textit{el}}{\Afootnote{+ ferdinn  L679.}} yf\textit{ir} \edtext{skóg}{\Afootnote{skoga  L679.}} o\textit{c} m\textit{er}kr, þ\textit{ar} t\textit{il} h\textit{ann}
q\textit{vo}m a\textit{t} Væn\textit{ir},  \edtext{v\textit{or}o þ\textit{ar}}{\Afootnote{var þar fijrir  L679, stódu  L1572.}} \edtext{h\textit{er}tjỏld}{\Afootnote{herlid  L679.}} \edtext{þ\textit{ei}rra}{\Afootnote{÷  L679.}} Svía  \edtext{þ\textit{ar} f\textit{yrir}}{\Afootnote{÷  L679, fyrir  L2404.}}, beiþ  \edtext{þá}{\Afootnote{þad  L1572.}} e\textit{ck}i lengi, þ\textit{ar} t\textit{il} \edtext{a\textit{t}}{\Afootnote{÷  L679, L1572.}} hỏfst m\textit{eþ} þ\textit{ei}m  b\textit{ar}dagi,
o\textit{c} \edtext{hin sn\textit{ar}pasta}{\Afootnote{og hórd  L679.}} orrusta. 
Sví\textit{ar} v\textit{or}o ákafafull\textit{ir}
o\textit{c} íll\textit{ir} viþr eign\textit{ar}; þ\textit{ei}r
g\textit{en}go í liþ Ol\textit{afs}
K\textit{óngs} oc  \edtext{britjuþo}{\Afootnote{+ þad L679.}} niþr  
\edtext{liþ h\textit{an}s}{\Afootnote{÷  L679.}}, s\textit{em} gras á jỏrþo. 
 Ol\textit{afr} K\textit{óngr} sókti  á móti þ\textit{ei}m Sví\textit{om} m\textit{eþ} meirsta drengskap;
h\textit{ann} sl\textit{ær} o\textit{c} leggr  \edtext{á báþ\textit{ar} hendr}{\Afootnote{til beggja handa  L679.}} o\textit{c} v\textit{ar} hin\textit{n} ỏruggasti. 
Helgi Frækni v\textit{ar}  Oþdviti   þ\textit{eir}ra Svía,
e\textit{n} Bíldr o\textit{c} V\textit{ole} þ\textit{eir}ra N\textit{or}þm\textit{anna}, sn\textit{er}i þá m\textit{ann}fallino á
Ol\textit{af} K\textit{óng} sva h\textit{ann} t\textit{óc} a\textit{t}  tvíla   u\textit{m} sigor sæld sína, þ\textit{ví} þá t\textit{óc} a\textit{t}
fæ\textit{ck}a  \edtext{liþ}{\Afootnote{herlid  L679.}} h\textit{an}s,  \edtext{sva hópum  gégndi}{\Afootnote{÷  L679.}}, lá  \edtext{þá}{\Afootnote{÷  L1572.}} \edtext{þ\textit{ei}m}{\Afootnote{÷  L679, L2404.}} \edtext{N\textit{or}þm\textit{ỏnnom}}{\Afootnote{+ þá  L1572.}} v\textit{it} a\textit{t} flýa. Ol\textit{afr} K\textit{óngr} 
b\textit{at} \edtext{m\textit{enn}}{\Afootnote{þá  L679.}} ei  uppgéfast, drógst sva und\textit{an} a\textit{t}
flóttan\textit{om} \edtext{yrþi}{\Afootnote{varþ ei  L1572.}} \edtext{fram\textit{m}qvæmt}{\Afootnote{vidkomid  L679.}}, leiþ sva af dagurin\textit{n} e\textit{n} nóttin t\textit{óc}
v\textit{it}, v\textit{ar} þá haldit  \edtext{uppi}{\Afootnote{upp  L1572.}} \edtext{friþ\textit{ar}skyldi}{\Afootnote{fridskildi  L1572.}}, lét þá K\textit{óngr} k\textit{ann}a liþ sitt, hafþi þá 
fall\textit{it} tæpr þriþjúngr af N\textit{or}þm\textit{ỏnnom}, sendi þá
Ol\textit{afr} K\textit{óngr} ept\textit{ir} liþi heim t\textit{il} sín. 
An\textit{n}\textit{an} \edtext{dag}{\Afootnote{+ eptir  L1572.}} t\textit{óc}st aptr hin sn\textit{ar}pasta  orrusta, féllo þá
m\textit{ar}g\textit{ir} af  \edtext{hv\textit{or}j\textit{om} f\textit{yrir}
	s\textit{ic}}{\Afootnote{hverotveggiom  L2404.}} sva fyrn\textit{om} gégndi.  
Helgi sókti betst
fram\textit{m}, o\textit{c} \edtext{stóþ}{\Afootnote{stódst  L679, L2404.}} engin\textit{n} \edtext{v\textit{it}}{\Afootnote{fyrir  L1572.}} h\textit{ỏnom}, h\textit{ann} drap bæþi  m\textit{enn}
o\textit{c} hesta, o\textit{c} \edtext{alla}{\Afootnote{hvórja  L679.}} fylkíng 
f\textit{or} \edtext{h\textit{ann}}{\Afootnote{hun  L2404.}} \edtext{í}{\Afootnote{÷  L679.}} gégn\textit{om}. Þetta sjer  Ol\textit{afr} K\textit{óngr}
o\textit{c} ríþr
h\textit{ar}þt fram\textit{m}, o\textit{c} biþr m\textit{enn} e\textit{ck}i
und\textit{an}leita. Þat  heyrþi Bíldr o\textit{c} \edtext{snýr í}{\Afootnote{sneri  L679.}} móti Helga, o\textit{c} hỏggr t\textit{il} h\textit{an}s af 
ỏllo megni. Helgi  \edtext{brá}{\Afootnote{b[or]  L679.}} f\textit{yrir} skyldin\textit{om},  \edtext{e\textit{n}}{\Afootnote{so  L679, oc  L2404.}} \edtext{sv\textit{er}þ\textit{it} festist}{\Afootnote{festiz sverþit  L2404.}} \edtext{í}{\Afootnote{+ honum  L2404.}};  þ\textit{at} sá Helgi, o\textit{c} \edtext{snar\textit{ar}}{\Afootnote{sneri  L1572.}} \edtext{sva h\textit{ar}þt skjỏldin\textit{n}}{\Afootnote{skildinum so hart  L679, sva hart skyldinom  L2404, so fast skiöldinn  L1572.}}, a\textit{t} \edtext{Bíld}{\Afootnote{Býldi  L2404.}} v\textit{ar}þ  laust sv\textit{er}þ\textit{it}; h\textit{ann}
blés þá vindi  \edtext{mót}{\Afootnote{til  L1572.}} Helga,  \edtext{sva}{\Afootnote{oc  L1572.}} \edtext{h\textit{ỏnom} réþi}{\Afootnote{redi honum  L1572.}} \edtext{t\textit{il} falls}{\Afootnote{vid  L679.}} o\textit{c} misti vopna,  \edtext{h\textit{ann} sjer hv\textit{ar} lá ein eykar  kylfa, h\textit{an}a þrífr Helgi}{\Afootnote{Helgi þrifur þá eina eikar kilfu  L679.}}, o\textit{c} \edtext{reir\textit{ir}}{\Afootnote{keirir  L679, L2404, L1572.}} \edtext{h\textit{an}a}{\Afootnote{÷  L679.}} í hỏf\textit{ut} á  \edtext{Bíld}{\Afootnote{Býldi  L2404.}}, 
||
\ledouternote{119v}
sva hausin\textit{n} molaþist í p\textit{ar}ta; Helgi lamdi  \edtext{h\textit{ann}}{\Afootnote{+ frá  L679.}} allan\textit{n} í sundr, e\textit{n} á m\textit{ed}an þetta
skéþi, áttust þ\textit{ei}r v\textit{it} Ol\textit{afr} K\textit{óngr} o\textit{c} an\textit{n}\textit{ar} \edtext{K\textit{óngr} þ\textit{eir}ra Svía}{\Afootnote{Svija Kongr  L679.}}; 
Ol\textit{afr} K\textit{óngr} \edtext{b\textit{ar} eitt h\textit{it} betsta}{\Afootnote{hefþi ágætt  L2404.}} sv\textit{er}þ o\textit{c} \edtext{herklæþi}{\Afootnote{+ in besto  L2404.}}, h\textit{ann} hỏggr  \edtext{þ\textit{ví}}{\Afootnote{÷  L679.}} t\textit{il} \edtext{Hálfdáns}{\Afootnote{Halfdanar  L679, + kongs  L2404.}}, e\textit{n} h\textit{ann} brá v\textit{it} \edtext{skyldi}{\Afootnote{skilldinum  L679.}}, oc   klofn\textit{aþ}i   h\textit{ann} í
tven\textit{n}t. Þá hjó Hálfdán t\textit{il} Ol\textit{afs} K\textit{óngs}
o\textit{c} stemndi á hjálmin\textit{n}. 
Þat sá Ol\textit{afr} K\textit{óngr} o\textit{c} sn\textit{er}i v\textit{it} hestin\textit{om}, oc
q\textit{vo}m hỏgg\textit{it} á hestin\textit{n} f\textit{yrir} fram\textit{an} bog\textit{an}a o\textit{c} þ\textit{ar} í sundr u\textit{m} þv\textit{er}t. 
Olafr K\textit{óngr}  stỏ\textit{ck} af baki fyrr  \edtext{e\textit{n}}{\Afootnote{+ hann  L679, L1572.}} gat fall\textit{it}, o\textit{c} q\textit{vo}m 
st\textit{an}d\textit{an}di niþr, o\textit{c} hjó á \edtext{ỏxlina}{\Afootnote{øxl  L1572.}} \edtext{á Hálfdáni K\textit{óngi}}{\Afootnote{Halfdanar kongs  L1572.}} oc  sneiþ  \edtext{of\textit{an}}{\Afootnote{af  L1572.}} \edtext{h\textit{er}þ\textit{ar}}{\Afootnote{herda  L679.}} blaþ\textit{it} o\textit{c} \edtext{síþuna}{\Afootnote{+ og  L679, L2404.}}, fylgdi þ\textit{ar} m\textit{eþ} hỏndin,  
féll þá Hálfdán  \edtext{K\textit{óngr}}{\Afootnote{+ og   L2404.}} v\textit{ar} \edtext{þá}{\Afootnote{÷  L2404.}} Ol\textit{afr} K\textit{óngr} \edtext{m\textit{ic}it}{\Afootnote{mióg  L679.}} móþr, e\textit{n} \edtext{lítt}{\Afootnote{ecki  L679.}} sár, sjer h\textit{ann} þá fall m\textit{anna} sin\textit{n}a, er
Helgi o\textit{c} Hálfdán  g\textit{jỏr}þo, e\textit{n} sin\textit{n} \edtext{átrún\textit{at}}{\Afootnote{átrunadr  L679.}} Vola, sá h\textit{ann} hv\textit{or}gi, þeinkti K\textit{óngr} \edtext{a\textit{t}}{\Afootnote{÷  L2404.}} han\textit{n} m\textit{un}di fallin\textit{n} v\textit{er}a,  \edtext{t\textit{óc} þá}{\Afootnote{þá tók  L1572.}} Ol\textit{afi} K\textit{óngi} a\textit{t} sýnast óvís sigorin\textit{n},
þ\textit{ar} s\textit{em} \edtext{a\textit{t} báþ\textit{ir}}{\Afootnote{÷  L679.}} 
\edtext{v\textit{ær}o falln\textit{ir}}{\Afootnote{fallnir vorú  L679.}} Bíldr o\textit{c} Vole.  
\pend \endnumbering  

\beginnumbering \pstart  
\vspace{5mm}\subsection*{XII kap\textit{ítuli}} 
Nú er a\textit{t} s\textit{egja} frá sendim\textit{ann}i
Ol\textit{afs} K\textit{óngs} a\textit{t} h\textit{ann} kémr heim,  o\textit{c} \edtext{heimt\textit{ar}}{\Afootnote{heimtir  L679, L1572.}} liþ sam\textit{an}, o\textit{c} s\textit{egir} Ol\textit{afr}
K\textit{óngr} n\textit{ær} þ\textit{ví} yf\textit{ir}
un\textit{n}in\textit{n};  \edtext{h\textit{onom} v\textit{er}þr}{\Afootnote{og verþr honum  L2404.}} lít\textit{it} t\textit{il} liþs, o\textit{c} f\textit{ær} \edtext{ccc}{\Afootnote{iijc  L2404.}} m\textit{ana}; þetta  \edtext{fregn\textit{ar}}{\Afootnote{friettir  L679, L1572.}} Svanh\textit{vít} \edtext{hún}{\Afootnote{og  L679.}} lætr t\textit{ac}a  \edtext{jv}{\Afootnote{vj  L679.}} far\textit{ar}skjóta, o\textit{c} sỏþla sér o\textit{c} \edtext{f\textit{er}þast}{\Afootnote{fer  L679.}} h\textit{it} sn\textit{ar}asta á fund Hróm\textit{undar}
þ\textit{ei}r feþg\textit{ar} fagna hen\textit{n}i m\textit{eþ} meirsto blíþo, o\textit{c} qváþo stór\textit{ar} nýúng\textit{ar}
v\textit{it} b\textit{er}a, a\textit{t} \edtext{slík\textit{ir} géstir vitjuþo}{\Afootnote{slikr gestr vitjadi  L1572.}} \edtext{sin\textit{n}a}{\Afootnote{þeira  L1572.}} \edtext{hýbýla}{\Afootnote{heim kinna  L679.}}, o\textit{c} sỏgþo a\textit{t} byrgja mætti  polla  \edtext{o\textit{c}}{\Afootnote{enn  L679.}} brúa t\textit{or}f\textit{ær}or. Gr\textit{eipr} bóndi b\textit{at}
h\textit{an}a þ\textit{ar} dvelja  o\textit{c} hjá sér  \edtext{a\textit{t}}{\Afootnote{÷  L679.}} þyggja  \edtext{þ\textit{at}}{\Afootnote{+ hun  L679.}} villdi. Svanh\textit{vít} \edtext{s\textit{egir} þa\textit{ck}a}{\Afootnote{þackar  L679, L1572.}} bónda  \edtext{f\textit{yrir} betsto sæmd\textit{ar}}{\Afootnote{gott  L679.}} boþ, o\textit{c} segist leynd\textit{ar}mál v\textit{it} Hróm\textit{und} ega, 
Vel er þ\textit{at} s\textit{egir} Gr\textit{eipr} o\textit{c} \edtext{man h\textit{ann}}{\Afootnote{munum vér  L1572.}} 
\edtext{skyldr}{\Afootnote{+ ad veita ydr  L679, skildir  L1572.}}, fremr ỏþr\textit{om},  \edtext{yþr a\textit{t} v\textit{ei}ta}{\Afootnote{÷  L679.}}. Hróm\textit{undr} o\textit{c} Svanh\textit{vít} g\textit{en}go  \edtext{þá}{\Afootnote{÷  L679.}} \edtext{bæþi}{\Afootnote{÷  L2404.}} á einmæli,  Hún m\textit{ælir} þ\textit{ess} eyrindis  \edtext{q\textit{vo}m e\textit{c} nú at fin\textit{n}a
	þ\textit{ic}}{\Afootnote{er eg kominn hingad  L679.}}, a\textit{t} mér er  sagt a\textit{t} Br\textit{óþir}
m\textit{inn} sé n\textit{ær} \edtext{þ\textit{ví}}{\Afootnote{÷  L679.}} á flótta  \edtext{r\textit{ic}in\textit{n}}{\Afootnote{kominn  L679.}},  \edtext{e\textit{n}}{\Afootnote{og  L679.}} m\textit{ar}gt af h\textit{onom} ||\ledouternote{120r} fall\textit{it} liþ h\textit{an}s, villda e\textit{c} \edtext{nú}{\Afootnote{÷  L679.}} a\textit{t} þú f\textit{yrir} min\textit{n} vinskap, adstoþ   veitt\textit{ir}  Ol\textit{afi} \edtext{K\textit{óngi}}{\Afootnote{÷  L679.}} \edtext{br\textit{óþr} m\textit{ínom}}{\Afootnote{÷  L2404.}}  
Ofús  er e\textit{c} \edtext{Ol\textit{afi} K\textit{óngi}}{\Afootnote{honum  L2404.}} liþ a\textit{t} veita, þó  e\textit{c} gjæti, s\textit{egir} \edtext{Hróm\textit{undr}}{\Afootnote{+ oc hellz  L2404.}} f\textit{yrir} þá skuld  \edtext{sérdeilisl\textit{e}ga}{\Afootnote{sierdeilis  L679, L1572, ÷  L2404.}} a\textit{t} h\textit{ann}  úthellti  blóþi þíno m\textit{eþ} \edtext{heipt\textit{ar}fullri}{\Afootnote{heiptar  L679.}} hendi, mín vegna,  
o\textit{c} er  \edtext{þ\textit{ví}}{\Afootnote{þad nú  L679.}} réttvíst a\textit{t} h\textit{ann} \edtext{nú}{\Afootnote{÷  L679, L2404.}} njóti þ\textit{ess}. Svanh\textit{vít} m\textit{ælir} fyrst sva  v\textit{ar},  \edtext{a\textit{t}}{\Afootnote{÷  L679.}} e\textit{c}  útstóþ  \edtext{ỏrkumslit}{\Afootnote{órkumsl  L679.}} m\textit{eþ}  vilja  f\textit{yrir}
þín\textit{ar} sak\textit{ir}, 
þá  manto f\textit{yrir} \edtext{mín\textit{ar}}{\Afootnote{+ sakir  L679.}} \edtext{t\textit{il}lỏgr}{\Afootnote{÷  L679.}} \edtext{liþsin\textit{n}a}{\Afootnote{+ kongi  L1572.}}. 
Hróm\textit{undr} m\textit{ælir}  jafn\textit{an}  sýn\textit{it} \edtext{þ\textit{it}}{\Afootnote{þer  L2404, L1572.}}, Jómfrú! yþ\textit{ar}  géþgjæþsko, 
e\textit{n}
v\textit{it} \edtext{dreng\textit{ir}}{\Afootnote{÷  L1572.}} okkar  h\textit{ar}þlyndi. 
Svanh\textit{vít} gé\textit{ck} þá  \edtext{a\textit{t} Hróm\textit{undi}}{\Afootnote{Rómundi at  L2404.}} o\textit{c} lagþi hendr u\textit{m} háls h\textit{onom}, o\textit{c} m\textit{ælir} e\textit{ck}i manto
þ\textit{ess}i orþ  \edtext{mæla af alvỏro}{\Afootnote{af alvóru tala  L679.}},  heldr í gamni. 
\edtext{Já}{\Afootnote{ju  L1572.}}, s\textit{egir} Hróm\textit{undr} f\textit{yrir} alv\textit{ỏr}o,
þ\textit{ví} f\textit{yrir} fulla  alv\textit{ỏr}o pústraþi
Ol\textit{afr} K\textit{óngr} þ\textit{ic}, o\textit{c} kallaþi þ\textit{ic} dubbo, o\textit{c} f\textit{er} e\textit{ck}i
ílla þó  \edtext{at h\textit{onom}, eþr h\textit{an}s góþo}{\Afootnote{hans  L679, honum oc hanz miclo  L2404, ad honum og  L1572.}} \edtext{g\textit{jỏr}p\textit{om}}{\Afootnote{+ hans  L1572.}}, blæþi or  \edtext{skrokk\textit{om}}{\Afootnote{kroppnum  L1572.}} líkt  s\textit{em} yþr, or ásjóno,  \edtext{o\textit{c}}{\Afootnote{eda  L679.}} \edtext{seppa}{\Afootnote{racka  L1572.}} mín\textit{om} Hrók t\textit{il} fulls Svanh\textit{vít} m\textit{ælir} o\textit{c} héldt e\textit{n} hỏnd\textit{om} u\textit{m} háls  \edtext{h\textit{an}s}{\Afootnote{honum  L679.}};  
tefþo  nú e\textit{ck}i
a\textit{t} gégna  mér, þ\textit{ess} er e\textit{c}
t\textit{il} \edtext{þín}{\Afootnote{÷  L679.}} mæli. Sva skal v\textit{er}a, s\textit{egir} Hróm\textit{undr} \edtext{o\textit{c}}{\Afootnote{÷  L2404.}} þó  e\textit{c} vissi,  \edtext{a\textit{t}}{\Afootnote{÷  L679, L1572.}} e\textit{c} mundi ei lífs aptr q\textit{vo}ma, þá vil e\textit{c}
gj\textit{ar}n\textit{an}, yþar vegna, Ol\textit{afi}
K\textit{óngi} liþ v\textit{ei}ta slíkt s\textit{em} e\textit{c} má; 
h\textit{ann} laut at  \edtext{Svanh\textit{vít}}{\Afootnote{henni  L679.}} o\textit{c} min\textit{n}tist v\textit{it} h\textit{an}a. 
Nú  \edtext{g\textit{jỏr}þ\textit{ir}}{\Afootnote{gjörir L679.}} þú v\textit{el}, s\textit{egir} hún,  \edtext{o\textit{c} er}{\Afootnote{nu er hier  L679, oc er hér  L2404, L1572.}} ein\textit{n} skjỏldr, er  \edtext{e\textit{c} hefi hug\textit{at} a\textit{t}}{\Afootnote{vil  L2404.}} géfa þér, o\textit{c} þygg h\textit{ann} \edtext{nú}{\Afootnote{÷  L1572.}},  h\textit{ann} man þér no\textit{ck}ot duga,  \edtext{kun\textit{n}\textit{ir}}{\Afootnote{kunna þá  L2404.}} \edtext{þú}{\Afootnote{+ berst  L2404.}} v\textit{it} Helga Frækna  \edtext{a\textit{t} b\textit{er}jast}{\Afootnote{÷  L2404.}}, o\textit{c} lát h\textit{ann} e\textit{ck}i þér frá hendi losna.
Hróm\textit{undr} m\textit{ælir} hefr þú heyrt a\textit{t} Helgi Frækni  \edtext{sé}{\Afootnote{muni vera  L679.}} í þ\textit{ess}\textit{om} b\textit{ar}daga?  Já,
s\textit{egir} hún. Þá er tvísýnt u\textit{m} sigr, s\textit{egir}
Hróm\textit{undr} þ\textit{ví} \edtext{a\textit{t}}{\Afootnote{÷  L679, L2404.}} h\textit{ann} er  hin  \edtext{meirsta}{\Afootnote{besta  L1572.}} kémpa. Lífs man\textit{om} v\textit{it} aptr sjást, s\textit{egir} hún, o\textit{c} t\textit{ac} nú v\textit{it} sokkab\textit{an}di
míno, o\textit{c} vef u\textit{m}  úlfliþ  þin\textit{n}, n\textit{ær} þú hỏggr t\textit{il} Helga,
o\textit{c} mano þér fá hỏgg  mistakast. Hróm\textit{undr}
þa\textit{ck}aþi hen\textit{n}i, o\textit{c} b\textit{at} h\textit{an}a  \edtext{orþlofs}{\Afootnote{orlofs  L2404.}} a\textit{t} f\textit{ar}a. Hún m\textit{ælti} f\textit{ar} þú
m\textit{ann}a lukko samastr, o\textit{c} q\textit{vo}m  heill  \edtext{eins}{\Afootnote{heim  L1572.}} aptr; skildo þau sva  \edtext{tal\textit{it}}{\Afootnote{+ og  L1572.}}, f\textit{er}þaþist Svanh\textit{vít} 
||\ledouternote{120v} 
heim  \edtext{aptr}{\Afootnote{÷ L679.}}  
m\textit{eþ} st\textit{ær}sto  \edtext{vyrþíng}{\Afootnote{vjrdinu  L679.}} af Gr\textit{eipi} bónda, o\textit{c} son\textit{om} h\textit{an}s. En\textit{n} Hróm\textit{undr} bjó f\textit{er}þ sína
austr t\textit{il} Væn\textit{ir}s; þ\textit{ei}r  \edtext{b\textit{ræþ}r  sp\textit{yrja}
	h\textit{ann}}{\Afootnote{spyrja hann brædr hans  L1572.}} \edtext{a\textit{t}}{\Afootnote{÷  L2404, L1572.}}, hv\textit{or}t h\textit{ann} vilji þ\textit{ei}r fylgi h\textit{onom}?
Han\textit{n} s\textit{egir} nei, þ\textit{ví} e\textit{ck}i vil 
e\textit{c} \edtext{a\textit{t} þ\textit{at}}{\Afootnote{þid  L679, ad þid  L1572.}} \edtext{hljót\textit{it}}{\Afootnote{hafid  L1572.}} slis f\textit{yrir} mín\textit{ar} sak\textit{ir}. Þat skal aldrei
v\textit{er}þa, s\textit{egja} þ\textit{ei}r, a\textit{t} þú  \edtext{ein\textit{n}}{\Afootnote{÷  L679.}} útdrag\textit{ir} t\textit{il} stríþs, e\textit{n} v\textit{it}
sitj\textit{om} \edtext{heima}{\Afootnote{÷  L679.}} ept\textit{ir},  eþa þú fall\textit{ir},  \edtext{e\textit{n}}{\Afootnote{og  L679.}} v\textit{it} ept\textit{ir} lif\textit{om}. Hróm\textit{undr}
m\textit{ælir} \edtext{Forlỏg\textit{om}}{\Afootnote{forlögunum  L679.}} v\textit{er}þr  frestat,  \edtext{e\textit{n}}{\Afootnote{+ ei  L1572.}} f\textit{yrir} \edtext{þau e\textit{ck}i}{\Afootnote{÷  L1572.}} \edtext{q\textit{vo}mist}{\Afootnote{+ og  L679.}}. Ept\textit{ir} \edtext{þetta}{\Afootnote{þad  L679.}} t\textit{ac}a þ\textit{ei}r b\textit{ræþ}r  a\textit{t} búa
f\textit{er}þ sína m\textit{eþ} Hróm\textit{undi} snúa þ\textit{ei}r þá
all\textit{ir} austr t\textit{il} Væn\textit{ir}s, v\textit{ar} þá Ol\textit{afr} K\textit{óngr}
a\textit{t} meirsto flúin\textit{n},  \edtext{s\textit{em} fyrr segir}{\Afootnote{÷  L2404.}}. \pend \endnumbering  
\beginnumbering \pstart  
\vspace{5mm}\subsection*{XIII kap\textit{ítuli}} 
Hróm\textit{undr} \edtext{lætr setja}{\Afootnote{lætr nú setja L679, setr  L1572.}} tjald sitt  \edtext{hino}{\Afootnote{hinum  L679.}} megin vatnsins, o\textit{c} lét v\textit{er}a hraukstjald;
þetta v\textit{ar}
\edtext{á áliþn\textit{om}}{\Afootnote{ad lidnum L679.}} 
 degi,  h\textit{ann} leggr
s\textit{ic} t\textit{il} svefns, o\textit{c} lætr ílla í svefni;
a\textit{t} m\textit{or}gni vakn\textit{ar} h\textit{ann},
o\textit{c} blæs mæþil\textit{e}ga; þ\textit{ei}r b\textit{ræþ}r
sp\textit{yrja} h\textit{ann} ept\textit{ir} draum\textit{om} sín\textit{om},  \edtext{e\textit{n}}{\Afootnote{÷  L679.}} h\textit{ann} q\textit{vaþ}st eng\textit{om} s\textit{egja} o\textit{c} man e\textit{c} ei út f\textit{ar}a á þ\textit{ess}\textit{om} degi,\footnote{L2404 repeats the sentence \enquote{hann leggr sic til svefns oc lętr illa i	svefni}.}  \edtext{s\textit{egir} h\textit{ann}}{\Afootnote{og  L679.}}, þótt aþr\textit{ir} \edtext{f\textit{ar}i}{\Afootnote{+ segir hann  L679.}}. Þeir b\textit{ræþ}r s\textit{egja} þá skul\textit{om} v\textit{it} samt f\textit{ar}a,  er o\textit{ss} þ\textit{at} m\textit{ei}ri
skỏm\textit{m}, a\textit{t} sitja hér í tjaldi v\textit{or}o, þeg\textit{ar}
aþr\textit{ir} v\textit{it} egast, e\textit{n} þ\textit{at}, 
\edtext{þó}{\Afootnote{þótt  L2404, L1572.}} vér fỏll\textit{om} m\textit{eþ} góþ\textit{om} orþstýr. 
Hróm\textit{undr} m\textit{ælir} þ\textit{it} manot ráþa  \edtext{v\textit{er}þa}{\Afootnote{+ en  L1572.}}, e\textit{c} hefi sjálfr mín ráþ;  v\textit{er}þr nú
Hróm\textit{undr} ept\textit{ir}, e\textit{n} þ\textit{ei}r b\textit{ræþ}r  \edtext{h\textit{an}s}{\Afootnote{÷  L679, L1572.}} ríþa t\textit{il} m\textit{anna} Ol\textit{afs} K\textit{óngs} o\textit{c} sækja  \edtext{bæþi v\textit{el} o\textit{c} lengi}{\Afootnote{÷  L1572.}} h\textit{ar}þt fram\textit{m},  \edtext{fella}{\Afootnote{feldu  L1572.}} þ\textit{ei}r  \edtext{þá}{\Afootnote{so  L1572.}} m\textit{ar}g\textit{an}\textit{n} m\textit{an}\textit{n},  \edtext{sva}{\Afootnote{ad  L1572.}} undr\textit{om} gégndi,  \edtext{v\textit{ar}þ}{\Afootnote{var  L1572.}} þá m\textit{ar}gr maþr  klofin\textit{n} a\textit{t}  endilaungo, 
o\textit{c} m\textit{ar}gr fótlestr o\textit{c} \edtext{h\textit{an}dar}{\Afootnote{hand  L679, L2404.}} hỏggvin\textit{n},  \edtext{ásamt hỏfutlaus}{\Afootnote{oc höfutlaus  L2404, ÷  L1572.}}, mátti þá vaþa  \edtext{bæþi}{\Afootnote{÷  L2404, L1572.}} \edtext{hesta o\textit{c} m\textit{anna} blóþ}{\Afootnote{manna blod og hesta  L679, manna blóþ  L2404.}} s\textit{em} vatn v\textit{ær}i, héldo þ\textit{ei}r
þ\textit{ess}\textit{om} ákafa o\textit{c} fram\textit{m}sókn t\textit{il} mids apt\textit{an}s, þá dró upp þykkni  \edtext{í}{\Afootnote{á  L1572.}} \edtext{lopt\textit{it}}{\Afootnote{lopti  L679.}}, sva  \edtext{a\textit{t} skuggaþi}{\Afootnote{picknade  L679.}}, þ\textit{ei}r lito þá í lopt upp, o\textit{c} sáo 
st\textit{or}a Álft fljúga yf\textit{ir} sér, hún blés sva st\textit{er}kom
anda a\textit{t} þ\textit{ei}m b\textit{ræþ}r\textit{om},  \edtext{sva}{\Afootnote{ad  L679.}} þá t\textit{óc} a\textit{t}  krimpa   \edtext{v\textit{it}}{\Afootnote{÷  L1572.}} af hita, 
||\ledouternote{121r} o\textit{c} stund\textit{om} af kulda; þ\textit{ess}i álft hafþi sva
st\textit{or}a vængi,  a\textit{t} þ\textit{ei}r náþo
yf\textit{ir} \edtext{M}{\Afootnote{C  L679.}} feta  \edtext{breiþa}{\Afootnote{÷  L679.}} j\textit{ar}þ\textit{ar} víþd á iv vega  \edtext{mælt}{\Afootnote{÷  L1572.}}, stund\textit{om} saung þ\textit{ess}i fugl sva fagurt, a\textit{t}
m\textit{enn} gáþo ei  dauþa síns o\textit{c} stóþo s\textit{em} \edtext{a\textit{t}}{\Afootnote{÷  L2404.}} skémtun v\textit{ær}i á f\textit{er}þ\textit{om},  \edtext{o\textit{c}}{\Afootnote{÷  L679, L2404.}} \edtext{h\textit{it} besta}{\Afootnote{hid mesta  L679, ÷  L2404, hin mesta  L1572.}} 
\edtext{gleþi spil}{\Afootnote{÷  L2404, gledi  L1572.}}. 
Vit þetta v\textit{er}þa þ\textit{ei}r b\textit{ræþ}r v\textit{ar}ir at
Helgi  Frækni hỏggr niþr liþ Ol\textit{afs} K\textit{óngs}
þ\textit{ei}r  \edtext{ríþa}{\Afootnote{+ þá  L1572.}} fram\textit{m} hv\textit{or} s\textit{em} \edtext{betst}{\Afootnote{mest  L679, meira  L1572.}} gétr, o\textit{c} hỏggva o\textit{c} leggja þá s\textit{em}
f\textit{yrir} v\textit{er}þa; þetta  \edtext{sá}{\Afootnote{sier  L679, L1572.}} Helgi, o\textit{c} ríþr á móti þ\textit{ei}m m\textit{eþ} meirsta
kappi, dró  þá yf\textit{ir} \edtext{skugg\textit{an}\textit{n}}{\Afootnote{scugga  L2404.}} o\textit{c} \edtext{dym\textit{m}o}{\Afootnote{dimmuna  L679.}} á sólun\textit{n}i m\textit{eþ} sæt\textit{om} o\textit{c}
yndisl\textit{e}gom Svana  \edtext{lát\textit{om}}{\Afootnote{saung  L2404.}}, sva alla furþaþi, bæþi N\textit{or}þm\textit{enn} oc  Svía;
syn\textit{ir} Gr\textit{eips} fylgþost all\textit{ir} \edtext{a\textit{t}}{\Afootnote{÷  L2404.}} í ein\textit{om} flo\textit{ck}i, o\textit{c} gafo sic  minst
a\textit{t} \edtext{álft\textit{ar}}{\Afootnote{svana  L679.}} raustin\textit{n}i, 
e\textit{n} hún lét sva fagort, a\textit{t} m\textit{ỏnnom} \edtext{féllst u\textit{m}}{\Afootnote{flesta fiellust  L679.}} v\textit{er}k sín; 
Miþt í þ\textit{ess}ari saunglyst kémr 
Helgi a\textit{t} þ\textit{ei}m b\textit{ræþ}r\textit{om} o\textit{c} hỏggr
t\textit{il} Haka o\textit{c}  klýfr  h\textit{an}n at  \edtext{endilaungo}{\Afootnote{+ oc fell Haki  L2404.}}; þá  \edtext{hitti}{\Afootnote{hittir  L679.}} h\textit{ann} Gaut,  \edtext{af h\textit{onom} sneiþ h\textit{ann} hỏnd o\textit{c}
	síþuna}{\Afootnote{og hjö af hónum sijduna og hóndina  L679, oc af honum sneiþ hann höndina meþ sýþonni  L2404.}}; þ\textit{ar} ept\textit{ir} v\textit{ar}þ Haukr
f\textit{yrir} h\textit{onom},  \edtext{e\textit{n}}{\Afootnote{÷  L679, L1572, oc  L2404.}} \edtext{Helgi vóg}{\Afootnote{vóg Helgi  L2404.}} h\textit{ann}, oc  kastaþi  \edtext{h\textit{onom}}{\Afootnote{+ daudum  L679, L1572, daudum  L2404.}} í valin\textit{n}; þá drap h\textit{ann} Þrỏst,  \edtext{þá}{\Afootnote{+ vóg hann  L2404.}} Helga  \edtext{oc}{\Afootnote{þa  L679,  + lagþi i gegnom  L2404.}} Loga, 
þá klauf h\textit{ann} Héþin\textit{n}, o\textit{c} síþ\textit{ar}st allra  \edtext{þ\textit{eir}ra}{\Afootnote{÷  L679.}} \edtext{hjó}{\Afootnote{vog  L679.}} h\textit{an}n  Angantýr  \edtext{t\textit{il} b\textit{an}a}{\Afootnote{÷  L679.}}, v\textit{or}o þá falln\textit{ir} all\textit{ir} b\textit{ræþr}
Hróm\textit{undar} o\textit{c} hỏfþo enga v\textit{or}nveitt Helga eþr mótstỏþo,
o\textit{c} er  fróþra m\textit{anna} sỏgn, a\textit{t}
alft\textit{ar} saungurin\textit{n} \edtext{mani h\textit{afa}}{\Afootnote{hafi  L679.}}  glapit   þá; ept\textit{ir}
þetta slotaþi sv\textit{an}a látun\textit{om}, o\textit{c} dró burt 
þykkn\textit{it} o\textit{c} dym\textit{m}una sn\textit{er}ist þá
m\textit{an}\textit{n}fall\textit{it} á Ol\textit{af} K\textit{óng} oc  v\textit{ar}aþi  \edtext{þ\textit{at}}{\Afootnote{+ allt  L679.}} t\textit{il} qvỏlds,  \edtext{o\textit{c}}{\Afootnote{÷  L679.}} lauk sva b\textit{ar}dagan\textit{om} ásamt 
degin\textit{om}, lét þá Ol\textit{afr} K\textit{óngr} \edtext{binda}{\Afootnote{+ um  L679, L1572.}} sár m\textit{anna} sin\textit{n}a, 
þ\textit{ví} m\textit{ar}g\textit{ir} v\textit{or}o sár\textit{ir}, hafþi þá sva fall\textit{it} liþ
h\textit{an}s, a\textit{t} \edtext{ept\textit{ir} v\textit{ar} ekki}{\Afootnote{ecki var eptir  L679, ei voro eptir  L2404.}} nema hálft an\textit{n}\textit{at} hundraþ, þókti h\textit{onom} \edtext{e\textit{n} s\textit{em} opt\textit{ar}}{\Afootnote{þá sem flestum  L679, enn  L1572.}}, ílla  \edtext{fyr\textit{ir} sér áh\textit{or}fast}{\Afootnote{áhorfast fijrir sier og  L679, + og  L2404, L1572.}}, v\textit{ar} þ\textit{ví} h\textit{ar}tn\textit{ær} flúin\textit{n}.  || \ledouternote{121v}
\pend \endnumbering
  \beginnumbering \pstart 
\vspace{5mm}  \subsection*{XIV kap\textit{ítuli}}
  Þetta sama qvỏld q\textit{vo}m sendim\textit{aþr} Ol\textit{afs}
K\textit{óngs} m\textit{eþ} liþ þ\textit{at} er h\textit{ann} hafþi t\textit{il} qvaþt v\textit{ar}þ  \edtext{þá Ol\textit{afr}}{\Afootnote{÷  L679.}} K\textit{óngr} þ\textit{ess}\textit{om} fegin\textit{n} o\textit{c} þóktist  nú allra meina bættr, e\textit{n} þó h\textit{ar}maþi
h\textit{ann} meirst þá Bíld  o\textit{c} Vola. Nú q\textit{vo}mo
Hróm\textit{undi} \edtext{frétt\textit{ir}}{\Afootnote{fregnir  L679.}}, a\textit{t} \edtext{þ\textit{ei}r}{\Afootnote{÷  L1572.}} b\textit{ræþ}r h\textit{an}s v\textit{ær}o  falln\textit{ir}
f\textit{yrir} Helga, h\textit{ann} qv\textit{aþ} s\textit{ic} sva
grun\textit{at} \edtext{hefþi}{\Afootnote{hafa  L679, L2404, L1572.}} a\textit{t} fara  m\textit{un}di; e\textit{n} qv\textit{aþ}
h\textit{ann} \edtext{hin\textit{n} meirsta kappa}{\Afootnote{hinna mestan kapps  L679, L2404, L1572.}} o\textit{c} hreysti  \edtext{man\textit{n}}{\Afootnote{vegna  L679, L2404, L1572.}}.  Um m\textit{or}gunin\textit{n} \edtext{býst}{\Afootnote{bióst  L679.}} Hróm\textit{undr} snem\textit{m}a t\textit{il} orrusto, o\textit{c}
t\textit{ec}r  vopn sín.  \edtext{K\textit{óngr} Ol\textit{afr}}{\Afootnote{Olafur Kongur  L679, L2404, L1572.}} v\textit{ei}t þá e\textit{ck}i af Hróm\textit{undi} a\textit{t} segja,
þ\textit{ví} h\textit{ann} hafþi tjald\textit{at} hinomegin\textit{n}
Væn\textit{ir}s, o\textit{c} \edtext{þó}{\Afootnote{þad  L679.}} afsíþis. Ol\textit{afr} K\textit{óngr} lét  \edtext{t\textit{il} orrusto}{\Afootnote{í lúdr  L1572.}} blása,  \edtext{furþaþi}{\Afootnote{furdar  L679.}} þá Svía, a\textit{t} K\textit{óngr} skyldi  sva
m\textit{ar}gt liþ e\textit{n} \edtext{þá}{\Afootnote{÷  L679, L2404.}} \edtext{ept\textit{ir}}{\Afootnote{÷  L679.}} h\textit{afa}, dróu þ\textit{ei}r  \edtext{út sva}{\Afootnote{so út  L679, L1572, oc út meþ  L2404.}} sitt líþ  m\textit{eþ} lúþra blæstri o\textit{c} vopna braki,
\edtext{t\textit{óc}}{\Afootnote{tókst  L679, L2404.}} nú  \edtext{t\textit{il}}{\Afootnote{÷  L679, L2404.}} hin\textit{n} h\textit{ar}þasti b\textit{ar}dagi,  \edtext{sva}{\Afootnote{÷  L679.}} s\textit{em} fyrri, t\textit{óc} \edtext{sva}{\Afootnote{nú  L2404, þá enn  L1572.}} a\textit{t} snúast  \edtext{m\textit{an}\textit{n}fall\textit{it}}{\Afootnote{mannfallinu  L679.}} á  Ol\textit{af} K\textit{óng} þ\textit{ví} Halfdán
o\textit{c} Helgi feldo alla  \edtext{þá}{\Afootnote{÷  L679.}} \edtext{s\textit{em}}{\Afootnote{er  L1572.}} f\textit{yrir} urþo,  o\textit{c} \edtext{stóþst}{\Afootnote{stód  L679.}} engin\textit{n} v\textit{it} þ\textit{ei}m af
N\textit{or}þm\textit{ỏnnom}. Hróm\textit{undr} v\textit{er}þr
þ\textit{ess}a  vís o\textit{c} t\textit{ec}r  \edtext{h\textit{ann}}{\Afootnote{÷  L679.}} \edtext{þá til}{\Afootnote{þá  L679, til  L2404.}} a\textit{t} britja  \edtext{stór\textit{om}}{\Afootnote{stort  L1572.}}, \edtext{bæþi a\textit{t} hỏggva oc leggja}{\Afootnote{÷  L679.}}, hrukko þá all\textit{ir} und\textit{an} s\textit{em} gáto
f\textit{or}þ\textit{at} sér, o\textit{c} víldo  flýa,
e\textit{n} \edtext{Fyr\textit{ir}sjón\textit{ar}m\textit{enn}}{\Afootnote{fyrer menn  L679, höfþingiar  L2404.}} Svía báþo ei uppgéfast, sýndist  þá fleyrst\textit{om}
a\textit{t} sýna dád o\textit{c} \edtext{drengskap}{\Afootnote{dugnad  L1572.}}, ept\textit{ir} þ\textit{ví} s\textit{em} hvỏr  v\textit{ar}
m\textit{aþr} t\textit{il}. Nú kémr Hróm\textit{undr} þ\textit{ar}
a\textit{t} s\textit{em} \edtext{þ\textit{ei}r}{\Afootnote{÷  L679, L1572.}} \edtext{b\textit{ræþ}r h\textit{an}s}{\Afootnote{hanns brædur  L679.}} hỏfþo  fall\textit{it} \edtext{dagin\textit{n}}{\Afootnote{deginum  L1572.}} áþr í Valin\textit{n}\footnote{In L679 dittography \enquote{í valinn}.},  \edtext{h\textit{ann} brá}{\Afootnote{brá hann  L679.}} sér lítt v\textit{it}, o\textit{c} m\textit{ælir} þ\textit{at} man  líkast a\textit{t} e\textit{c} leiti ept\textit{ir} \edtext{y\textit{ck}r hefndar}{\Afootnote{hefndir yckar  L679, yckar hefndum  L1572.}}, 
sva s\textit{em} Helgi ept\textit{ir} Hraungviþ, litast h\textit{ann} nú  \edtext{u\textit{m}}{\Afootnote{vid  L1572.}} a\textit{t} fin\textit{n}a h\textit{ann}, o\textit{c} loksins sér  h\textit{ann} 
a\textit{t} Helgi er skamt   \edtext{á burt}{\Afootnote{i burt þadan  L679, i burt L1572.}}; 
h\textit{ann} ren\textit{n}r  \edtext{á skunda}{\Afootnote{med skunda  L679, ÷  L2404, af skunda  L1572.}} á  \edtext{móts v\textit{it} h\textit{ann}}{\Afootnote{á möti hónum enn  L679, + og  L1572.}}; er Helgi  \edtext{sjer}{\Afootnote{sá  L679.}} h\textit{ann},  \edtext{s\textit{egir}}{\Afootnote{mælir  L1572.}} h\textit{ann}: hér erto nú q\textit{vo}min\textit{n}, 
s\textit{em} gafst Hraungv\textit{iþ} \edtext{f\textit{or}þ\textit{om}}{\Afootnote{÷  L2404.}} hỏgg\textit{it} m\textit{eþ} kylfun\textit{n}i, ut\textit{an} und||\ledouternote{122r}ir v\textit{an}g\textit{an}\textit{n}, 
sva h\textit{ann} hefr
e\textit{ck}i vakn\textit{at} síþ\textit{an},  \edtext{hefi}{\Afootnote{hef  L679, L2404, L1572.}} e\textit{c} lítit hefnt h\textit{an}s á
b\textit{ræþ}r\textit{om} þín\textit{om}, jafnv\textit{el} þó þú
lét\textit{ir} græþa sár mín  Hróm\textit{undr} m\textit{ælir}
m\textit{ỏr}g\textit{om} hefr betr  \edtext{far\textit{it}}{\Afootnote{farist  L679.}} e\textit{n} þér, o\textit{c} manto ílla  \edtext{u\textit{m}bun}{\Afootnote{umbuna  L1572.}} \edtext{hljóta}{\Afootnote{Helgi mælir  L1572.}}. 
Heit mín v\textit{ar}þ e\textit{c} a\textit{t} efna,  \edtext{s\textit{egir} Helgi}{\Afootnote{÷  L679, L1572.}} hv\textit{aþ} s\textit{em} ỏþro leiþ, o\textit{c} mátto vyrþa á
þ\textit{ann} veg, s\textit{em} þú villt,  e\textit{n}
e\textit{ck}i hræþist e\textit{c} þ\textit{ic},  \edtext{þó a\textit{t}}{\Afootnote{þott  L679.}} þú haf\textit{ir} sókt Mistiltein í  haug Þráins, o\textit{c}
n\textit{ær} mátto g\textit{an}ga ef þú villt hefna 
Fr\textit{ænda} þin\textit{n}a. Hróm\textit{undr} m\textit{ælir} þú mant
lítt  \edtext{t\textit{il}}{\Afootnote{÷  L679, L1572.}} f\textit{ær} f\textit{yrir} þá sỏk,  a\textit{t} þú hefr  \edtext{m\textit{eþ} f\textit{er}þis sva þúngt}{\Afootnote{þungt vopn med ferdis  L679.}} o\textit{c} breiþt  \edtext{sverþ}{\Afootnote{÷  L679.}}, a\textit{t} þú k\textit{an}t  þ\textit{ví} e\textit{ck}i
a\textit{t} valda,  \edtext{o\textit{c}}{\Afootnote{÷  L679.}} \edtext{f\textit{yrir}}{\Afootnote{÷  L2404.}} þ\textit{ví} v\textit{il} e\textit{c} ljá þér Mistiltein s\textit{em} er
m\textit{ic}lo  létt\textit{ar}i. Helgi m\textit{ælir}
e\textit{ck}i þ\textit{ar}fto a\textit{t} frýa mér u\textit{m} burþa leysi
þ\textit{ví} m\textit{eþ} þ\textit{ess}o  \edtext{sama}{\Afootnote{÷  L679.}} vopni helsló e\textit{c} \edtext{alla þína}{\Afootnote{allar þínar  L2404.}} b\textit{ræþ}r, o\textit{c} v\textit{ar}þ þ\textit{at} \edtext{alblóþugt}{\Afootnote{blodúgt  L679.}} a\textit{t} hjỏlt\textit{om}. 
J  \edtext{þ\textit{ess}o}{\Afootnote{þessum  L679, L2404.}} \edtext{sama}{\Afootnote{÷  L1572.}} bili sló 
yf\textit{ir} \edtext{dym\textit{m}o o\textit{c} myrkri o\textit{c} sæt\textit{om} sv\textit{an}a hljóþ\textit{om}}{\Afootnote{miklum og yndislegúm svana látum og dimmum  L679.}},  
\edtext{f\textit{or} þá s\textit{em} fyrr, a\textit{t}}{\Afootnote{sem fyrr so  L1572.}} m\textit{enn} gáþo sín e\textit{ck}i, s\textit{em} \edtext{skyldi}{\Afootnote{skilldu  L679.}}, e\textit{n} Hróm\textit{undr} gaf s\textit{ic} e\textit{ck}i a\textit{t} \edtext{þ\textit{ess}o}{\Afootnote{þvj  L679.}}; þá t\textit{óc} \edtext{álftin}{\Afootnote{+ helldr  L679.}} a\textit{t} v\textit{er}þa  \edtext{heldr}{\Afootnote{÷  L679.}} m\textit{ann}vỏn, o\textit{c} settist  \edtext{niþr}{\Afootnote{+ helldr  L679.}} n\textit{ær}ri þ\textit{ei}m, e\textit{n} stund\textit{om} flaug  \edtext{hún}{\Afootnote{+ upp  L679.}} skam\textit{m}t  \edtext{í lopti}{\Afootnote{÷  L679.}}, f\textit{yrir} of\textit{an} hỏfuþ m\textit{anna}. 
Nú t\textit{óc}o
þ\textit{ei}r Helgi o\textit{c} Hróm\textit{undr} a\textit{t} b\textit{er}jast m\textit{eþ}
hin\textit{om} meirsta  \edtext{ákafa}{\Afootnote{+ og á eggjann  L679.}}, o\textit{c} var  eigi gam\textit{an}
a\textit{t} h\textit{or}fa á þ\textit{ei}rra v\textit{it}skipti,
v\textit{or}o  \edtext{þ\textit{ar}}{\Afootnote{þeir  L679, L2404.}}  báþ\textit{ir} jafn  \edtext{vopnfim\textit{ir}}{\Afootnote{vigfimir  L679.}} o\textit{c} h\textit{ar}þfeng\textit{ir}; þá v\textit{ar}þ Helga
lit\textit{it} á hỏnd Hr\textit{omundi} \edtext{h\textit{ann}}{\Afootnote{og  L1572.}} m\textit{ælir} ílla manto v\textit{it} una  \edtext{kuld\textit{an}\textit{n}}{\Afootnote{kulda  L679.}},  \edtext{þá}{\Afootnote{þar  L679, því  L2404.}} þú  \edtext{⸌hefr⸍}{\Afootnote{+ tekid  L679.}} \edtext{so\textit{ck}ab\textit{an}d}{\Afootnote{sockabandid  L679.}} \edtext{kj\textit{ær}osto
	þin\textit{n}\textit{ar}}{\Afootnote{af unnústú þinni  L679.}} 
\edtext{tek\textit{it}}{\Afootnote{÷  L679.}} m\textit{eþ} þér a\textit{t} skilnaþi, o\textit{c} er þ\textit{at}
þ\textit{ví} a\textit{t} þa\textit{ck}a, e\textit{n} e\textit{ck}i þin\textit{n}i hreysti, þó a\textit{t} þú
v\textit{er}þ\textit{ir} hv\textit{or}ki sár né móþr, o\textit{c} er þ\textit{at} \edtext{san\textit{n}f\textit{er}þugt}{\Afootnote{sannferdugast  L679, satt  L2404, sannast  L1572.}}, a\textit{t} \edtext{þú}{\Afootnote{÷  L2404.}} trú\textit{ir} á  \edtext{h\textit{an}a}{\Afootnote{þá qonu  L2404.}}. Vit þetta reiþdist  Hróm\textit{undr} o\textit{c} hỏggr
t\textit{il} Helga o\textit{c} í skjỏldin\textit{n}, o\textit{c} klýfr
h\textit{an}n,  Þesso veldr trú þín á
so\textit{ck}ab\textit{an}dino, s\textit{egir} Helgi, o\textit{c} h\textit{afa}
e\textit{ck}i  all\textit{ir} slíko a\textit{t} hrósa. Þá sleit
Hróm\textit{undr} b\textit{an}d\textit{it} af hendi sér, ||\ledouternote{122v} o\textit{c} kastaþi ⸌niþr⸍ Helgi hjó t\textit{il}
Hróm\textit{undar} o\textit{c} \edtext{stemndi}{\Afootnote{hió  L1572.}} á ỏxlina, 
Hróm\textit{undr} brá sér lít\textit{it}
und\textit{an} \edtext{á hliþina}{\Afootnote{÷  L679.}}, o\textit{c} gat e\textit{ck}i sv\textit{er}þ\textit{it} ram\textit{m}at  \edtext{h\textit{an}a}{\Afootnote{hann  L679, L2404.}}, þó fé\textit{ck} Hróm\textit{undr} sár á l\textit{ær}it.
Hróm\textit{undr} villdi þá aptr hỏggva t\textit{il} Helga,
e\textit{n} h\textit{ann} \edtext{b\textit{ar}}{\Afootnote{bra  L679.}} \edtext{f\textit{yrir} skjỏldin\textit{n}}{\Afootnote{vid skilldinum  L679.}}, o\textit{c} rendi af  h\textit{onom} fjórþa  \edtext{p\textit{ar}tin\textit{n}}{\Afootnote{part  L679.}} o\textit{c} q\textit{vo}m  \edtext{v\textit{it} kálf\textit{an}}{\Afootnote{÷  L679.}},  \edtext{o\textit{c} nam af ut\textit{an} fótar vỏþva}{\Afootnote{á fotinn útann verdann og nam af vódvann  L679.}}. Þá hjó Helgi t\textit{il} Hróm\textit{undar} o\textit{c} stemndi á hjálmin\textit{n}. 
Hróm\textit{undr} ljóp í lopt upp aptr á bak,  \edtext{v\textit{ar}þ þá}{\Afootnote{og vard  L679.}} Alftin  f\textit{yrir} h\textit{onom}, h\textit{ann}
hrỏslaþist yf\textit{ir} h\textit{an}a, e\textit{n} Helgi sókti  \edtext{á}{\Afootnote{÷  L679.}} eptir,  Alftin flaug upp  \edtext{líka}{\Afootnote{i þessu  L1572.}} o\textit{c} v\textit{ar}þ f\textit{yrir} sv\textit{er}þino o\textit{c}
t\textit{óc} þ\textit{at} í sundur í  hen\textit{n}i
l\textit{ær}leggin\textit{n}, datt hún þá í kono líki daud niþr.  Þá m\textit{ælir} Hróm\textit{undr} níþíngr erto  \edtext{Helgi}{\Afootnote{÷  L679.}} í  \edtext{fleyro}{\Afootnote{fleiri  L1572.}} e\textit{n}eino,  fyrst a\textit{t} launa ílla lækníng þína
v\textit{it} m\textit{ic}, o\textit{c} síþ\textit{an} a\textit{t} fjỏr fletta  \edtext{f\textit{an}gkono}{\Afootnote{filgikonu  L679.}} \edtext{þína}{\Afootnote{+ Helgi mælti  L679.}}. Engin\textit{n} \edtext{em}{\Afootnote{er  L679, L1572.}} e\textit{c} níþíngr f\textit{yrir} græþslo þína,  \edtext{s\textit{egir} Helgi}{\Afootnote{÷  L679, L2404, L1572.}}, heldr sýndi e\textit{c} æro í  \edtext{loforþi}{\Afootnote{+ mijnu  L679, L2404.}} v\textit{it} Fr\textit{ænda} min\textit{n}, hv\textit{or}n e\textit{c} elskaþi meir e\textit{n} \edtext{líf}{\Afootnote{lijfid  L679.}} mitt; 
e\textit{n} þ\textit{at} s\textit{em} Vinkono  mína
hendti, man hún  \edtext{mér f\textit{or}láta}{\Afootnote{forlata mier fridsamlega  L679.}},  \edtext{þá}{\Afootnote{+ er  L679.}} v\textit{it} \edtext{fin\textit{n}umst}{\Afootnote{finnustum  L679.}} \edtext{friþsaml\textit{e}ga}{\Afootnote{÷  L679, L2404.}} í Valhỏll, o\textit{c} gyldti eino hv\textit{or} h\textit{ann}
v\textit{ar}, s\textit{em} \edtext{lauk  upp}{\Afootnote{upp lauk  L679.}} f\textit{yrir} hen\textit{n}i, e\textit{n} e\textit{c} man q\textit{vo}ma
t\textit{il} hen\textit{n}\textit{ar} sn\textit{ar}l\textit{e}ga. Þá ljóp
H\textit{elgi} a\textit{t} Hróm\textit{undi} o\textit{c} hjó sva tíþt a\textit{t}
h\textit{onom}, a\textit{t} Hróm\textit{undr} g\textit{jỏr}þi ei betr
e\textit{n} \edtext{v\textit{er}jast}{\Afootnote{verja sig fijrir hónum  L679.}} 
m\textit{eþ} skyldin\textit{om} þ\textit{ar} t\textit{il} \edtext{a\textit{t} Hróm\textit{undr}}{\Afootnote{hann  L679, Hrómundr  L1572.}} héldt  \edtext{ept\textit{ir}}{\Afootnote{+ á  L679.}} tóm\textit{om} mundriþ\textit{an}um, o\textit{c} litl\textit{om} p\textit{ar}ti þ\textit{ar} m\textit{eþ}. 
Þá s\textit{egir} Helgi: nú er ec 
búin\textit{n} a\textit{t} hỏggva Fulltrúa þin\textit{n} \edtext{hin\textit{n}}{\Afootnote{÷  L679, þann hinn  L1572.}} an\textit{n}\textit{an}, o\textit{c} sjerþo nú  betst,
hv\textit{aþ} h\textit{ann} dug\textit{ir}. Hróm\textit{undr} gleimdi þá
orþ\textit{om} Svanhv\textit{ítar} o\textit{c} fleygþi  \edtext{niþr}{\Afootnote{÷  L1572.}} þ\textit{ví} s\textit{em} \edtext{h\textit{ann} héldt}{\Afootnote{÷  L679.}} ept\textit{ir} \edtext{af skyldin\textit{om}}{\Afootnote{var  L679.}}, o\textit{c} hỏggur  \edtext{af reiþi mót Helga}{\Afootnote{á moti Helga af reidi  L679.}}, e\textit{n} h\textit{ann} v\textit{ar}þist ágjæta v\textit{el},
o\textit{c} gé\textit{ck} sva lengi þ\textit{eir}ra á milli, o\textit{c} þókti  \edtext{fleyrst\textit{om}}{\Afootnote{÷  L1572.}} s\textit{em} \edtext{eldíngar}{\Afootnote{ellding  L679.}} flýgi af  \edtext{sv\textit{er}þ\textit{om} þ\textit{eir}ra}{\Afootnote{sverdi hans  L679.}}. 
Eitt sin\textit{n} hjó Helgi t\textit{il} Hróm\textit{undar} oc  stemndi í andlit h\textit{an}s. Hróm\textit{undr} veik sér  \edtext{á}{\Afootnote{um  L1572.}} \edtext{hop}{\Afootnote{hæl  L679.}} aptr á bak,  e\textit{n} sv\textit{er}þ\textit{it} gat
sn\textit{or}t\textit{it} á Hróm\textit{undi} hold\textit{it} á  \edtext{bríngun\textit{n}i}{\Afootnote{bringuna  L679.}}, o\textit{c} \edtext{sva ljóp}{\Afootnote{hliop so  L1572.}} ||\ledouternote{123r} \edtext{þ\textit{at}}{\Afootnote{÷  L1572.}} of\textit{an} all\textit{an} qviþin\textit{n}, o\textit{c} risti
h\textit{ann} in\textit{n} a\textit{t} \edtext{Jstro}{\Afootnote{Jstrum  L679.}} o\textit{c} \edtext{sva á kaf}{\Afootnote{sóck þad  L1572.}} of\textit{an} í  vỏllin\textit{n} upp a\textit{t} hỏnd\textit{om}.  
\edtext{Helgi}{\Afootnote{Hellga hann  L679, L1572.}} m\textit{ælir} nú er  auþna  mín  úti  \edtext{o\textit{c} enduþ}{\Afootnote{÷  L2404.}}, a\textit{t} e\textit{c} misti þín. Helga v\textit{ar}þ  \edtext{ófljótt}{\Afootnote{eij fljott  L679.}} a\textit{t} fá  los\textit{at} sv\textit{er}þ\textit{it} or
j\textit{or}þun\textit{n}i,  \edtext{e\textit{n}}{\Afootnote{÷  L1572.}} þ\textit{at} sá Hróm\textit{undr} o\textit{c} brá  \edtext{Mistilteini}{\Afootnote{Mistiltein  L1572.}} í  lopt  \edtext{upp}{\Afootnote{+ og  L679.}} á Háls Helga, sva í eino aftók hỏfuþ\textit{it},  féll sva  \edtext{Helgi}{\Afootnote{+ hinn fræknasti  L679.}},  \edtext{ein\textit{n} hin\textit{n}}{\Afootnote{og  L679.}} ágjætasti  \edtext{o\textit{c} fræknasti}{\Afootnote{÷  L679.}} m\textit{aþr}, er  ver\textit{it} hefr, o\textit{c} er h\textit{an}s víþa gét\textit{it} í \edtext{ỏþr\textit{om}}{\Afootnote{÷  L2404.}} sỏg\textit{om}, \edtext{þókti nú Sví\textit{om} st\textit{or} miss\textit{ir} a\textit{t} h\textit{onom} v\textit{er}þa}{\Afootnote{÷  L1572.}}. 
J viþreign þ\textit{eir}ra Hróm\textit{undar} oc  Helga, 
hafþi Hr\textit{ómundr} eitt sin\textit{n} hỏggv\textit{it} 
\edtext{af}{\Afootnote{á  L679.}} hjálmb\textit{ar}þ\textit{it} Helga, 
o\textit{c} \edtext{hỏfþo}{\Afootnote{+ þá  L679.}} \edtext{brotn\textit{at}}{\Afootnote{+ þá  L2404.}} m\textit{ỏr}g sk\textit{ỏr}þ  
\edtext{þá}{\Afootnote{÷ L679, L2404, L1572.}} í eggina á Mistilteini. 
Ept\textit{ir} dauþa Helga,
t\textit{óc} Hróm\textit{undr} \edtext{a\textit{t} hỏggva m\textit{eþ}}{\Afootnote{÷  L2404.}} \edtext{ákafl\textit{e}gri}{\Afootnote{kallmannlegri  L679, ákaflega  L2404.}} \edtext{fram\textit{m}gaungo}{\Afootnote{at beriaz  L2404.}} þá Svía; suma klauf h\textit{ann} at  endilaungo; af
sum\textit{om} hjó h\textit{ann} hendr  \edtext{eþa}{\Afootnote{og  L679, L1572.}} \edtext{fætr}{\Afootnote{+ oc af  L2404.}}, sum\textit{om} hỏfuþ, suma sundr í miþjo; h\textit{ann} helsló bæþi m\textit{enn}
o\textit{c} hesta,  o\textit{c} valkỏsturin\textit{n} \edtext{v\textit{ar}þ}{\Afootnote{var  L679.}} h\textit{onom} í miþti, 
\edtext{o\textit{c} báþ\textit{ar}}{\Afootnote{÷  L1572.}} \edtext{sín\textit{ar}}{\Afootnote{÷  L2404, L1572.}} \edtext{alblóþugar}{\Afootnote{÷  L679, L2404, L1572.}} \edtext{hendr}{\Afootnote{÷  L1572.}} \edtext{b\textit{ar}}{\Afootnote{hafþi  L2404, ÷  L1572.}} \edtext{h\textit{ann}}{\Afootnote{+ bloduga  L679, + blóþugar  L2404, ÷  L1572.}} \edtext{a\textit{t} ỏxl\textit{om}}{\Afootnote{til axla  L679, L2404, ÷  L1572.}}, gé\textit{ck} \edtext{sva}{\Afootnote{÷  L1572.}} þ\textit{ess}i  \edtext{h\textit{ar}þa}{\Afootnote{÷  L679, L2404, L1572.}} \edtext{orrusta}{\Afootnote{+ allt  L1572.}} t\textit{il} nætor.  Ol\textit{afr} K\textit{óngr} hafþi
b\textit{ar}ist m\textit{eþ} \edtext{betsta}{\Afootnote{mesta  L679.}} drengskap, o\textit{c} allt liþ  h\textit{an}s u\textit{m}
dagin\textit{n},  \edtext{þá}{\Afootnote{þó  L2404, L1572.}} v\textit{ar} m\textit{ar}gt af þ\textit{ví} fall\textit{it},  \edtext{lauk sva degin\textit{om}}{\Afootnote{÷  L1572.}},  \edtext{e\textit{n}}{\Afootnote{ad  L679, L2404, ÷  L1572.}} 
\edtext{dym\textit{m}a}{\Afootnote{÷  L1572.}} \edtext{næturin\textit{n}\textit{ar}}{\Afootnote{÷  L679, L1572.}} \edtext{t\textit{óc}}{\Afootnote{÷  L1572.}} \edtext{v\textit{it}}{\Afootnote{÷  L679, L1572.}}, 
sn\textit{er}o  \edtext{þá}{\Afootnote{nú  L679, L1572.}} all\textit{ir} 
t\textit{il} h\textit{er}búþa  \edtext{sin\textit{n}a}{\Afootnote{÷  L679.}}, oc  tóco náþ\textit{ir},  \edtext{e\textit{n}}{\Afootnote{og  L1572.}} \edtext{þ\textit{ei}r s\textit{ær}þo}{\Afootnote{÷  L1572.}} léto binda  \edtext{u\textit{m}}{\Afootnote{÷  L1572.}} sár sín. Hróm\textit{undr} snýr þá t\textit{il} tjalds síns yf\textit{ir} u\textit{m}
vatn\textit{it}; h\textit{ann} g\textit{en}gr á Jsin\textit{n} oc  ætl\textit{ar} heim, sjer  \edtext{h\textit{ann}}{\Afootnote{+ þá  L679.}} \edtext{m\textit{an}\textit{n} ein\textit{n}}{\Afootnote{eirn mann  L679.}} st\textit{an}da á Jsn\textit{om}, o\textit{c} er a\textit{t} rista  rún\textit{ir} á  \edtext{ísin\textit{n}}{\Afootnote{hann  L679.}}, sva  \edtext{a\textit{t}}{\Afootnote{÷  L679.}} allr v\textit{ar} ísin\textit{n} s\textit{em} ein skrifuþ bók;
þ\textit{at} sá  e\textit{ck}i Hróm\textit{undr} a\textit{t}
rún\textit{ir}n\textit{ar} \edtext{v\textit{or}o}{\Afootnote{urdu  L679.}} f\textit{yrir} fót\textit{om} \edtext{h\textit{an}s}{\Afootnote{+ enn  L679.}}; h\textit{ann} þykist  \edtext{kén\textit{n}a þ\textit{ar}}{\Afootnote{þar kienna  L679, þar  L2404.}} Vola kompán  \edtext{sin\textit{n}}{\Afootnote{+ kenna  L2404.}}, m\textit{eþ} \edtext{stál sprota}{\Afootnote{stafbroti  L1572.}} í hendi sér, m\textit{eþ} \edtext{hv\textit{or}j\textit{om}}{\Afootnote{hvorju  L1572.}} h\textit{ann} risti rún\textit{ir}n\textit{ar}. Hróm\textit{undr} leypr
þeg\textit{ar} a\textit{t} Vola, e\textit{n} han\textit{n} blæs sva h\textit{ar}þt a\textit{t} Hróm\textit{undi} a\textit{t}
Mistiltein\textit{n} hraut  \edtext{l\textit{an}g\textit{an} veg}{\Afootnote{langt  L679, ÷  L2404, langa leid  L1572.}} or hendi  \edtext{h\textit{an}s}{\Afootnote{hönum langan veg  L2404.}}, o\textit{c} þ\textit{at} í lopti, sva lengi saung í  \edtext{því}{\Afootnote{honum  L2404.}}. ||\ledouternote{123v} \edtext{Þat b\textit{ar} t\textit{il}, a\textit{t}}{\Afootnote{enn  L679.}} þ\textit{ar} s\textit{em} þ\textit{at} q\textit{vo}m niþr,  \edtext{v\textit{ar}þ}{\Afootnote{var  L679.}} f\textit{yrir} þ\textit{ví} \edtext{st\textit{or}}{\Afootnote{÷  L679.}}  rifo   sprúnga,  á
Jsn\textit{om}, þ\textit{at} hitti  \edtext{á rifuna}{\Afootnote{sprunguna  L679.}}, o\textit{c} féll  \edtext{sva}{\Afootnote{÷  L679.}} t\textit{il} \edtext{botns}{\Afootnote{+ enn  L679.}}; vatn\textit{it} \edtext{v\textit{ar}}{\Afootnote{er  L679.}} lx  \edtext{faþma}{\Afootnote{feta  L1572.}} djúpt; Vole hló o\textit{c} m\textit{ælir}: nú erto   feigr   \edtext{þú hinn}{\Afootnote{þinn  L679, L2404, L1572.}} f\textit{an}tr  o\textit{c}
qven\textit{n}a  skratti! a\textit{t} þú mist\textit{ir} sv\textit{er}þ\textit{it} or hendi þér.  Ovíst er þ\textit{at}, s\textit{egir} Hróm\textit{undr} Vole vill þá
hỏggva t\textit{il} h\textit{an}s m\textit{eþ} sv\textit{er}þi,  
e\textit{n} Hróm\textit{undr} ljóp a\textit{t} 
\edtext{Vola}{\Afootnote{hónum L679.}}, 
o\textit{c} náþi hỏnd\textit{om} u\textit{m} háls  \edtext{á Vola}{\Afootnote{hans  L679, L1572.}},  e\textit{n} h\textit{ann} t\textit{óc} á  \edtext{móti}{\Afootnote{+ og  L679.}}, urþo  \edtext{þá}{\Afootnote{÷  L679.}} e\textit{ck}i l\textit{an}g\textit{ar} sviptíng\textit{ar}, þ\textit{ar}
t\textit{il} \edtext{a\textit{t}}{\Afootnote{÷  L679.}} Hróm\textit{undr} f\textit{ær}þi Vola  \edtext{í}{\Afootnote{a  L679.}} lopt upp,  \edtext{o\textit{c}}{\Afootnote{+ færþi  L2404.}} síþ\textit{an} \edtext{niþr flat\textit{an}\textit{n}}{\Afootnote{flatann nidr  L679.}} á Jsin\textit{n}. Þá  m\textit{ælir} Hróm\textit{undr} Nú
lítst mér a\textit{t} þú sért feigr engu síþr  e\textit{n}
e\textit{c}, o\textit{c} skal þér nú a\textit{t} gagni v\textit{er}þa
l\textit{an}gr rógr o\textit{c} \edtext{álýgi}{\Afootnote{lijgi  L679.}},  líka  \edtext{hefi}{\Afootnote{hef  L679, L2404.}} e\textit{c} \edtext{lengi}{\Afootnote{÷  L679.}} hugsat þér kaup f\textit{yrir} Hrók
min\textit{n}, s\textit{e}m  þú myrdt\textit{ir}
f\textit{or}þ\textit{om} saklausan\textit{n} af tómri ỏfund v\textit{it}
m\textit{ic}. 
Vole  t\textit{óc} þá a\textit{t} blása  \edtext{veþri}{\Afootnote{vedr  L1572.}} \edtext{á Hróm\textit{und}}{\Afootnote{ad Hromundi  L679.}} e\textit{n} Hróm\textit{undr} h\textit{er}þti þá  \edtext{a\textit{t}}{\Afootnote{÷  L2404.}} hálsi  h\textit{an}s m\textit{eþ} greipun\textit{om},
þ\textit{ar} t\textit{il} \edtext{a\textit{t}}{\Afootnote{÷  L679.}} h\textit{ann} hengdi Vola m\textit{eþ} háls kyrkíngo, o\textit{c} lauk  \edtext{sva Vole æfi sin\textit{n}i}{\Afootnote{hann svo sijnu lijfi  L679.}} m\textit{eþ} \edtext{aum\textit{om}}{\Afootnote{illom  L2404, L1572.}} dauþdaga.  Ept\textit{ir} þetta settist Hróm\textit{undr}
niþr á Jsin\textit{n}, styrþr, sár o\textit{c} móþr, e\textit{n} myrkr o\textit{c} frosth\textit{ar}ka
v\textit{ar} m\textit{ic}il, þóktist h\textit{ann} nú ílla 
haldin\textit{n}; Vatn\textit{it} v\textit{ar} rifn\textit{at} u\textit{m}
þv\textit{er}t, e\textit{n} qviþr h\textit{an}s ristr,  \edtext{o\textit{c}}{\Afootnote{enn  L679, ÷  L1572.}} \edtext{sjálfor sár  ok  móþr}{\Afootnote{÷  L1572.}},  e\textit{n} þ\textit{at} \edtext{v\textit{er}sta þókti h\textit{onom}}{\Afootnote{þotti honum verst  L679.}}, a\textit{t} Mistiltein\textit{n} v\textit{ar} \edtext{h\textit{onom}}{\Afootnote{÷  L679.}} tapaþr; t\textit{óc} h\textit{ann} þá a\textit{t} telja sér tỏlr,
o\textit{c} s\textit{egir} v\textit{it} sjálf\textit{an} s\textit{ic}: nú
hefi e\textit{c} feng\textit{it} m\textit{ỏr}g sár,
f\textit{yrir} orþ Svnah\textit{vítar} e\textit{c} \edtext{hefi}{\Afootnote{+ og  L679.}} mist mína  viij b\textit{ræþ}r, o\textit{c} mitt góþa
sv\textit{er}þ fæ e\textit{c} aldrei aptr  \edtext{á min\textit{n}i  æfi}{\Afootnote{á æfi minni  L679, ÷  L1572.}}, o\textit{c} nú  \edtext{sýnist}{\Afootnote{+ mier  L679, L2404, L1572.}} vísastr  \edtext{dauþin\textit{n}}{\Afootnote{daudi fijrstr minna lijfsdaga  L679, daudi  L2404, L1572.}}.  \edtext{Hróm\textit{undr} leggst}{\Afootnote{legst hann  L679, L2404, L1572.}} \edtext{nú}{\Afootnote{þá  L679.}} \edtext{f\textit{yrir}}{\Afootnote{nidr  L1572.}} u\textit{m} stund, o\textit{c} t\textit{ec}or h\textit{ann} þá
t\textit{il} a\textit{t} styrþna o\textit{c} kólna, skreiþist  \edtext{h\textit{ann} þá}{\Afootnote{þá Hromundr  L679, L1572, Romundr þa  L2404.}} á  \edtext{fót}{\Afootnote{fætr  L679, L2404.}} o\textit{c} villdi eitthvaþ t\textit{il} ráþs t\textit{ac}a,  \edtext{v\textit{ar}þ}{\Afootnote{verdr  L679.}} \edtext{h\textit{onom}}{\Afootnote{+ þar  L2404.}} þ\textit{at} fyrst, a\textit{t} h\textit{ann} \edtext{hratt}{\Afootnote{hrindir  L679.}} \edtext{in\textit{n}}{\Afootnote{÷  L1572.}} Jstrun\textit{n}i í qviþin\textit{n}, o\textit{c} tók  linda
kníf sin\textit{n} o\textit{c} stakk f\textit{yrir} m\textit{eþ}
oþdin\textit{om}, h\textit{ann} rakti  g\textit{ar}n  or 
armb\textit{an}dino o\textit{c} sauma\textit{þ}i m\textit{eþ}
þ\textit{ví}, síþ\textit{an} vafþi h\textit{ann} a\textit{t} sér  \edtext{klæþ\textit{om}}{\Afootnote{klæþinn  L2404.}}, o\textit{c} leggst f\textit{yrir} aptr, o\textit{c} v\textit{ar} nú
n\textit{ær}sta v\textit{it} \edtext{hnoll}{\Afootnote{ hroll   L679, L1572.}},  \edtext{þó}{\Afootnote{þá  L679, L2404, L1572.}} \edtext{var}{\Afootnote{÷  L2404.}} ||\ledouternote{124r} \edtext{bj\textit{ar}t heiþi}{\Afootnote{heiþbiart  L2404.}} \edtext{o\textit{c}}{\Afootnote{÷  L679.}} \edtext{frostit}{\Afootnote{frost  L679, micit frost  L2404.}},  \edtext{s\textit{em}}{\Afootnote{og  L679, sva vatnit  L2404, ÷  L1572.}} \edtext{t\textit{óc}}{\Afootnote{+ nu til  L679.}} a\textit{t} leggja  \edtext{Vatn\textit{it}}{\Afootnote{÷  L2404.}}, lá Hróm\textit{undr} \edtext{sva}{\Afootnote{þá  L2404.}} u\textit{m} stund, a\textit{t} e\textit{ck}ert v\textit{ar}þ t\textit{il}
a\textit{d}stoþ\textit{ar}.  
\pend 
\endnumbering  
\beginnumbering 
\pstart  
\vspace{5mm}\subsection*{XV kap\textit{ítuli}} 
Litlo ept\textit{ir} þetta, 
v\textit{er}þr Hróm\textit{undr} v\textit{ar} v\textit{it} \edtext{þ\textit{at}}{\Afootnote{÷  L679.}} a\textit{t} tek\textit{it} er á  \edtext{h\textit{onom}}{\Afootnote{fotum hans  L679.}},  
o\textit{c} \edtext{sp\textit{urþ} a\textit{t}}{\Afootnote{sagt  L679.}}:  \edtext{gétr}{\Afootnote{tekr  L2404.}} þú tal\textit{at} v\textit{it} m\textit{ic} \edtext{Hróm\textit{undr}}{\Afootnote{÷  L679.}} eþr e\textit{ck}i? Hróm\textit{undr} m\textit{ælir} v\textit{ar}la er þ\textit{at} \edtext{sva}{\Afootnote{÷  L1572.}} \edtext{af}{\Afootnote{÷  L679, L1572.}} \edtext{megi segja}{\Afootnote{÷  L1572.}}, eþa hv\textit{or} er sá  \edtext{er}{\Afootnote{sem  L1572.}} leit\textit{ar} máls  \edtext{af mér}{\Afootnote{vit mic  L2404.}}?  \edtext{h\textit{onom} v\textit{ar} s\textit{agt}}{\Afootnote{÷  L679.}} Hagall er nafn mitt, o\textit{c} á e\textit{c} eyrindi a\textit{t} fin\textit{n}a þ\textit{ic}, o\textit{c}
b\textit{er} þ\textit{ic} a\textit{t} setjast upp o\textit{c} tala
v\textit{it} \edtext{m\textit{ic}}{\Afootnote{+ ef gietr  L679.}},  þ\textit{ví} e\textit{ck}i má  tefja. 
Hróm\textit{undr} b\textit{ar} s\textit{ic} þá a\textit{t}
setjast upp, o\textit{c} leit  í kríng\textit{om} s\textit{ic},  \edtext{sá h\textit{ann}}{\Afootnote{hann sá  L679, L2404.}} \edtext{þá}{\Afootnote{÷  L2404.}} ein\textit{n} gaml\textit{an} m\textit{ann} o\textit{c} ij Jómfrúr, kéndi h\textit{ann} \edtext{þ\textit{ar}}{\Afootnote{þær  L1572.}} systr Ol\textit{afs} K\textit{óngs} Svanh\textit{vít} o\textit{c} \edtext{Dagnýo}{\Afootnote{Dagni  L679.}}, v\textit{ar} þ\textit{an}g\textit{at} q\textit{vo}me\textit{n} vagn o\textit{c} \edtext{hest\textit{ar}}{\Afootnote{hestr  L1572.}}; þ\textit{ær} qveþja Hróm\textit{und} e\textit{n} h\textit{ann} gleþst
v\textit{it} q\textit{vo}mo  þ\textit{eir}ra, þ\textit{ær} biþja
Hagal a\textit{t} \edtext{t\textit{ac}a}{\Afootnote{færa  L2404.}} Hróm\textit{und} upp á vagnin\textit{n}, e\textit{n}
h\textit{ann} \edtext{segist t\textit{il} reyna a\textit{t} q\textit{vo}mast
	sjálfr}{\Afootnote{letz mundi sialfr til reyna  L2404.}}, vildi þá  Hróm\textit{undr} uppst\textit{an}da,  \edtext{o\textit{c}}{\Afootnote{en  L1572.}} gat e\textit{ck}i. Þá m\textit{ælir} Hróm\textit{undr} k\textit{ar}li  hv\textit{or}j\textit{om} \edtext{kémr}{\Afootnote{+ ad  L679, L2404, L1572.}} ỏrv\textit{er}pi, o\textit{c} sama hend\textit{ir}
m\textit{ic} nú. Þá f\textit{or}o  þau þrjú t\textit{il}, Hagall
o\textit{c} Jómfrún\textit{n}\textit{ar}, a\textit{t} q\textit{vo}ma  \edtext{Hróm\textit{undi}}{\Afootnote{+ upp  L679.}} í Vagnin\textit{n} o\textit{c} gáto  \edtext{v\textit{ar}la}{\Afootnote{þad naumast  L679.}}, f\textit{yrir} þýngsl\textit{om} h\textit{an}s; síþ\textit{an} óko þau
h\textit{onom} t\textit{il} skóg\textit{ar}, í eitt fagurt
rjóþr,  \edtext{v\textit{ar}þ þ\textit{ar}}{\Afootnote{þar var  L679.}} f\textit{yrir} þ\textit{ei}m lítill  Bær, o\textit{c}
v\textit{el} \edtext{u\textit{m}vendtr}{\Afootnote{um buid  L679, umvendt  L1572.}}, þ\textit{ar} kom  \edtext{ein}{\Afootnote{÷  L2404.}} kérlíng  \edtext{út}{\Afootnote{÷  L679.}} t\textit{il} þ\textit{eir}ra  o\textit{c} \edtext{hjálpa\textit{þ}i Hróm\textit{undi}}{\Afootnote{hromundi hiálpadi  L1572.}} or  \edtext{vagni}{\Afootnote{vagninu  L679, vagninum  L1572.}} t\textit{il} hvílo, v\textit{ar} \edtext{þá dreg\textit{it} af Hróm\textit{undi}}{\Afootnote{so af honum dregid  L1572.}} 
\edtext{a\textit{t}}{\Afootnote{so  L679.}} \edtext{v\textit{ar}la mátti mæla}{\Afootnote{varla mátti hann mæla L679, L2404, hann mátti valla mæla  L1572.}} f\textit{yrir} kulda sak\textit{ir}, o\textit{c} blóþrás; 
t\textit{óc}o þ\textit{ær} þá a\textit{t} tendra  \edtext{elda}{\Afootnote{elld  L679.}} o\textit{c} t\textit{il} búa heita dry\textit{ck}i a\textit{t} lífga  h\textit{ann} á, líka a\textit{t} þvo h\textit{ann} o\textit{c}
sauma qviþ h\textit{an}s, v\textit{ar} \edtext{þá}{\Afootnote{÷  L679.}} k\textit{ar}lin\textit{n} Hagall hin\textit{n} \edtext{f\textit{or}nemsti o\textit{c}}{\Afootnote{÷  L2404.}} ráþugasti  \edtext{í þ\textit{ví}}{\Afootnote{÷  L679, L2404.}} a\textit{t} lífga Hróm\textit{und} t\textit{óc} h\textit{ann} \edtext{hraþt}{\Afootnote{nú braþr  L2404.}} v\textit{it} a\textit{t} hressast  \edtext{o\textit{c}}{\Afootnote{÷  L2404.}} \edtext{hj\textit{ar}na}{\Afootnote{+ og  L679, ÷  L2404.}}. Þá gé\textit{ck} Svanh\textit{vít} t\textit{il} ||\ledouternote{124v} h\textit{an}s o\textit{c} m\textit{ælir} Nú v\textit{er}þ\textit{om}
v\textit{it} systr, a\textit{t} skilja v\textit{it} þ\textit{ic},
o\textit{c} hefi e\textit{c} feng\textit{it} þér hús o\textit{c} hjúkrún hér á laun;  \edtext{v\textit{it} feng\textit{om}}{\Afootnote{höfom viþ fengit  L2404.}} þen\textit{n}\textit{an}\textit{n} m\textit{ann} a\textit{t} vitja þín, o\textit{c} eins a\textit{t} græþa þ\textit{ic},  \edtext{o\textit{c}}{\Afootnote{÷  L679.}} man k\textit{er}líng h\textit{an}s  \edtext{kun\textit{n}a}{\Afootnote{gieta  L679.}} \edtext{a\textit{t} fræþa}{\Afootnote{frætt  L679.}} \edtext{m\textit{enn}}{\Afootnote{þig  L679.}} á eina o\textit{c} aþra luti, þ\textit{ví} \edtext{a\textit{t}}{\Afootnote{÷  L679.}} hún er f\textit{or}n í brỏgþ\textit{om}. 
Hróm\textit{undr} m\textit{ælir} mest  \edtext{þyk\textit{ir}}{\Afootnote{þiki  L679.}} mér u\textit{m} Mistiltein, a\textit{t} \edtext{ec  misti}{\Afootnote{missa  L679.}} h\textit{ann}.  \edtext{Svanh\textit{vít} m\textit{ælir}}{\Afootnote{÷  L679.}} fleyra v\textit{er}þr a\textit{t} nýta e\textit{n} gott eitt, o\textit{c} k\textit{an}to a\textit{t} fá þ\textit{at} bætt, ef lengi
lif\textit{ir}. Hróm\textit{undr} þakk\textit{ar} hen\textit{n}i  \edtext{f\textit{yrir}}{\Afootnote{÷  L679.}} all\textit{an} sin\textit{n} st\textit{ar}fa, o\textit{c} skiljast þau  \edtext{a\textit{t}}{\Afootnote{÷  L679.}} síþ\textit{an}. Svanh\textit{vít} m\textit{ælir} t\textit{il} Hagals: Nú biþ e\textit{c} \edtext{þ\textit{ic}}{\Afootnote{÷  L679.}} f\textit{yrir} \edtext{Hróm\textit{und}}{\Afootnote{÷  L679.}} a\textit{t} þú o\textit{c} kona þín  \edtext{g\textit{jỏr}it}{\Afootnote{veitit honum  L2404.}} \edtext{þ\textit{at}}{\Afootnote{þid  L679, it  L2404.}} betsta  \edtext{þ\textit{it} gét\textit{it}}{\Afootnote{÷  L679, sem kunnit  L2404.}} \edtext{t\textit{il}}{\Afootnote{÷  L2404.}} \edtext{v\textit{it} Hróm\textit{und}}{\Afootnote{hans  L679, ÷  L2404.}} \edtext{sva lengi}{\Afootnote{÷  L2404.}} \edtext{s\textit{em}}{\Afootnote{÷  L679, L2404.}} \edtext{h\textit{ann} þ\textit{ar}f  þ\textit{ess}}{\Afootnote{÷  L2404.}} \edtext{v\textit{it}}{\Afootnote{med  L679, ÷  L2404.}},  \edtext{e\textit{n} e\textit{c} man sjálf}{\Afootnote{oc man ek  L2404.}} láta y\textit{ck}or vita, a\textit{t} mér  \edtext{betr k\textit{an} a\textit{t}}{\Afootnote{ betr  L679, kann betr  L2404.}} \edtext{þykja}{\Afootnote{þickir  L679.}} \edtext{a\textit{t}}{\Afootnote{ef  L2404.}} h\textit{ann} lif\textit{ir}. Þau  \edtext{qváþo}{\Afootnote{+ göra  L2404.}} slíkt  \edtext{skyldo}{\Afootnote{skillda  L679, ÷  L2404, slikt skildum  L1572.}} \edtext{t\textit{il} láta}{\Afootnote{er mátto  L2404.}},  \edtext{er hefþo}{\Afootnote{ok til hefþo  L2404.}}; f\textit{er}þuþust þ\textit{ær} \edtext{síþ\textit{an}}{\Afootnote{sijstr i  L679.}} \edtext{burt}{\Afootnote{heim  L2404.}},  \edtext{Svanh\textit{vít} o\textit{c} Dagný}{\Afootnote{÷  L679.}} o\textit{c} er ei gét\textit{it} u\textit{m} f\textit{er}þ þ\textit{eir}ra.
En\textit{n} Hróm\textit{undr} t\textit{óc} a\textit{t} hressast, sva  h\textit{ann} komst af re\textit{ck}jo,  \edtext{þeg\textit{ar}}{\Afootnote{þá hann  L679, + hann  L2404, L1572.}} hafþi leg\textit{it} hálf\textit{an} mánuþ,  lækna\textit{þ}i
k\textit{er}líng Hagals  \edtext{h\textit{ann}}{\Afootnote{hans  L679.}} \edtext{⸌s\textit{em} h\textit{ie}t⸍ Haddýs}{\Afootnote{sem Haddijs hiet  L679.}}, o\textit{c} veitti h\textit{onom} betsta  \edtext{umbúþ}{\Afootnote{um bun  L679.}} \edtext{a\textit{t}}{\Afootnote{af  L679, L1572.}} kosti o\textit{c} klæþ\textit{om}, o\textit{c} ỏllo þ\textit{ví}
s\textit{em} \edtext{v\textit{it}}{\Afootnote{hann  L679, med  L1572.}} þurfti,  Einosin\textit{n}i q\textit{vo}m hún a\textit{t}
máli v\textit{it} Hróm\textit{und} o\textit{c} m\textit{ælir} hv\textit{or}nig k\textit{an}to v\textit{it} þ\textit{ic} hér í kotino?
V\textit{el} s\textit{egir} h\textit{ann}, o\textit{c} \edtext{einkis}{\Afootnote{eckert  L679.}} þykir  mér ábóta v\textit{an}t, nema m\textit{ic}
v\textit{an}t\textit{ar} Mistiltein. 
E\textit{ck}i  er
h\textit{ann} a\textit{t} s\textit{or}ga, 
s\textit{egir} Haddýs,
þ\textit{ví} h\textit{ann} er í þ\textit{ei}m st\textit{at}, s\textit{em}
engin\textit{n} hefr hỏnd á. 
Hróm\textit{undr}  \edtext{m\textit{ælir}}{\Afootnote{segir L679.}}  \edtext{hv\textit{aþ}}{\Afootnote{÷  L679.}} s\textit{egir} þú mér 
af  Ol\textit{afi} K\textit{óngi}
o\textit{c} \edtext{sigrsæld}{\Afootnote{sigrsældom  L2404.}} h\textit{an}s? 
Hún m\textit{ælir} þ\textit{at} hygg e\textit{c}, 
at  \edtext{an\textit{n}\textit{at}hv\textit{or}t séo þ\textit{ei}r}{\Afootnote{þeir sieu annad hvórt  L679.}} sætt\textit{ir} Hálfdán o\textit{c} h\textit{ann}, eþa þ\textit{ei}r  \edtext{halda}{\Afootnote{halldi  L679.}} uppi b\textit{ar}dag\textit{an}um t\textit{il} an\textit{n}\textit{ar}s
árs, man þ\textit{at} þ\textit{ví} valda a\textit{t} 
\edtext{Hel\-gi}{\Afootnote{+landvarnar madr L679.}} er fall\-in\textit{n} 
\edtext{Land\-v\textit{ar}n\textit{ar}\-ma\-þr}{\Afootnote{÷  L679}} 
||\ledouternote{125r} \edtext{Hálfdán\textit{ar} K\textit{óngs}}{\Afootnote{÷  L679, Olaf Kongs  L2404, hans  L1572.}}. 
Ekki lífn\textit{ar} h\textit{ann} aptr, s\textit{egir} Hróm\textit{undr}. 
Hún  \edtext{m\textit{ælir}}{\Afootnote{+ sá  L679, L2404.}} m\textit{aþr} er  \edtext{sá}{\Afootnote{÷  L679, L2404.}} m\textit{eþ} \edtext{Halfdáni}{\Afootnote{+ kongi  L679, kongi  L2404.}}, er Blindr heit\textit{ir}, h\textit{ann} er fárra  líki,
en\textit{n} fleyrstra jafni, a\textit{t} þ\textit{ví} s\textit{em} íllt  \edtext{er}{\Afootnote{+ þvj  L679.}}, h\textit{ann} er bæþi  gỏldróttr  o\textit{c} falskr, h\textit{ann} er  \edtext{heimugl\textit{e}gt}{\Afootnote{÷  L2404.}} \edtext{Ráþaneyti}{\Afootnote{ráþgjafi  L2404.}} \edtext{Hálfdánar K\textit{óngs}}{\Afootnote{kongs Hálfdanar  L679.}} hefr  \edtext{K\textit{óngr}}{\Afootnote{÷  L679.}} beþ\textit{it} h\textit{ann} \edtext{a\textit{t} leita upp þan\textit{n} m\textit{an}\textit{n} s\textit{em} feldi  Helga}{\Afootnote{÷  L2404.}}, o\textit{c} \edtext{heit\textit{ir}}{\Afootnote{heitid  L679, L2404.}} \edtext{h\textit{onom} hin\textit{om}}{\Afootnote{÷  L2404.}} st\textit{ær}sto  \edtext{gjỏf\textit{om}}{\Afootnote{launum  L679, L2404, L1572.}}, liggr h\textit{ann} úti  \edtext{u\textit{m} þ\textit{at}}{\Afootnote{÷  L679.}} m\textit{eþ} f\textit{or}neskjoskap, a\textit{t} frétta  \edtext{h\textit{ann} upp}{\Afootnote{÷  L679.}} hv\textit{ar} h\textit{ann} muni niþor  \edtext{q\textit{vo}min\textit{n}}{\Afootnote{+ vera  L679.}},  \edtext{o\textit{c}}{\Afootnote{÷  L1572.}} man h\textit{ann} v\textit{er}þa þ\textit{ess} \edtext{skjótt}{\Afootnote{fliótt  L1572.}} vís, þ\textit{ví} \edtext{a\textit{t}}{\Afootnote{÷  L679, L2404.}} h\textit{ann} er bæþi  slægr  o\textit{c} vitr. 
Þat b\textit{ar} t\textit{il} opt á þ\textit{ei}m tíma s\textit{em} \edtext{a\textit{t}}{\Afootnote{÷  L679, L2404.}} Hróm\textit{undr} lá í græþslo hjá Haddýsi, a\textit{t} Hagall f\textit{or}
t\textit{il} \edtext{veiþa}{\Afootnote{veidi  L679.}} o\textit{c} fiski abla; h\textit{ann} gé\textit{ck} á
Jsin\textit{n} \edtext{á}{\Afootnote{÷  L1572.}} \edtext{Væn\textit{ir}}{\Afootnote{+ oc for at veiþa  L2404.}} ept\textit{ir} \edtext{v\textit{an}a}{\Afootnote{+ sijnum  L679.}}, o\textit{c} abla\textit{þ}i  jafn\textit{an} v\textit{el}.
Eitt qvöld q\textit{vo}m h\textit{ann} heim, o\textit{c} dró ept\textit{ir} sér  \edtext{sva}{\Afootnote{÷  L2404.}} st\textit{or}a  Fliþro, \edtext{a\textit{t} h\textit{ann} gat}{\Afootnote{og gat hann  L679.}} e\textit{ck}i b\textit{or}it; k\textit{er}líng fagn\textit{ar}
h\textit{onom} blíþl\textit{e}ga  o\textit{c} qveþr
h\textit{ann} f\textit{ær}a sér t\textit{il} mat\textit{ar}. Hagall  \edtext{m\textit{ælir}}{\Afootnote{qvad  L2404, L1572.}} \edtext{þ\textit{at}}{\Afootnote{+ ætla eg  L679.}} víst v\textit{er}a,  \edtext{a\textit{t} nú}{\Afootnote{oc  L1572.}} \edtext{meina e\textit{c} a\textit{t}}{\Afootnote{÷  L2404.}} \edtext{e\textit{c} f\textit{ær}i}{\Afootnote{færi ek  L2404.}} þér hlut\textit{ar}bót, o\textit{c} \edtext{þarf e\textit{ck}i a\textit{t} tvíla a\textit{t}}{\Afootnote{÷  L2404, L1572.}}  \edtext{þ\textit{at} er sú st\textit{ær}sta sképna, s\textit{em} e\textit{c} hefi}{\Afootnote{hef ek alldrei stęrri scepno  L2404.}} \edtext{no\textit{ck}orn tíma áþr}{\Afootnote{÷  L679, L2404, L1572.}} \edtext{feng\textit{it} or vatni a\textit{t} draga}{\Afootnote{or Vatni dreigit  L2404, L1572.}}. Þat  \edtext{er}{\Afootnote{+ og  L679.}} hv\textit{or}ttveggja  \edtext{s\textit{egir} h\textit{on}}{\Afootnote{÷  L679.}},  a\textit{t} \edtext{fleyri ero nú}{\Afootnote{þad eru fleiri til  L679.}} \edtext{a\textit{t} eyþa}{\Afootnote{veidi menn  L1572.}}, e\textit{n} v\textit{an}i er  \edtext{t\textit{il}}{\Afootnote{segir hun  L679.}}, enda kémr nú í  \edtext{g\textit{ar}þ}{\Afootnote{+ minn  L679.}} m\textit{ei}ri veiþi e\textit{n} áþr, o\textit{c} skal  \edtext{e\textit{c}}{\Afootnote{nu  L679, L1572.}} f\textit{ar}a a\textit{t} \edtext{kokka}{\Afootnote{matreida  L1572.}}; t\textit{óc} þá  Hag\textit{all} sax sitt o\textit{c}
sundraþi  \edtext{m\textit{eþ} þ\textit{ví}}{\Afootnote{÷  L679.}} fliþruna, e\textit{n} s\textit{em} han\textit{n} risti  \edtext{h\textit{an}a}{\Afootnote{÷  L679.}} á qviþin\textit{n}, v\textit{ar}þ f\textit{yrir} h\textit{onom} \edtext{sv\textit{er}þ\textit{it}}{\Afootnote{sverþ  L2404.}} í mag\textit{an}um, þ\textit{at} s\textit{em} Hróm\textit{undr} \edtext{mist hafþi}{\Afootnote{hafdi mist  L679.}}; 
þá hló Haddýs o\textit{c} m\textit{ælir} a\textit{t} þ\textit{ví} má  gagn v\textit{er}þa s\textit{em} á l\textit{an}d kémr, o\textit{c}
greip sv\textit{er}þ\textit{it} o\textit{c} b\textit{ar} \edtext{þ\textit{at}}{\Afootnote{÷  L679, L2404.}} t\textit{il} Hróm\textit{undar} o\textit{c} m\textit{ælir} kén\textit{n}\textit{ir} þú  \edtext{no\textit{ck}ot}{\Afootnote{þetta  L679.}} eþr ei? Já s\textit{egir} Hróm\textit{undr},  \edtext{e\textit{c} kén\textit{n}i v\textit{el}}{\Afootnote{÷  L679.}}, o\textit{c} ætla\textit{þ}i e\textit{c} \edtext{þ\textit{at} e\textit{ck}i}{\Afootnote{÷  L679.}} \edtext{a\textit{t} e\textit{c}}{\Afootnote{÷  L679, at  L2404.}} \edtext{m\textit{un}di}{\Afootnote{+ eij  L679.}} kén\textit{n}a héþ\textit{an} í frá. Nú er s\textit{em} q\textit{vo}m\textit{it} er s\textit{egir} \edtext{k\textit{er}líng}{\Afootnote{hun  L679.}},  \edtext{o\textit{c}}{\Afootnote{÷  L679.}} t\textit{ac} v\textit{it} eign þin\textit{n}i. Hróm\textit{undr} brosti
o\textit{c} kysti á hjalt\textit{it};  \edtext{þá t\textit{óc} h\textit{ann} v\textit{it} sv\textit{er}þino}{\Afootnote{÷  L679, þá hann tók vit sverþino  L2404, tók vid sverdinu  L1572.}} o\textit{c} m\textit{ælir} \edtext{njóti}{\Afootnote{man  L2404.}} \edtext{Hag\textit{all} ||\ledouternote{125v} betst}{\Afootnote{best Hagall  L2404.}} h\textit{an}da sin\textit{n}a, e\textit{n} þér man þa\textit{ck}a
v\textit{er}þugast v\textit{er}a,  \edtext{veit e\textit{c} nú a\textit{t} e\textit{c} er í góþra m\textit{ann}a hús\textit{om}, n\textit{ær} s\textit{em}
	e\textit{c} gét ásan\textit{n}ast lát\textit{it}}{\Afootnote{÷  L2404.}}, feldo þau sva tal\textit{it},
a\textit{t} k\textit{er}líng lét upp  a\textit{t} sjóþa. 
\pend \endnumbering 

\beginnumbering \pstart
\vspace{5mm} 
\subsection*{XVI kap\textit{ítuli}} 
No\textit{ck}ro ept\textit{ir} þetta,  
\edtext{þeg\textit{ar}}{\Afootnote{þá  L679.}} Hróm\textit{undr} t\textit{óc} a\textit{t} v\textit{er}þa   gróinn   sára sin\textit{n}a, skéþi  \edtext{þ\textit{at}}{\Afootnote{+ ad  L679, L2404, L1572.}}: 
k\textit{er}líng t\textit{óc} t\textit{il} a\textit{t} fá  \edtext{sva}{\Afootnote{÷  L2404, L1572.}} \edtext{stóra}{\Afootnote{÷  L1572.}} geispa, 
a\textit{t} hún q\textit{vo}m  \edtext{v\textit{ar}la fram\textit{m}}{\Afootnote{framm valla einú  L679.}} orþi f\textit{yrir} þ\textit{ei}m; 
hún m\textit{ælir} einhvỏrstaþ\textit{ar} er e\textit{c} \edtext{nú}{\Afootnote{þar  L679.}} á  \edtext{orþi}{\Afootnote{+ sem eg er eij á bordi  L679.}}, 
o\textit{c} man þ\textit{at} san\textit{n}ast v\textit{er}a, a\textit{t} slíkt megi henda sic;  \edtext{man nú}{\Afootnote{nú man  L2404.}} Hálfdán  \edtext{K\textit{óngr}}{\Afootnote{÷  L679.}} v\textit{er}a  \edtext{á tali}{\Afootnote{ad tala  L679.}} v\textit{it} Blind vin sin\textit{n} ept\textit{ir} venjo. Nú v\textit{er}þr frá þ\textit{ví} a\textit{t} s\textit{egja} a\textit{t} \edtext{þetta sama qvỏld}{\Afootnote{÷  L679.}} \edtext{q\textit{vo}m Bl\textit{indr}}{\Afootnote{Blindr kom  L679.}} \edtext{á tal}{\Afootnote{ad mali  L679.}} v\textit{it} Hálfd\textit{án} K\textit{óng} o\textit{c} m\textit{ælir} frétt\textit{ir} h\textit{afa} fræþt m\textit{ic} u\textit{m}
þ\textit{at}, hv\textit{aþ} heit\textit{it} hafi sá 
m\textit{aþr},  \edtext{er}{\Afootnote{sem  L679.}} vo  \edtext{Helga}{\Afootnote{+ in  L2404.}} Frækna. K\textit{óngr} m\textit{ælir} hv\textit{aþ}
h\textit{ei}t\textit{ir} h\textit{ann}? Bl\textit{indr} s\textit{egir}
Hróm\textit{undr} h\textit{ei}t\textit{ir} h\textit{ann}, 
ættaþr or Noregi. Veitsto hv\textit{ar} h\textit{ann} er nú? s\textit{egir}
K\textit{óngr} Han\textit{n} er  \edtext{e\textit{n}}{\Afootnote{nú  L679.}} á lífi, s\textit{egir} Bl\textit{indr} hjá Hag\textit{ali}
k\textit{ar}li, hér á skógn\textit{om}. Þá  skalto
fin\textit{n}a h\textit{ann}, o\textit{c} láta h\textit{ann} ei und\textit{an}
sleppa, s\textit{egir} K\textit{óngr} oc  manto m\textit{enn}
t\textit{il} kalla,  \edtext{a\textit{t} leita  h\textit{an}s á m\textit{or}gun}{\Afootnote{snema á morgun ad þú skallt finna hann  L679.}}. 
Sva skal  g\textit{jỏr}t
\edtext{v\textit{er}þa}{\Afootnote{vera  L679.}} s\textit{egir} Bl\textit{indr} e\textit{n} u\textit{m}
m\textit{or}gunin\textit{n} ept\textit{ir} v\textit{ar} Bl\textit{indr}
snem\textit{m}a  á fót\textit{om}, o\textit{c} heimt\textit{ar}
xxx m\textit{anna} \edtext{t\textit{il} sín}{\Afootnote{med ser  L1572.}}, o\textit{c} heldr t\textit{il} skógar;  þ\textit{ei}r
q\textit{vo}ma þeg\textit{ar} fyrst byrti. Bl\textit{indr} s\textit{egir} \edtext{vilja}{\Afootnote{÷  L679.}} r\textit{an}sókn uppi h\textit{afa}, o\textit{c} leita
ept\textit{ir} \edtext{þ\textit{ei}m}{\Afootnote{+ manni  L1572.}} s\textit{em} Helga veg\textit{it} \edtext{hafi}{\Afootnote{hafdi  L679.}}, því  Hálfd\textit{án} K\textit{óngr} vilji sva v\textit{er}a
láta. Hag\textit{all} s\textit{egir} heimil  \edtext{skulo}{\Afootnote{÷  L679.}} hús  \edtext{mín}{\Afootnote{sijn  L679.}} hv\textit{or}j\textit{om} m\textit{ann}i, e\textit{n} þó er  \edtext{skyldast}{\Afootnote{skyldt  L2404.}}, a\textit{t} e\textit{c} \edtext{þ\textit{ei}m}{\Afootnote{÷  L679.}} betst leyfi,  \edtext{s\textit{em}}{\Afootnote{er  L679.}} frá Hálfd\textit{áni} K\textit{óngi} q\textit{vo}ma, o\textit{c}
e\textit{ck}i vil e\textit{c} vin\textit{n}a þ\textit{at} f\textit{yrir} eins m\textit{an}s vinátto, a\textit{t} missa vinátto
y\textit{ck}ar  K\textit{ón}gs, o\textit{c} skal yþr
t\textit{il} reiþo  ljós o\textit{c}
hv\textit{aþ} an\textit{n}\textit{at}, er v\textit{it} þurf\textit{it}.  Bl\textit{indr} m\textit{ælir} v\textit{er}þi eins raun á
reyndin\textit{n}i,  \edtext{s\textit{em}}{\Afootnote{og  L679, L1572.}} þú lætr á ||\ledouternote{126r} heyrast, þá er betr tal\textit{at} e\textit{n} e\textit{ck}i.
Hag\textit{all} biþr þá  \edtext{a\textit{t}}{\Afootnote{÷  L679.}} qveikja ljós, o\textit{c} \edtext{lýsa}{\Afootnote{leita  L1572.}} í h\textit{or}n hús\textit{om}, s\textit{em} dagr byrti e\textit{ck}i.
Haddýs  leypr t\textit{il}  óns  a\textit{t}  tendra  ljós, v\textit{ar} þ\textit{ar}
Hróm\textit{undr} f\textit{yrir}, hún b\textit{at} \edtext{h\textit{an}n}{\Afootnote{Hromund  L679, L2404, L1572.}} f\textit{ar}a niþr und\textit{ir} soþkétilin\textit{n}, þ\textit{at}
g\textit{jỏr}þi  \edtext{Hróm\textit{undr}}{\Afootnote{hann  L679.}} e\textit{n} hún hvolfdi  \edtext{katlin\textit{om}}{\Afootnote{+ ofann  L679.}} yf\textit{ir} h\textit{ann} o\textit{c} \edtext{risti}{\Afootnote{ristir  L679.}} no\textit{ck}or f\textit{or}n  \edtext{v\textit{er}s}{\Afootnote{vess  L679, L2404, L1572.}} á  kélilbotnin\textit{n}, gé\textit{ck} síþ\textit{an}
t\textit{il} gést\textit{an}a m\textit{eþ} góþ\textit{an} qven\textit{n}sk\textit{ỏr}úngsskap, 
oc bauþ þ\textit{ei}m leit fram\textit{m}i a\textit{t} hafa; leitoþo þá  \edtext{m\textit{enn} K\textit{óng}s}{\Afootnote{Kongs menn  L679, L1572.}} hv\textit{ar} \edtext{s\textit{em}}{\Afootnote{þeir  L679.}} vildo o\textit{c} fundo e\textit{ck}i Hróm\textit{und} o\textit{c}  e\textit{ck}i  \edtext{heldr fundo þ\textit{ei}r}{\Afootnote{÷  L679.}} \edtext{kélilin\textit{n}}{\Afootnote{kietilinn  L679, L2404, L1572.}}\footnote{
	Note that a better reading is in L679, L2404, and L1572 than in B11109.}  \edtext{f\textit{yrir} sér}{\Afootnote{÷  L679.}}; t\textit{óc} þá Bl\textit{indr} a\textit{t} v\textit{er}þa  brúna þúngr  o\textit{c} áhyggjufullr, snýr h\textit{ann} þá í burto, o\textit{c} er h\textit{ann} \edtext{v\textit{ar}}{\Afootnote{er  L679, ÷  L1572.}} \edtext{q\textit{vo}min\textit{n} áleiþis}{\Afootnote{a leidis kominn  L679, kom á leidis  L1572.}}, m\textit{ælir} h\textit{ann} t\textit{il} m\textit{ann}a sin\textit{n}a: vissol\textit{e}ga er
Hróm\textit{undr} hér á laun; o\textit{c} skul\textit{om} \edtext{v\textit{it}}{\Afootnote{ver  L1572.}} aptr  \edtext{heim}{\Afootnote{÷  L679.}} snúa o\textit{c} leita betr,  
\edtext{g\textit{jỏr}a þ\textit{ei}r}{\Afootnote{þeir gjóra  L679.}} sva  \edtext{a\textit{t} þ\textit{ei}r snúa  aptr}{\Afootnote{÷  L679.}}. 
Þá s\textit{egir} k\textit{er}líng v\textit{it} Hróm\textit{und} nú
skalto f\textit{ar}a í fỏt Eldabusko min\textit{n}\textit{ar},
o\textit{c} skalto hr\textit{ær}a graut t\textit{il} \edtext{matg\textit{jỏr}þ\textit{ar}}{\Afootnote{matar  L1572.}}, fékk  hún h\textit{onom} þá Mistiltein f\textit{yrir}
graut\textit{ar} þv\textit{ỏr}o, 
o\textit{c} m\textit{ælir} m\textit{eþ} þvỏrun\textit{n}i  má v\textit{er}jast ef á liggr.
Litlo síþ\textit{ar} q\textit{vo}mo m\textit{enn} K\textit{ón}gs, 
o\textit{c} sỏgþust vilja betr leita, e\textit{n} fyrr. 
Hag\textit{all} s\textit{egir} þ\textit{ei}m heimila r\textit{an}sỏkun, leita
þ\textit{ei}r  nú lengi, o\textit{c} fin\textit{n}a
\edtext{e\textit{ck}i }{\Afootnote{+ helldr L679.}} Hróm\textit{und} \edtext{a\textit{t} heldr}{\Afootnote{÷  L679.}}; 
sáo þ\textit{ei}r nú  Eldabuskuna, a\textit{t} hún dró til  Möndulin\textit{n}, o\textit{c} leit \edtext{óhýrt}{\Afootnote{hijrlega  L679.}} t\textit{il} \edtext{K\textit{ón}gs m\textit{ann}a}{\Afootnote{þeira  L679.}}, 
snúa þ\textit{ei}r  \edtext{en\textit{n}}{\Afootnote{nu  L679.}} á burt m\textit{eþ} \edtext{stór\textit{om}}{\Afootnote{storann  L679, L1572, miklom  L2404.}} óþo\textit{ck}a, o\textit{c} er þ\textit{ei}r v\textit{or}o á  \edtext{miþja}{\Afootnote{÷  L679.}} leiþ q\textit{vo}mn\textit{ir}, t\textit{ec}or Bl\textit{indr} t\textit{il} orþa  \edtext{o\textit{c} s\textit{egir}}{\Afootnote{÷  L2404.}} m\textit{ic}il skỏm\textit{m} hend\textit{ir} o\textit{ss}, 
a\textit{t} \edtext{v\textit{it}}{\Afootnote{vér  L1572.}} \edtext{gétum e\textit{ck}i fund\textit{it}}{\Afootnote{finnum ei  L1572.}} Hróm\textit{und}
e\textit{n} h\textit{ann} er þó hjá k\textit{er}língo
Hagals, o\textit{c} er hún h\textit{it} meirsta galdra seiþi,
o\textit{c} skul\textit{om} \edtext{v\textit{it}}{\Afootnote{vér  L1572.}} e\textit{n} þá t\textit{il} hætta,  \edtext{o\textit{c} leita}{\Afootnote{÷  L679.}} í þriþja sin\textit{n}; 
snúa  \edtext{þ\textit{ei}r}{\Afootnote{÷  L2404, L1572.}} þá aptr,  
\edtext{o\textit{c} er þ\textit{ei}r v\textit{or}o skam\textit{m}t á leiþ aptr q\textit{vo}mn\textit{ir}}{\Afootnote{þá  L679.}}, 
 t\textit{óc} Haddýs  t\textit{il}
orþa, o\textit{c} m\textit{ælir} t\textit{il} Hróm\textit{undar} \edtext{e\textit{ck}i tjá\textit{ir} a\textit{t} tvíla}{\Afootnote{÷  L1572.}},  \edtext{e\textit{n} þá}{\Afootnote{ad  L679.}} q\textit{vo}ma  \edtext{þ\textit{ei}r}{\Afootnote{÷  L679.}} Bl\textit{indr} o\textit{c} \edtext{m\textit{enn} h\textit{an}s}{\Afootnote{hans menn  L679.}}, o\textit{c} f\textit{ar} þú nú a\textit{t} reka sam\textit{an} kindr
o\textit{ck}ar,  o\textit{c} brúka  vosbúþ\textit{ar}  kubl  \edtext{á þér}{\Afootnote{÷  L679.}}, o\textit{c} fjárpípo v\textit{it} munn ||\ledouternote{126v} þér, o\textit{c} sýng í hen\textit{n}i, e\textit{n} e\textit{c} man
g\textit{an}ga t\textit{il} mjalta. 
Þetta  g\textit{jỏr}ir Hróm\textit{undr} \edtext{h\textit{ann}}{\Afootnote{og  L679.}} g\textit{en}gr t\textit{il} smala m\textit{enn}sko, 
o\textit{c} t\textit{ec}or lúþr meþ  sér, 
o\textit{c} blæs  \edtext{nú}{\Afootnote{÷  L679, L1572.}} sam\textit{an} 
\edtext{hj\textit{or}þina}{\Afootnote{fienu  L679, hjordinni  L2404, L1572.}} s\textit{em} Hag\textit{all} átti; 
o\textit{c} s\textit{em} h\textit{ann}
er q\textit{vo}min\textit{n} á miþja  leiþ,
q\textit{vo}ma  \edtext{þ\textit{ei}r}{\Afootnote{þa  L2404.}} K\textit{óng}s m\textit{enn} o\textit{c} qveþja h\textit{ann}, 
e\textit{n} h\textit{ann} 
\edtext{t\textit{ec}or}{\Afootnote{+ þvj L679.}} seint  
\edtext{orþ\textit{om} þ\textit{eir}ra}{\Afootnote{÷  L679.}}; 
þ\textit{ei}r  \edtext{sp\textit{yrja}}{\Afootnote{+ hann  L679.}} ept\textit{ir} Hróm\textit{undi} 
\edtext{h\textit{ann} qveþst þenkja}{\Afootnote{enn hann þeinkir   L679.}}, 
\edtext{a\textit{t} h\textit{ann} mani þ\textit{ar} v\textit{er}it h\textit{afa}}{\Afootnote{hann hafi verid þar  L679.}} 
no\textit{ck}orn tíma,  \edtext{e\textit{n}}{\Afootnote{og  L679.}} nú \edtext{er}{\Afootnote{muni  L679.}} h\textit{ann} \edtext{víst}{\Afootnote{÷  L679.}} \edtext{á burt}{\Afootnote{i burt  L679.}}, s\textit{egir} h\textit{ann}, o\textit{c} man heim
q\textit{vo}min\textit{n}; g\textit{an}ga þ\textit{ei}r nú t\textit{il}
bæ\textit{ar}ins, e\textit{n} k\textit{er}líng t\textit{il}
mjalta. Hún m\textit{ælir} \edtext{t\textit{il} Smala man\textit{n}s, 
	m\textit{eþ} há\textit{om} hljóþ\textit{om}}{\Afootnote{med háum hljödum til smala mannsins  L679.}}: 
hvaþa fj\textit{an}d\textit{an}s v\textit{er}kun er þetta, s\textit{em}
þú  gj\textit{ỏr}ir, þin\textit{n} ólu\textit{ck}o þ\textit{or}p\textit{ar}i, a\textit{t} þú brúk\textit{ar}
þ\textit{an}\textit{n} b\textit{an}\textit{n}aþ\textit{an}\textit{n} ósiþ,
a\textit{t} styggja sképnr, o\textit{c} venja t\textit{il} \edtext{óspekt\textit{ar}}{\Afootnote{öspekta  L679.}}, 
o\textit{c} ef þú leggor  \edtext{ei þetta}{\Afootnote{þetta ecki  L1572.}} af,  
\edtext{þá}{\Afootnote{÷  L679, L1572.}} skalto fá a\textit{t} kén\textit{n}a á  \edtext{h\textit{or}þo}{\Afootnote{+ enn  L679.}}.  \edtext{Smalinn}{\Afootnote{smala madr  L679, L1572.}} þagþi o\textit{c} sv\textit{ar}aþi engo. Nú leita þ\textit{ei}r
Bl\textit{indr} o\textit{c} h\textit{an}s fỏronaut\textit{ar},  
\edtext{ut\textit{an} bæ\textit{ar} o\textit{c}	 in\textit{n}an\textit{n}}{\Afootnote{uti oc inni  L1572.}}, 
o\textit{c} fin\textit{n}a  \edtext{nú}{\Afootnote{÷  L679, L1572.}} e\textit{ck}i þ\textit{at} villdo,  \edtext{e\textit{n} s\textit{em} áþr}{\Afootnote{÷  L679.}},  \edtext{sneipast þ\textit{ei}r o\textit{c} snúa áleiþis heim aptr}{\Afootnote{÷  L2404.}}; 
o\textit{c} s\textit{em} þ\textit{ei}r  \edtext{e\textit{ro}}{\Afootnote{voru  L679.}} \edtext{í þ\textit{at} pláts}{\Afootnote{þangat  L2404.}} q\textit{vo}mn\textit{ir}, s\textit{em} þ\textit{ei}r v\textit{or}o
v\textit{an}ir aptr a\textit{t} \edtext{hv\textit{er}fa}{\Afootnote{snua  L679.}}, 
s\textit{egir} Bl\textit{indr} 
Nú v\textit{ei}t e\textit{c} hv\textit{aþ} v\textit{er}it hefr,
a\textit{t} Hróm\textit{undr} hefr sá m\textit{aþr} v\textit{er}it,  \edtext{s\textit{e}m}{\Afootnote{er  L679.}} stóþ m\textit{e}þ lúþurin\textit{n}, o\textit{c} 
skul\textit{om} \edtext{nú}{\Afootnote{vid  L679.}} 
leita h\textit{an}s m\textit{eþ} ỏll\textit{om} ákafa. Snúa þ\textit{ei}r  \edtext{e\textit{n}}{\Afootnote{nú  L679.}} aptr t\textit{il} bæ\textit{ar} Hagals. 
Þá m\textit{ælir} Hag\textit{all} v\textit{it} kono sína: m\textit{ic}it h\textit{ar}þna nú skór
a\textit{t} fót\textit{om} mér, at  þetta g\textit{en}gr sva
lengi, a\textit{t} þ\textit{ei}r Bl\textit{indr} o\textit{c}
h\textit{an}s  \edtext{félag\textit{ar}}{\Afootnote{menn  L679.}}, g\textit{jỏr}a  \edtext{oss}{\Afootnote{mier  L679.}} slík\textit{ar} óspekt\textit{ir} f\textit{yrir} litl\textit{ar}
sak\textit{ir}, o\textit{c} nú vil e\textit{c} \edtext{sjálfr g\textit{an}ga}{\Afootnote{ganga sjálfr  L679, L2404.}} á  veg á móti þ\textit{ei}m, e\textit{n} þú mant  \edtext{sjá t\textit{il}}{\Afootnote{vel sjá um  L679.}} a\textit{t} v\textit{el} \edtext{t\textit{il} t\textit{ec}ist}{\Afootnote{takist til  L679, takist  L1572.}}; g\textit{en}gr  nú Hag\textit{all} \edtext{á móti}{\Afootnote{móti  L1572.}} þ\textit{ei}m a\textit{t} g\textit{ar}þi út, e\textit{n} er þ\textit{ei}r
sáu h\textit{an}n  v\textit{ar}þ h\textit{ann}
und\textit{ar}l\textit{e}gr í aug\textit{om} þ\textit{eir}ra, sýndist þ\textit{ei}m  \edtext{þá, at}{\Afootnote{÷  L679, þá  L1572.}} Flugdr\textit{ec}i  \edtext{sækja}{\Afootnote{sækti  L2404.}} a\textit{t} þ\textit{ei}m, t\textit{óc}o þ\textit{ei}r þá a\textit{t}
leita  \edtext{aptr}{\Afootnote{÷  L679.}} t\textit{il} b\textit{ak}a, o\textit{c} snúa heim leiþis, e\textit{n} \edtext{Draconin\textit{n}}{\Afootnote{drekinn  L1572.}} sókti eptir ||\ledouternote{127r} þ\textit{ei}m af meirsta kappi, sýndist þ\textit{ei}m h\textit{ann}
blása eldi  o\textit{c} eytri, villdo þá sum\textit{ir} bíþa
h\textit{an}s, e\textit{n} sum\textit{ir} f\textit{or}þa sér, v\textit{ar}þ
þ\textit{at} \edtext{þá}{\Afootnote{÷  L1572.}} fleyrst\textit{om} f\textit{yrir} \edtext{a\textit{t}}{\Afootnote{+ forda sier og   L679.}} leita  \edtext{sér}{\Afootnote{÷  L1572.}} und\textit{an} f\textit{ær}slo, v\textit{ar}aþi þ\textit{ess}i adsókn fram\textit{m} a\textit{t} qvỏldi, q\textit{vo}most
þ\textit{ei}r  \edtext{sva}{\Afootnote{þá  L679.}} u\textit{m} síþ\textit{ir} heim til  Hálfd\textit{áns}
K\textit{óng}s o\textit{c} sỏgþo  \edtext{sin\textit{n}}{\Afootnote{honum  L1572.}}  hrakníng  o\textit{c} eyrindis
leysi,  a\textit{t} Hróm\textit{undr} hefþi e\textit{ck}i
fundin\textit{n} \edtext{orþ\textit{it}}{\Afootnote{ordinn  L679, verid  L1572.}}, þókti  \edtext{nú}{\Afootnote{÷  L1572.}} K\textit{óng}i þessi  f\textit{er}þ  \edtext{ílla}{\Afootnote{+ oc hæþilega  L2404.}} t\textit{il}t\textit{ec}ist h\textit{afa},  \edtext{o\textit{c} hin hæþil\textit{e}gasta}{\Afootnote{÷  L2404.}}, leiþ sva  \edtext{á}{\Afootnote{÷  L679, L2404, L1572.}} Veturin\textit{n} a\textit{t} \edtext{lít\textit{it}}{\Afootnote{ecki  L679.}} b\textit{ar} t\textit{il} tíþinda 
\pend\endnumbering  
\beginnumbering\pstart  
\vspace{5mm}\subsection*{XVII kap\textit{ítuli}}  Um v\textit{or}it, \edtext{stutto f\textit{yrir} sum\textit{ar}}{\Afootnote{÷  L679.}}, q\textit{vo}m Bl\textit{indr} a\textit{t} máli v\textit{it} \edtext{Hálfdán}{\Afootnote{÷  L1572.}} \edtext{K\textit{óng}}{\Afootnote{÷  L2404.}} o\textit{c} s\textit{egir} \edtext{h\textit{onom} a\textit{t}}{\Afootnote{÷  L679.}} s\textit{ic} hafi  \edtext{m\textit{ar}ga drauma}{\Afootnote{÷  L2404.}} dreymt  und\textit{ar}l\textit{e}ga á
þ\textit{ess}\textit{om} vetri, o\textit{c} 
\edtext{villdi}{\Afootnote{vil L679.}} e\textit{c},  
\edtext{s\textit{egir} h\textit{ann} a\textit{t} þér, Herra!}{\Afootnote{herra ad þier  L679, herra segir hann, at þér  L2404.}},  \edtext{vildot ráþa}{\Afootnote{rádid  L679.}} þá t\textit{il} gagns o\textit{c} gam\textit{an}s. K\textit{óngr}
s\textit{egir} \edtext{vilja heyra}{\Afootnote{heira vilja  L679.}}.  Bl\textit{indr} s\textit{egir} þ\textit{at} dreymdi
m\textit{ic} fyrst, a\textit{t} mér þókti q\textit{vo}ma ein\textit{n} st\textit{or} o\textit{c} hræþil\textit{e}gr hýþbj\textit{ỏr}n,
h\textit{ann} óþ a\textit{t} yþr o\textit{c} beit yþr mikla  und  \edtext{á}{\Afootnote{i  L2404.}} brjóst\textit{it}. 
K\textit{óngr} s\textit{egir} \edtext{híng\textit{at}}{\Afootnote{hier  L679.}} kémr an\textit{n}\textit{at}hv\textit{or}t K\textit{óng}r eþr  \edtext{an\textit{n}\textit{ar}}{\Afootnote{÷  L2404.}} kappi,  \edtext{s\textit{em} vekr hér}{\Afootnote{og vilt hefja  L679.}} styrjỏlþ v\textit{it} m\textit{ic},  \edtext{e\textit{n} þ\textit{ar} á}{\Afootnote{og þar L679.}} ept\textit{ir} sættist  h\textit{ann} o\textit{c} e\textit{c}.
Bl\textit{indr} m\textit{ælir} an\textit{n}\textit{an}\textit{n} draum
dreymdi m\textit{ic},  \edtext{s\textit{egir} Bl\textit{indr}}{\Afootnote{÷  L679, L2404, L1572.}} a\textit{t} ec  þóktist q\textit{vo}ma í eitt st\textit{or}t
o\textit{c} vænt hús, þ\textit{ar} sáto m\textit{ar}g\textit{ir} Fálkar á stỏll\textit{om}, v\textit{or}o þ\textit{ei}r all\textit{ir}
m\textit{eþ} \edtext{fỏgr\textit{om}}{\Afootnote{fóstum  L679.}} fjỏþr\textit{om}, e\textit{n} sá fálkin\textit{n} s\textit{em} þér attot,  \edtext{h\textit{ann}}{\Afootnote{÷  L679, L1572.}} v\textit{ar} \edtext{fjaþralaus, o\textit{c}}{\Afootnote{÷  L679.}} allr  \edtext{rifin\textit{n}}{\Afootnote{+ og  L679.}} af h\textit{onom} hamrin\textit{n}. 
K\textit{óng}r  \edtext{m\textit{ælir}}{\Afootnote{segir  L679.}} þ\textit{at} v\textit{er}þr m\textit{ic}ill st\textit{or}mvindr,
s\textit{em} hrist\textit{ir} o\textit{c} \edtext{skékur}{\Afootnote{+ allt  L2404.}} þ\textit{at} h\textit{ann} á blæs, v\textit{er}þ e\textit{c} þá staþdr
n\textit{ær}ri eldi o\textit{c} \edtext{fjúka}{\Afootnote{hagli  L679.}} \edtext{þá}{\Afootnote{÷  L1572.}} \edtext{gneist\textit{ar}n\textit{ir}}{\Afootnote{gneistrast  L679.}} \edtext{á}{\Afootnote{um  L679.}} m\textit{ic}. 
Þriþji  \edtext{draumorin\textit{n}}{\Afootnote{draumr minn  L2404.}} v\textit{ar} þ\textit{at}, s\textit{egir} Bl\textit{indr} a\textit{t} e\textit{c} þóktist sjá m\textit{ar}gt Svín
ren\textit{n}a hér a\textit{t} sun\textit{n}\textit{an}, þau rótoþo upp allri  \edtext{j\textit{ỏr}þun\textit{n}i}{\Afootnote{jörþu  L2404.}} \edtext{um l\textit{an}d\textit{it}}{\Afootnote{÷  L2404.}}. 
K\textit{óng}r m\textit{ælir} þ\textit{at} v\textit{er}þr fyr\textit{ir} ||\ledouternote{127v} sjáf\textit{ar} ókyrrleika o\textit{c} sólskyni, ásamt
f\textit{yrir} v\textit{ar}ma o\textit{c} vorgróþa. 
Þat er \edtext{min\textit{n}}{\Afootnote{÷ L679.}} sá fjórþi \edtext{draumr}{\Afootnote{+minn L679.}}, 
s\textit{egir} Bl\textit{indr} a\textit{t}
e\textit{c} sá  l\textit{an}g\textit{an} j\textit{ar}þ\textit{ar}orm skríþa  \edtext{hér}{\Afootnote{÷  L2404.}} 
a\textit{t} aust\textit{an} o\textit{c} in\textit{n} í hỏllina, sá beit yþr  \edtext{Herra}{\Afootnote{÷  L679.}} bæþi  
\edtext{djúpa o\textit{c} breiþa}{\Afootnote{breida og djupa  L679.}} und á brjóst\textit{it}.  
\edtext{Þat  er, s\textit{egir} K\textit{óngr}}{\Afootnote{Kongr segir þad er  L679.}} f\textit{yrir} einhv\textit{or}j\textit{om} stór\textit{om} hỏfþíngja
s\textit{em} \edtext{á min\textit{n} fund  sæk\textit{ir}}{\Afootnote{sækir á minn fund  L679.}}, h\textit{ann} man vilja skoþa ỏll vopn v\textit{or}, af hv\textit{or}ri h\textit{an}s forvitni  \edtext{a\textit{t}}{\Afootnote{÷  L679, L2404.}} e\textit{c} man reiþast. 
Þ\textit{at} 
\edtext{b\textit{ar} e\textit{n} t\textit{il}}{\Afootnote{dreimdi mig enn  L679.}} 
s\textit{egir} Bl\textit{indr} 
\edtext{a\textit{t} m\textit{ic} dreymdi  a\textit{t}}{\Afootnote{÷  L679.}} 
\edtext{hỏggormr ein\textit{n}}{\Afootnote{Eirn hóggorm  L679.}} 
\edtext{v\textit{ar} sva st\textit{or}}{\Afootnote{svo storann  L679.}}, a\textit{t} h\textit{ann} náþi  \edtext{hér}{\Afootnote{÷  L679.}} yf\textit{ir} \edtext{u\textit{m} þvert}{\Afootnote{allt  L679.}} l\textit{an}d\textit{it}.  K\textit{óng}r m\textit{ælir} 
þ\textit{at} v\textit{er}þr eitt \edtext{mj\textit{ỏc}  príþil\textit{e}gt}{\Afootnote{÷  L2404.}} \edtext{o\textit{c}}{\Afootnote{÷  L679, L2404.}} \edtext{blóml\textit{e}gt}{\Afootnote{÷  L679.}} dr\textit{ec}a  skip, s\textit{e}m híng\textit{at}
sigl\textit{ir} an\textit{n}\textit{ar}staþ\textit{ar} a\textit{t}. 
Sá er hin\textit{n} sjötti  draumr min\textit{n}, s\textit{egir}
Bl\textit{indr} a\textit{t} e\textit{c} sá kolsv\textit{ỏr}t ⸌ský⸍ q\textit{vo}ma 
\edtext{or}{\Afootnote{af  L679.}} n\textit{or}þri, 
þau hỏfþo klær o\textit{c} \edtext{bjúga vængi}{\Afootnote{vængi bjúga  L679.}}, 
þau flugo  \edtext{burto}{\Afootnote{búrt L679.}} m\textit{eþ} \edtext{þ\textit{ic}}{\Afootnote{jdr  L679.}},  
\edtext{oc}{\Afootnote{so  L679, L2404, L1572.}} e\textit{c} \edtext{vissi}{\Afootnote{sá  L679.}} e\textit{ck}i hvaþ af yþr v\textit{ar}þ, o\textit{c} þ\textit{ar}
ept\textit{ir} sá e\textit{c} yþ\textit{ar} betsta  \edtext{dr\textit{ec}askip}{\Afootnote{skip  L679.}}, m\textit{ar}a í  \edtext{miþjo}{\Afootnote{÷  L679.}} kafi í  \edtext{brimi}{\Afootnote{÷  L679, L1572.}} \edtext{o\textit{c}}{\Afootnote{÷  L679.}} \edtext{sjó\textit{ar}æþi g\textit{an}gi}{\Afootnote{sióargangi  L2404.}},  
e\textit{n} allr  \edtext{herin\textit{n} yþ\textit{ar}}{\Afootnote{ydar her  L679, her yþar  L2404.}} stóþ hỏfotlaus niþr í  \edtext{eino vatni}{\Afootnote{vatni eino  L2404.}},  síþ\textit{an} þóttist e\textit{c} staþdr n\textit{ær}ri bæ
Hagals,  \edtext{kom þ\textit{ar}}{\Afootnote{þar kom  L679.}} út  hỏggormr,  \edtext{h\textit{ann} v\textit{ar}}{\Afootnote{÷  L1572.}} íllúþl\textit{e}gr, h\textit{ann} beit  \edtext{m\textit{enn}}{\Afootnote{mic  L2404.}} t\textit{il} \edtext{b\textit{an}a}{\Afootnote{+ og  L679, L1572.}},  \edtext{síþ\textit{an}}{\Afootnote{÷  L1572.}} át  \edtext{h\textit{ann}}{\Afootnote{+ þá  L679, ÷  L1572.}} alla  \edtext{s\textit{em}}{\Afootnote{er  L1572.}} h\textit{ann} \edtext{beit}{\Afootnote{+ og  L679, L2404.}}, síþ\textit{ar}st át h\textit{ann} yþr o\textit{c}
m\textit{ic}. K\textit{óng}r  \edtext{m\textit{ælir}}{\Afootnote{segir  L679, L1572.}} þ\textit{at} er  \edtext{þ\textit{at}}{\Afootnote{÷  L679.}}, s\textit{em} e\textit{c} \edtext{lengi hefi}{\Afootnote{hefi leingi  L679, L2404, L1572.}} ætla\textit{t} a\textit{t} láta vin\textit{n}a,  \edtext{a\textit{t}}{\Afootnote{og  L679.}} drepa hýþbj\textit{ỏr}n þan\textit{n},  s\textit{em}
hj\textit{or}þ\textit{om} m\textit{ann}a hefr  \edtext{opt}{\Afootnote{ádr  L679.}} mein g\textit{jỏr}t, o\textit{c} er skam\textit{m}t frá  koti
Hag\textit{als}.  En\textit{n} dreymdi m\textit{ic},  \edtext{s\textit{egir} Bl\textit{indr}}{\Afootnote{÷  L679.}} a\textit{t} úngr hỏggormr v\textit{ar} dregin\textit{n} út or hỏllin\textit{n}i, o\textit{c} h\textit{an}gdi sv\textit{er}þ v\textit{it} h\textit{an}s  síþo, s\textit{em} Mistiltein\textit{n} h\textit{ie}t.  
Þetta er allt drauma rugl,  s\textit{egir}
K\textit{óng}r, þ\textit{ví} Mistiltein\textit{n} liggr  \edtext{í}{\Afootnote{á  L679.}} botnin\textit{om} á Væn\textit{ir}, o\textit{c} þarf 
þ\textit{ic} e\textit{ck}i þetta a\textit{t} skélfa; eþa erto
orþin\textit{n} sú bleyþa, a\textit{t} þú \edtext{hyggr}{\Afootnote{hugsar  L679.}} a\textit{t} allt mani teikna b\textit{ar}daga o\textit{c} blóþs
úthellíng||\ledouternote{128r}ar 
Bl\textit{indr} m\textit{ælir} fleyri drauma
hefr m\textit{ic} e\textit{n} dreymt, mætti  \edtext{ec}{\Afootnote{+ þá  L679.}} segja. K\textit{óng}r m\textit{ælir}, segja skalto, e\textit{n}
e\textit{c} man ráþa  \edtext{þá}{\Afootnote{sem  L1572.}} e\textit{c} vil; 
Þat  dreymdi m\textit{ic} s\textit{egir}
Bl\textit{indr} \edtext{mér þókti a\textit{t}}{\Afootnote{ad mier þotti  L679, L1572, at mier þotti at  L2404.}} slegin\textit{n} \edtext{v\textit{er}a}{\Afootnote{÷  L679.}} hríngr af rauda gulli  \edtext{betsta}{\Afootnote{÷  L679, L2404, L1572.}} u\textit{m} hálsin\textit{n} \edtext{á yþr}{\Afootnote{þier  L679.}}, þ\textit{ar} ept\textit{ir} vissi e\textit{c},  
\edtext{á}{\Afootnote{ad  L679, L2404, L1572.}} þér  riþot  \edtext{hesti}{\Afootnote{+ svo lótum  L679.}}, s\textit{em} þér gátot e\textit{ck}i  \edtext{nú\textit{it}}{\Afootnote{+ áframm  L679.}} f\textit{yrir} leti, o\textit{c} 
af þ\textit{ví} urþot þér   \edtext{sva móþ\textit{ir} o\textit{c} þreytt\textit{ir}}{\Afootnote{þreittir og modir  L679, módr og þreittr  L1572.}},  \edtext{sva}{\Afootnote{at  L2404, L1572.}} m\textit{ic} furþ\textit{at}i, o\textit{c} man v\textit{er}þa  \edtext{þúngt}{\Afootnote{langt  L679.}} a\textit{t} ráþa  \edtext{þ\textit{ess}a drauma}{\Afootnote{þá alla  L679.}}; b\textit{ar} \edtext{þá Bl\textit{indr}}{\Afootnote{Blindr þá  L2404.}} fram\textit{m} fleyri drauma, þó  \edtext{at}{\Afootnote{÷ L679.}} hér séo e\textit{ck}i ritaþ\textit{ir}, o\textit{c} réþi
Hálfd\textit{án} K\textit{óngr} þá  \edtext{alla}{\Afootnote{÷  L679.}} ept\textit{ir} \edtext{sin\textit{n}i  vild}{\Afootnote{villd sinni  L679.}}, e\textit{n} e\textit{ck}i  \edtext{ept\textit{ir} þ\textit{ví}}{\Afootnote{því eptir  L2404.}} \edtext{s\textit{em}}{\Afootnote{+ þeir  L1572.}} m\textit{er}kto, sva s\textit{em} síþ\textit{ar} \edtext{gaf raun}{\Afootnote{raun gaf  L679.}} vitni.  Þá  \edtext{s\textit{egir}}{\Afootnote{mælir  L1572.}} K\textit{óngr} hv\textit{aþ} hyggr þú  \edtext{a\textit{t}}{\Afootnote{÷  L679.}} þ\textit{ess}i  \edtext{hin\textit{n} síþ\textit{ar}sti}{\Afootnote{÷  L679.}} draumr hafi  a\textit{t} þýþa? 
Blindr s\textit{egir} gamall
o\textit{c} latr hestr m\textit{er}k\textit{ir} gálga, 
e\textit{n} hríngr af  \edtext{betsta}{\Afootnote{rauþa  L2404.}} gulli  \edtext{á hálsi}{\Afootnote{um háls  L679.}} \edtext{b\textit{or}in\textit{n}}{\Afootnote{+ merkir  L2404, L1572.}}, svívyrþíngar  sn\textit{or}ob\textit{an}d, o\textit{c}
man\textit{om} v\textit{it} báþ\textit{ir} skam\textit{m}a stund lifa.
K\textit{óngr} m\textit{ælir} sva  \edtext{rætist}{\Afootnote{rætaz  L2404.}} \edtext{draumr}{\Afootnote{draumar  L2404.}} s\textit{em} \edtext{h\textit{ann} er ráþin\textit{n}}{\Afootnote{ráþnir ero  L2404.}} \edtext{t\textit{il}}{\Afootnote{÷  L679, L2404, L1572.}}, 
o\textit{c} man sva v\textit{er}þa  a\textit{t} íll
samvitska á sér jafn\textit{an} Bỏþul, skildo  \edtext{þ\textit{ei}r}{\Afootnote{÷  L679, L2404.}} sva  tal\textit{it},  \edtext{a\textit{t} sitt}{\Afootnote{oc  L2404.}} þókti  \edtext{hv\textit{or}jum þeirra}{\Afootnote{hvorjum  L679, sitt hveriom  L2404.}}. 
\pend \endnumbering 
\beginnumbering \pstart 
\vspace{5mm} \subsection*{XVIII kap\textit{ítuli}}
 \edtext{Nú er a\textit{t} s\textit{egja} frá Hróm\textit{undi}}{\Afootnote{Um Romund er þat at segja  L2404.}} a\textit{t} h\textit{ann} hressist hjá Hagal\footnote{Note the written-out form \enquote{hjá Hagal} instead of \textit{hjá Hagali} or \textit{hjá Hagli}.}  k\textit{ar}li  \edtext{í koti h\textit{an}s}{\Afootnote{÷  L679, L2404, L1572.}}, þ\textit{ar} t\textit{il} \edtext{a\textit{t}}{\Afootnote{÷  L679, L2404, L1572.}} h\textit{ann} \edtext{er}{\Afootnote{var  L1572.}} gróin\textit{n} sára sin\textit{n}a, fýs\textit{ir} h\textit{ann} þá at \edtext{vitja}{\Afootnote{finna  L679.}} \edtext{fỏþ\textit{ur}s síns}{\Afootnote{fódur sinn  L679, L1572.}}; 
 o\textit{c} s\textit{em} h\textit{ann} sj\textit{er} Hró\-m\textit{und} q\textit{vo}ma, fagn\textit{ar} \edtext{Gr\textit{eipr}}{\Afootnote{hann  L679.}} h\textit{onom} \edtext{af meirsta blíþskap}{\Afootnote{blijdlega  L679, vel  L2404.}}, 
 o\textit{c} sp\textit{yr} tíþinda. 
 Hró\-m\textit{undr} s\textit{egir} slík  \edtext{s\textit{em}}{\Afootnote{er  L1572.}} t\textit{il} fall\textit{it} \edtext{hỏfþo}{\Afootnote{hafdi  L679.}}, a\textit{t} s\textit{ynir} h\textit{an}s  \edtext{v\textit{ær}o}{\Afootnote{ero  L679.}} falln\textit{ir}. 
 Gr\textit{eipr} s\textit{egir} a\textit{t} nú  \edtext{mano}{\Afootnote{sieu  L679, muni  L2404, L1572.}} \edtext{fram\textit{m}\-komn\textit{ir}}{\Afootnote{framkoma  L2404.}} 
 draum\textit{ar} sín\textit{ir} hin\textit{ir} f\textit{or}no í  \edtext{no\textit{ck}rom}{\Afootnote{sumum  L679.}} lutom. 
 Hróm\textit{undr} \edtext{dvelr}{\Afootnote{dvaldi  L1572.}} þá m\textit{eþ} fỏþr sín\textit{om}, þ\textit{at} ept\textit{ir}
v\textit{ar} \edtext{vetr\textit{ar}ins}{\Afootnote{vetursins  L679.}}; e\textit{n} u\textit{m} v\textit{or}it  \edtext{lætr}{\Afootnote{liet  L679.}} Ol\textit{afr} K\textit{óng}r  \edtext{h\textit{er}ỏr}{\Afootnote{herfór  L679.}} uppskéra, o\textit{c} blása t\textit{il} b\textit{ar}daga,
o\textit{c} segist vilja aptr t\textit{il} Svíþjóþ\textit{ar} a\textit{t}
vin\textit{n}a  Hálfd\textit{án} K\textit{óng} \edtext{þetta þókti}{\Afootnote{þotti þetta  L679, L2404.}} fleyrst\textit{om} und\textit{ar}l\textit{e}gt, þ\textit{ví} \edtext{a\textit{t}}{\Afootnote{÷  L679.}} ||\ledouternote{128v} þ\textit{ei}r  \edtext{Ol\textit{afr} K\textit{óngr} o\textit{c} Hálfd\textit{án}}{\Afootnote{÷  L679, + kongr  L1572.}} sættost sín á \edtext{mill\textit{om}}{\Afootnote{milli  L679.}}, 
ept\textit{ir} dauþa \edtext{Helga o\textit{c} þ\textit{eir}ra b\textit{ræþ}ra}{\Afootnote{÷  L679.}},  \edtext{Vola oc Bílds}{\Afootnote{Bilds og Vola  L679.}}, 
e\textit{n} af þ\textit{ei}m Gr\textit{eips} son\textit{om} kun\textit{n}o  fá\textit{ir} a\textit{t} segja;  \edtext{e\textit{n}}{\Afootnote{÷  L2404.}} ei a\textit{t} síþr drifo  \edtext{þá}{\Afootnote{÷ L679.}} m\textit{ar}g\textit{ir} m\textit{enn}
t\textit{il} \edtext{Ol\textit{afs}}{\Afootnote{÷  L679.}} K\textit{óngs} o\textit{c} v\textit{ar}þ  \edtext{h\textit{an}s}{\Afootnote{+ lids  L679.}} t\textit{il}búníngr  \edtext{bæþi}{\Afootnote{÷  L679.}} m\textit{ic}ill o\textit{c} l\textit{an}gr, f\textit{or} þetta allt
m\textit{eþ} fá\textit{om} fréttom  
\edtext{o\textit{c} afspurn\textit{om}}{\Afootnote{og afspúrdúm L679, ÷  L2404.}}. 
Þat v\textit{ar} eitt sin\textit{n},  \edtext{a\textit{t}}{\Afootnote{er  L1572.}} \edtext{Ol\textit{afr}}{\Afootnote{÷  L679.}} K\textit{óngr} v\textit{ar} a\textit{t} \edtext{láta}{\Afootnote{÷  L679, L1572.}} búa h\textit{er} sin\textit{n}, a\textit{t} h\textit{ann} \edtext{hefr uppræþo v\textit{it} m\textit{enn} sína,
	o\textit{c}}{\Afootnote{÷  L1572.}} sp\textit{yr} ept\textit{ir} hv\textit{or}t  \edtext{a\textit{t}}{\Afootnote{÷  L679, L1572.}} þ\textit{ei}r viti no\textit{ck}ot t\textit{il} Hróm\textit{undar}
o\textit{c} b\textit{ræþ}ra h\textit{an}s, síþ\textit{an} \edtext{a\textit{t}}{\Afootnote{÷  L2404.}} h\textit{ann} hv\textit{ar}f  \edtext{í}{\Afootnote{÷  L679, á  L1572.}} burt  héþ\textit{an},  \edtext{s\textit{egir} h\textit{ann}}{\Afootnote{÷  L679, L2404.}}?  \edtext{M\textit{enn} K\textit{óng}s}{\Afootnote{þeir  L679, menn  L2404.}} qváþo lít\textit{it} \edtext{f\textit{yrir}}{\Afootnote{frá  L1572.}} þ\textit{ví},  \edtext{o\textit{c}}{\Afootnote{at  L1572.}} segja  \edtext{a\textit{t}}{\Afootnote{÷  L2404.}} h\textit{ann} \edtext{sjest hafi}{\Afootnote{hafi siest  L679, séþ hafa  L2404.}} hino megin\textit{n} Væn\textit{ir}s o\textit{c} \edtext{þ\textit{ei}r}{\Afootnote{þá  L2404.}} b\textit{ræþ}r h\textit{an}s,  \edtext{e\textit{n}}{\Afootnote{÷  L679.}} ei visso  \edtext{m\textit{enn}, s\textit{egja}}{\Afootnote{÷  L679.}} þ\textit{ei}r, hv\textit{or}t  \edtext{heldr a\textit{t} h\textit{ann} fylgdi}{\Afootnote{þeir filgdu  L679, hann filgdi  L1572.}} \edtext{yþ\textit{ar} m\textit{ỏnnom}, eþr þ\textit{ei}m Sví\textit{om}}{\Afootnote{ydr eda Svjum  L679.}}. 
K\textit{óngr} m\textit{ælir} sá þá  engin\textit{n} hv\textit{aþ} þ\textit{ei}r a\textit{t} hỏfþost.  
Þá sv\textit{ar}aþi \edtext{t\textit{il} m\textit{aþr}}{\Afootnote{madr til  L679, L1572.}}, o\textit{c} s\textit{agþi} þ\textit{at} hygg e\textit{c},  \edtext{s\textit{egir} h\textit{ann}}{\Afootnote{÷  L2404.}}, a\textit{t} Hróm\textit{undr} \edtext{mani lifa}{\Afootnote{lifi  L679.}} e\textit{n} nú, o\textit{c} \edtext{dvelja}{\Afootnote{sie  L679, vera  L2404.}} hjá Gr\textit{eipi} e\textit{n} b\textit{ræþ}r  h\textit{an}s
drap  \edtext{Helgi}{\Afootnote{+ hinn frækni  L679.}} v\textit{it} Væn\textit{ir}s vatn í vetr, o\textit{c} er sú t\textit{il} \edtext{géta}{\Afootnote{gáta  L2404, L1572.}},  a\textit{t} Hróm\textit{undr} \edtext{mani h\textit{afa}}{\Afootnote{hafi  L679.}} \edtext{vegit}{\Afootnote{drepid  L1572.}} Helga. Hv\textit{aþ} man því  valda, s\textit{egir}
K\textit{óngr} a\textit{t} \edtext{h\textit{ann}}{\Afootnote{Hromundr  L679.}} kémr e\textit{ck}i á  \edtext{fund v\textit{or}n}{\Afootnote{minn fund  L679, vorn fund  L2404.}}? Þat man þ\textit{ví} valda, s\textit{egir} \edtext{þ\textit{ess}i m\textit{aþr}}{\Afootnote{madrinn  L679.}}, a\textit{t} h\textit{ann} 
\edtext{vill e\textit{ck}i vitja yþ\textit{ar}, a\textit{t} h\textit{ann}}{\Afootnote{÷  L679, L1572.}} þurfti at f\textit{or}þa lífi síno \edtext{seinast}{\Afootnote{sijdast L679, L2404, L1572.}}, 
þá h\textit{ann}
flýþi  \edtext{héþ\textit{an}}{\Afootnote{÷  L679.}} f\textit{yrir} reiþi yþ\textit{ar},  o\textit{c} \edtext{sein\textit{n}a}{\Afootnote{seint  L679.}} man  \edtext{þ\textit{at} v\textit{er}þa, a\textit{t} h\textit{ann}
	híng\textit{at} q\textit{vo}mi}{\Afootnote{hann hijngad koma  L679.}}, fyrr e\textit{n} h\textit{ann} hefr  hefnt síns
v\textit{an}sa, s\textit{em} h\textit{ann} þóktist fá hjá yþr, f\textit{yrir}
álygar  Bílds o\textit{c} Vola. 
K\textit{óngr} m\textit{ælir} \edtext{Viljiþ}{\Afootnote{vilie  L2404.}} 
\edtext{þér}{\Afootnote{þid L679.}} láta Hróm\textit{und} 
v\textit{er}a  yþ\textit{ar} \edtext{h\textit{er}fyl\-kín\-ga yf\textit{ir}\-m\textit{an}\textit{n}}{\Afootnote{oddvita  L2404.}}, eþr ei? 
ef þér þ\textit{at} viljiþ,  \edtext{þá}{\Afootnote{÷  L679, L1572.}} \edtext{skal}{\Afootnote{+ eg  L1572.}} ept\textit{ir} h\textit{onom} senda, o\textit{c} sæmda boþ bjóþa.
Men\textit{n} qváþost ei  \edtext{v\textit{el}}{\Afootnote{÷  L2404.}} vita, 
hv\textit{or}t h\textit{ann} villdi sætt\textit{om} t\textit{ac}a,
eþr e\textit{ck}i, þ\textit{ví} h\textit{ann} hefr þúnga lund o\textit{c} þétta. K\textit{óngr} biþr  \edtext{þá}{\Afootnote{÷  L2404.}} \edtext{m\textit{enn}}{\Afootnote{+ ad  L679, + sina  L2404.}} f\textit{ar}a, oc  fin\textit{n}a Hróm\textit{und}
o\textit{c} b\textit{er}a h\textit{onom} orþ sín, a\textit{t} h\textit{ann} \edtext{býþr}{\Afootnote{bjódi  L679.}} h\textit{onom} all\textit{ar} þ\textit{ær} sæmd\textit{ir},  \edtext{er}{\Afootnote{sem  L679.}} h\textit{ann} \edtext{þykist}{\Afootnote{frekast  L679.}} v\textit{ei}ta  \edtext{mega}{\Afootnote{meigi  L679.}}. 
Nú  \edtext{q\textit{vo}ma}{\Afootnote{komu  L2404.}} m\textit{enn} á  \edtext{Bæ}{\Afootnote{fund  L2404, L1572.}} Hróm\textit{undar} o\textit{c} 
\edtext{qveþja}{\Afootnote{kvóddú L679.}}\footnote{
	An uncertain reading in L2404 due to the abbreviation \enquote{qv}.}
 h\textit{ann} t\textit{il} máls,
e\textit{n} h\textit{ann} qveþst  \edtext{hafa ||\ledouternote{129r} an\textit{n}\textit{at} ráþstofn\textit{at}}{\Afootnote{annad rád stofnad hafa  L679.}},  \edtext{heldr}{\Afootnote{÷  L679, L2404.}} e\textit{n} \edtext{a\textit{t}}{\Afootnote{÷  L679, L1572.}} láta fram\textit{an}di m\textit{enn} tefja s\textit{ic} eyrinda sin\textit{n}a. Þeir ⸌s\textit{egja}⸍ h\textit{onom} orþ Ol\textit{afs} K\textit{óngs} o\textit{c} sæmda boþ h\textit{an}s.
Hróm\textit{undr} s\textit{egist} \edtext{e\textit{ck}i h\textit{afa} f\textit{ar}it neina}{\Afootnote{enga  L1572.}} 
\edtext{sæmd\textit{ar} f\textit{er}þ}{\Afootnote{sæmda fór  L679, L2404, + hafa farid  L1572.}} \edtext{f\textit{yrir}}{\Afootnote{siþaz fyrir  L2404.}} K\textit{óng} \edtext{híng\textit{at} t\textit{il}}{\Afootnote{÷  L2404.}}, o\textit{c} sva  \edtext{mani}{\Afootnote{man  L2404, L1572.}} e\textit{n} v\textit{er}þa,  \edtext{þó}{\Afootnote{÷  L2404.}} \edtext{h\textit{ann}}{\Afootnote{eg  L679, ÷  L2404.}} \edtext{fin\textit{n}i h\textit{ann}}{\Afootnote{÷  L2404.}}, hafi  \edtext{K\textit{óngr} siþ\textit{ar}st}{\Afootnote{hann oc  L2404.}} h\textit{ei}t\textit{it} sér gálga  \edtext{n\textit{ær}}{\Afootnote{þá  L2404.}} h\textit{ann} hafi sér blygþ á hendr b\textit{or}it, o\textit{c} \edtext{sín leitaþi}{\Afootnote{sín leitad  L679, L1572, leytat sin  L2404.}} í h\textit{er}b\textit{er}gjom  Svanh\textit{vítar} líka
kun\textit{n}i h\textit{an}s ráþaneyti lengi a\textit{t} duga, 
s\textit{em} \edtext{a\textit{t}}{\Afootnote{÷  L679.}} séo  \edtext{þ\textit{ei}r}{\Afootnote{+ bræþr  L2404.}} Bíldr o\textit{c} Vole. Sendim\textit{enn} qváþo þá dauþa báþa
f\textit{yrir} laungo. Hróm\textit{undr} létst þ\textit{at} ei vita, o\textit{c}
b\textit{at} \edtext{þá}{\Afootnote{÷  L2404.}} b\textit{er}a  
\edtext{Ol\textit{afi}}{\Afootnote{÷  L679.}} K\textit{óngi} orþ  sín. 
 \pend \endnumbering  
 \beginnumbering \pstart  
\vspace{5mm} \subsection*{XIX kap\textit{ítuli}}
 Sendim\textit{enn} Ol\textit{afs} K\textit{óng}s ríþa nú sva  \edtext{bún\textit{ir}}{\Afootnote{buit  L2404.}} heim aptr, o\textit{c} s\textit{egja} K\textit{óngi} orþ Hróm\textit{undar} K\textit{óngr} m\textit{ælir} ílla f\textit{or} þ\textit{at}, a\textit{t} Hróm\textit{undr} \edtext{f\textit{er}}{\Afootnote{for  L679.}} e\textit{ck}i t\textit{il} \edtext{f\textit{er}þ\textit{ar}}{\Afootnote{fundr  L679.}} \edtext{m\textit{eþ}}{\Afootnote{vid  L679.}} o\textit{ss},  þ\textit{ví} \edtext{a\textit{t}}{\Afootnote{÷  L679.}} óvíst  \edtext{v\textit{er}þr}{\Afootnote{er þat  L679.}}, 
 hv\textit{or}t  \edtext{vér}{\Afootnote{vid  L679.}} fá\textit{om} \edtext{Hálfd\textit{án} K\textit{óng} sigr\textit{at}}{\Afootnote{sigrad Halfdan Kong  L679.}} m\textit{eþ} v\textit{or}t eindæmi. 
 Sendim\textit{enn}
sỏgþost  \edtext{halda}{\Afootnote{+ övijst  L679.}},  \edtext{a\textit{t}}{\Afootnote{÷  L2404.}} Hrom\textit{undr} m\textit{un}di hv\textit{or}gi  \edtext{f\textit{ar}it}{\Afootnote{farad  L679.}} h\textit{afa}, hv\textit{or} boþ s\textit{em} \edtext{feng\textit{it} hefþi}{\Afootnote{hefþi fengit  L2404.}}. Ol\textit{afr} K\textit{óngr} lætr nú \edtext{m\textit{enn}}{\Afootnote{÷   L679.}} búast t\textit{il} burtf\textit{er}þ\textit{ar} t\textit{il} Svía r\textit{ic}is, v\textit{or}o  \edtext{þó}{\Afootnote{÷  L1572.}} m\textit{enn} \edtext{m\textit{ic}it}{\Afootnote{÷  L679, þess miok  L2404, til þess miog  L1572.}} ófús\textit{ir}, þ\textit{ví} \edtext{þ\textit{ei}r þeinkto}{\Afootnote{þeim þokti örvænt  L2404.}},  \edtext{a\textit{t} síþr m\textit{un}di}{\Afootnote{÷  L2404.}} \edtext{sigurin\textit{n} fást}{\Afootnote{fást sigr  L679, um sigr  L2404.}} \edtext{e\textit{n} áþr}{\Afootnote{÷  L2404.}}.  
Nú er a\textit{t} s\textit{egja} frá Sv\textit{an}h\textit{víti} a\textit{t} hún g\textit{jỏr}ir orþ Hróm\textit{undi} o\textit{c} biþr  h\textit{ann} q\textit{vo}ma á sin\textit{n} fund,  \edtext{o\textit{c} tefja ei}{\Afootnote{sem bráþaz  L2404.}} \edtext{v\textit{it} a\textit{t}}{\Afootnote{svo  L679, ÷  L2404.}} \edtext{tím\textit{om} skipti}{\Afootnote{÷  L2404.}}. 
Hróm\textit{undi} brá v\textit{it} \edtext{h\textit{it} sn\textit{ar}asta}{\Afootnote{skjótt  L679.}}, o\textit{c} t\textit{óc} \edtext{h\textit{er}klæþi sín o\textit{c} reiþskjóta}{\Afootnote{reidskiota sinn og herklædi  L679.}}, o\textit{c} q\textit{vo}m  á fund Svanh\textit{vítar} 
\edtext{Svanh\textit{vít}}{\Afootnote{hún  L679, L2404.}} \edtext{fagnaþi}{\Afootnote{fagnar  L679.}} \edtext{Hróm\textit{undi} o\textit{c} h\textit{ann}
	hen\textit{n}i}{\Afootnote{hónum vel  L679, honum med allri blíþu  L2404, Hromundi  L1572.}},  
\edtext{o\textit{c}}{\Afootnote{Hrómundr  L679, L2404, hann  L1572.}} sp\textit{yr} \edtext{hvaþa}{\Afootnote{at  L2404, hvort  L1572.}} eyrindi  \edtext{hún vilji, 
	a\textit{t} h\textit{ann} útrétti}{\Afootnote{÷  L2404.}}. Hún m\textit{ælir} \edtext{nú vil e\textit{c}}{\Afootnote{ek vil  L2404.}} a\textit{t} þú b\textit{or}g\textit{ir} mér  \edtext{húsaleigo}{\Afootnote{husaleiguna  L679.}}, s\textit{em} e\textit{c} veitti  \edtext{þér}{\Afootnote{+ þá þú varst  L679.}} hjá  Hagali k\textit{ar}li, a\textit{t} þú f\textit{ar}ir
\edtext{nú}{\Afootnote{÷  L679.}} t\textit{il} Svíþjóþ\textit{ar}, m\textit{eþ} br\textit{óþr}
mín\textit{om} f\textit{yrir} mín orþ, o\textit{c} skul\textit{om} v\textit{it} \edtext{nú}{\Afootnote{÷  L679.}} bæþi g\textit{an}ga á K\textit{óng}s fund, o\textit{c} vita 
hv\textit{or}jo  \edtext{a\textit{t}}{\Afootnote{÷  L679, L2404, L1572.}} \edtext{h\textit{ann} vill góþo}{\Afootnote{godu hann vill  L679, L1572.}} t\textit{il} sv\textit{ar}a. Gánga þau þá  \edtext{bæþi}{\Afootnote{÷  L679.}} t\textit{il} K\textit{óng}s  K\textit{óngr} fagn\textit{ar}
þ\textit{ei}m  \edtext{v\textit{el}}{\Afootnote{+ enn  L679.}}, 
þó betst Hróm\textit{undi} o\textit{c} bauþ h\textit{onom} t\textit{il}
sætis,  \edtext{e\textit{n}}{\Afootnote{÷  L679.}} 
Hróm\textit{undr} qvaþst ei  \edtext{sæti þurfa}{\Afootnote{sætis þúrfi  L679.}} \edtext{f\textit{yrir} lúa sak\textit{ir}}{\Afootnote{÷  L2404.}}. K\textit{óngr} m\textit{ælir} þá  \edtext{manto vilja}{\Afootnote{villtu  L679.}} þyggja góþ\textit{ar} \edtext{sæmd\textit{ir}}{\Afootnote{gjafir af oss  L679.}} o\textit{c} dry\textit{ck}jo v\textit{or}a e\textit{ck}i síþr ||\ledouternote{129v} e\textit{n} \edtext{aþr\textit{ir}}{\Afootnote{÷  L679.}}, þ\textit{ei}r  \edtext{eþ}{\Afootnote{sem  L679, L1572, ÷  L2404.}} o\textit{ss} heimsækja. Hróm\textit{undr} qveþst  \edtext{góþ\textit{om} sæmd\textit{om} aldrei}{\Afootnote{alldrei godum sæmdum  L679.}} neita, e\textit{n} þó seigist  \edtext{h\textit{ann}}{\Afootnote{+ fijrst  L679.}} vilja vita eyrindi þ\textit{ei}rra  m\textit{anna} er
h\textit{ann} hafi  \edtext{t\textit{il} sín sendt}{\Afootnote{sendt til sijn  L679.}} \edtext{f\textit{yrir} stutto}{\Afootnote{÷  L2404.}}. K\textit{óngr} s\textit{egir} þ\textit{at} v\textit{ar} þ\textit{ar}
undir  fal\textit{it}, a\textit{t} þú villd\textit{ir} f\textit{ar}a
m\textit{eþ} o\textit{ss} t\textit{il} Svía r\textit{ic}is,
o\textit{c} h\textit{er}ja á Hálfd\textit{án} K\textit{óng} \edtext{Enga eþr litla}{\Afootnote{litla eda eingva  L679.}} sỏk á e\textit{c} 
v\textit{it}  \edtext{Hálfd\textit{án} }{\Afootnote{hann L679.}} 
s\textit{egir} Hróm\textit{undr} o\textit{c} st\textit{ær}ri sỏk á e\textit{c} v\textit{it} yþr, f\textit{yrir} last o\textit{c} \edtext{álýgi}{\Afootnote{á mæli ligi  L1572.}}, s\textit{em} þér haf\textit{it} bæþi mér  veitt o\textit{c}
a\textit{t} \edtext{hyllst}{\Afootnote{helst  L679, hillt  L1572.}}, þyki mér nú o\textit{c} betst, a\textit{t} þér vyrþ\textit{it} yþ\textit{ar} trúnaþ\textit{ar}m\textit{enn}  Bíld o\textit{c} Vola fram\textit{ar} e\textit{n} m\textit{ic}, þ\textit{ví} a\textit{t}
þ\textit{ei}r h\textit{afa} t\textit{il} þ\textit{ess} un\textit{n}\textit{it} af yþr,  \edtext{þ\textit{ar} eþ}{\Afootnote{er  L2404, þar  L1572.}} þ\textit{ei}r hafa,  \edtext{hv\textit{or} í sin\textit{n} st\textit{at}}{\Afootnote{sva forkunnar vel  L2404.}}, fægt  \edtext{á yþr eyron}{\Afootnote{eýru yþar  L2404.}}.  Þá t\textit{óc} Svanh\textit{vít} t\textit{il} orþa,  \edtext{o\textit{c} m\textit{ælir} t\textit{il}
	K\textit{óngs}}{\Afootnote{÷  L2404.}} sva er  \edtext{þ\textit{ví}}{\Afootnote{þier  L679.}} háttaþ,  \edtext{s\textit{egir} hún}{\Afootnote{÷  L2404.}}, a\textit{t} þér eg\textit{it} Hróm\textit{undi} \edtext{góþo}{\Afootnote{gott  L2404.}} a\textit{t} launa, þ\textit{ar} \edtext{eþ}{\Afootnote{÷  L679.}} h\textit{ann} hefr feng\textit{it} bæþi frænda skaþa, o\textit{c}
fjármisso, líka t\textit{ec}it litla vyrþíng af yþr f\textit{yrir}
st\textit{ar}fa sin\textit{n}, bæþi í haugi Þráins
o\textit{c} opt\textit{ar}, o\textit{c} er yþr nú betst a\textit{t} bjóþa h\textit{onom} sættir 
o\textit{c} góþ  \edtext{boþ}{\Afootnote{sæmdar bod  L679.}} s\textit{em} h\textit{ann} \edtext{má}{\Afootnote{meigi  L679, L1572.}} \edtext{v\textit{el} f\textit{yrir} g\textit{an}gast}{\Afootnote{gangast fijrer  L679,  fyrir gangast  L1572.}},  \edtext{e\textit{n} e\textit{ck}i nein svik eþor yf\textit{ir} hylmíng}{\Afootnote{÷  L1572.}}. 
Sva skal g\textit{jỏr}t, s\textit{egir} K\textit{óngr} stóþ  \edtext{þá Ol\textit{afr} K\textit{óngr}}{\Afootnote{hann þá  L2404, L1572.}} or  sæti síno o\textit{c} t\textit{óc} í hỏnd Hróm\textit{undar} o\textit{c} m\textit{ælir} v\textit{el} má e\textit{c}
biþja þ\textit{ic} \edtext{Hróm\textit{undr}}{\Afootnote{÷  L2404.}} a\textit{t} géfa mér upp reiþi þína, o\textit{c} mæla t\textit{il} vinátto v\textit{it} m\textit{ic}, o\textit{c} beiþst þ\textit{ess} er
þú villt af  \edtext{mér}{\Afootnote{+ þiggja  L679, + fá  L1572.}}, o\textit{c} skal e\textit{c} æv\textit{ar}andi níþíngs nafn
b\textit{er}a,  ef e\textit{c} g\textit{en}g á heit mín,  \edtext{þau  er e\textit{c}}{\Afootnote{÷  L2404.}} \edtext{v\textit{ei}ti}{\Afootnote{heiti  L679, ÷ L2404.}} \edtext{o\textit{c} lofi}{\Afootnote{og lofa  L679, ÷  L2404.}}. Þá m\textit{ælir} Hróm\textit{undr} m\textit{eþ} þ\textit{ví}
a\textit{t} Svanh\textit{vít} \edtext{vill hér t\textit{il} leggja}{\Afootnote{leggr til  L679.}} sín góþ orþ, þá skul\textit{om} v\textit{it} vinátto m\textit{eþ}
o\textit{ck}or  binda, a\textit{t} þ\textit{ví} \edtext{skilyrþi}{\Afootnote{skilorþi  L2404.}} \edtext{fast játto}{\Afootnote{÷  L2404.}}, a\textit{t} e\textit{c} \edtext{t\textit{il} ekta fái}{\Afootnote{fái til egta  L2404.}} systir yþ\textit{ar} Svanh\textit{vít} o\textit{c} vil
e\textit{c} nú heyra hen\textit{n}\textit{ar} jáyrþi, sva  framt  \edtext{s\textit{em}}{\Afootnote{÷  L679, L2404.}} þér viljiþ v\textit{ei}ta. Svanh\textit{vít} \edtext{m\textit{ælir}}{\Afootnote{segir  L1572.}} þ\textit{at} þykist e\textit{c} vita, a\textit{t} br\textit{óþir} min\textit{n} \edtext{mani}{\Afootnote{mun  L679, muni  L2404, L1572.}} \edtext{þ\textit{at} h\textit{it} sama vilja o\textit{c}}{\Afootnote{þetta  L679.}} samþy\textit{ck}ja  \edtext{í þ\textit{ess}o efni}{\Afootnote{÷  L2404.}}, s\textit{em} e\textit{c} sjálf t\textit{il}\edtext{mæli}{\Afootnote{+ i þesso efni  L2404.}}, sérdeilis þ\textit{ar} \edtext{eþ}{\Afootnote{÷  L679, L2404.}} \edtext{h\textit{onom} ríþ||\ledouternote{130r}or sva m\textit{ic}it á, a\textit{t}}{\Afootnote{÷  L2404.}} líf o\textit{c} sæmd h\textit{an}s liggr
v\textit{it}, 
\edtext{hins o\textit{c} an\textit{n}ars}{\Afootnote{oc  L2404.}}, a\textit{t} góþr kostr er f\textit{yrir} hendi, o\textit{c} gyrnist
e\textit{c} \edtext{ei}{\Afootnote{+ annann  L679.}} fremr, e\textit{n} \edtext{þen\textit{n}\textit{an}\textit{n}, a\textit{t}}{\Afootnote{÷  L679, L2404.}} Hróm\textit{undr} sé m\textit{eþ} þ\textit{ess}o sæmdr, ásamt mỏrgo ỏþro góþo \edtext{s\textit{em} þ\textit{ar} a\textit{t} lítr}{\Afootnote{÷  L2404.}}. K\textit{óngr} b\textit{at} þá hyrþm\textit{enn} \edtext{sína}{\Afootnote{+ ad  L1572.}} v\textit{er}a  min\textit{n}uga  \edtext{þ\textit{ei}rr\textit{ar}}{\Afootnote{þessarar  L679, L2404, L1572.}} sátt\textit{ar} gj\textit{ỏr}þ\textit{ar} \edtext{sín á milli}{\Afootnote{÷  L2404, milli sin  L1572.}} \edtext{o\textit{c} Hróm\textit{undar}}{\Afootnote{÷  L2404.}} Fastn\textit{aþ}i þá Hróm\textit{undr} Svanh\textit{vít} \edtext{þ\textit{ar}}{\Afootnote{at  L679, ÷  L1572.}} ỏll\textit{om} v\textit{it} v\textit{er}andi, e\textit{n} sá  skilmáli fylgdi, a\textit{t} h\textit{ann} \edtext{skal}{\Afootnote{skyldi  L2404.}} l\textit{an}dv\textit{ỏr}n \edtext{halda}{\Afootnote{÷  L2404.}} \edtext{o\textit{c}}{\Afootnote{÷  L679, L2404, L1572.}} \edtext{h\textit{afa}}{\Afootnote{÷  L679, L1572.}},  \edtext{þeg⟨ar⟩}{\Afootnote{þá  L679.}} Ol\textit{afr} K\textit{óngr} \edtext{vill o\textit{c} m\textit{eþ} þ\textit{ar}f}{\Afootnote{þarf med  L679, meþ þarf  L2404.}},  
\edtext{e\textit{n} Hróm\textit{undr} aptr þau ljeni}{\Afootnote{÷  L2404.}},  \edtext{er}{\Afootnote{+ hann  L679, L1572, ÷  L2404.}} \edtext{vill o\textit{c} t\textit{il}kall\textit{ar}}{\Afootnote{÷  L2404.}},  \edtext{þ\textit{ar}}{\Afootnote{÷  L679, L2404.}} \edtext{í r\textit{ic}ino}{\Afootnote{÷  L2404.}}. 
Settost nú all\textit{ir} niþr, oc  t\textit{óc}o
a\textit{t} dre\textit{ck}a  \edtext{gott min\textit{n}is ỏl}{\Afootnote{÷  L2404.}}, e\textit{n} Brúþkaup skal bíþa  \edtext{t\textit{il} þ\textit{ess} a\textit{t}}{\Afootnote{þar til  L679, þar til þeir  L2404.}} K\textit{óng}r o\textit{c} Hróm\textit{undr} \edtext{heim aptr}{\Afootnote{÷  L2404.}} \edtext{q\textit{vo}ma}{\Afootnote{kæmi aptr  L2404.}} frá Sví\textit{om}. 
\pend \endnumbering 
\beginnumbering \pstart  
\vspace{5mm}\subsection*{XX kap\textit{ítuli}} 
Strax u\textit{m} m\textit{or}gunin\textit{n},  \edtext{þá}{\Afootnote{er  L679, ÷  L1572.}} \edtext{dagr v\textit{ar}}{\Afootnote{÷  L1572.}}, stóþ  Hrom\textit{undr}  \edtext{fyrstr m\textit{anna} upp}{\Afootnote{upp fijrstr manna  L679, firzr manna á fætr  L2404.}}, o\textit{c} \edtext{heimt\textit{aþ}i}{\Afootnote{heimtar  L1572.}} sam\textit{an} liþ\textit{it},  \edtext{þ\textit{at} s\textit{em}}{\Afootnote{er  L679, sem  L2404.}} Ol\textit{afi} K\textit{óng}i  \edtext{tilheyrþi}{\Afootnote{tillagdi  L679.}}, o\textit{c} b\textit{at} \edtext{þ\textit{at}}{\Afootnote{þá  L679, L2404, L1572.}} búast t\textit{il} h\textit{er}f\textit{er}þ\textit{ar}, urþo nú
m\textit{enn} t\textit{il} í \edtext{fleyra}{\Afootnote{fijrra L679.}} lagi, 
o\textit{c} bjuggost  \edtext{allfiml\textit{e}ga}{\Afootnote{fimliga  L1572.}}. K\textit{óngr} \edtext{s\textit{egir}}{\Afootnote{baþ  L2404.}} \edtext{a\textit{t}}{\Afootnote{÷  L679, L2404.}} \edtext{þ\textit{ei}r skuli}{\Afootnote{menn  L2404.}} f\textit{ar}a  hljóþl\textit{e}ga, Hróm\textit{undr}
m\textit{ælir} v\textit{it} skul\textit{om} \edtext{ríþa}{\Afootnote{halda  L1572.}} \edtext{nú}{\Afootnote{÷  L679, L1572.}} an\textit{n}\textit{an}\textit{n} veg,  \edtext{heldr}{\Afootnote{÷  L679, L1572.}} \edtext{e\textit{n} áþr}{\Afootnote{÷  L679.}}, o\textit{c} \edtext{skul\textit{om}}{\Afootnote{÷  L679, L1572.}} \edtext{halda}{\Afootnote{÷  L1572.}} sun\textit{n}\textit{an} a\textit{t} Svíþjóþ, o\textit{c} \edtext{q\textit{vo}mom}{\Afootnote{koma á  L2404.}} óv\textit{ar}t, ef  \edtext{ské mætti a\textit{t}}{\Afootnote{÷  L679.}} sigr fengist. 
Þeir  \edtext{g\textit{jỏr}a}{\Afootnote{+ svo  L679, L2404.}} s\textit{em} Hróm\textit{undr} f\textit{yrir} sagþi; 
snúa þ\textit{ei}r þá
f\textit{er}þ sin\textit{n}i suþr fyrir  Kjỏl, o\textit{c} q\textit{vo}ma  \edtext{í}{\Afootnote{sunnan ad þvj  L679.}} Svíþjóþ  \edtext{a\textit{t}}{\Afootnote{÷  L679.}} \edtext{óvỏr\textit{om}}{\Afootnote{vorom  L2404.}}, bren\textit{n}a  \edtext{þ\textit{ei}r}{\Afootnote{þar  L679, L2404.}} bæi  u\textit{m} nætr, 
e\textit{n} ríþa t\textit{il} \edtext{skóga}{\Afootnote{skogar  L679, L1572.}} u\textit{m} daga, sva lít\textit{it} t\textit{il}spyrst, 
þ\textit{ar} t\textit{il} eina nótt, a\textit{t} þ\textit{ei}r
q\textit{vo}mo a\textit{t} Uppsỏl\textit{om}, v\textit{ar} þá Hálfd\textit{án} K\textit{óngr} \edtext{í úti\-ském\-\textit{m}o}{\Afootnote{uti i skemmu  L1572.}} m\textit{eþ} no\textit{ck}ra hỏfþíngja, t\textit{óc} þá Hróm\textit{undr} þ\textit{at} \edtext{fyrst t\textit{il} ráþs}{\Afootnote{rad  L679, first ráþz  L2404.}},  
\edtext{a\textit{t}}{\Afootnote{+ hann  L679.}}  \edtext{t\textit{ac}a}{\Afootnote{tok L679.}}  fólki  \edtext{síno}{\Afootnote{fijrst  L679.}} strandhỏgg 
o\textit{c} ||\ledouternote{130v} vist\textit{ir} í utibúr\textit{om},  \edtext{ept\textit{ir} venjo h\textit{er}víkíngja}{\Afootnote{÷  L679, L2404.}};  \edtext{Visso þá ei m\textit{enn} h\textit{an}s fyrri t\textit{il}}{\Afootnote{vissu þá menn hans ei fyrri til  L1572, ÷  L2404.}}; e\textit{n} Hálfd\textit{án} K\textit{óngr} v\textit{ar} þ\textit{ar}
in\textit{n}i f\textit{yrir}, t\textit{óc} þá m\textit{ỏnnom} a\textit{t}
þykja  \edtext{betri}{\Afootnote{batna  L679.}} \edtext{veiþorin\textit{n}}{\Afootnote{veidinn  L679, L1572.}}. 
Hálfd\textit{án} \edtext{K\textit{óngr}}{\Afootnote{÷ L679.}}
 kall\textit{ar} t\textit{il} þ\textit{eir}ra  \edtext{er}{\Afootnote{sem  L679.}} úti  \edtext{stóþo}{\Afootnote{vorú  L679.}},  
 o\textit{c} sp\textit{yr} \edtext{hv\textit{or}j\textit{ir}}{\Afootnote{+ þeir  L679, + þeir menn  L2404.}} \edtext{v\textit{ær}i}{\Afootnote{værú  L679, L2404, L1572.}} \edtext{þ\textit{ei}r m\textit{enn}}{\Afootnote{÷  L2404, þeir  L1572.}} 
 \edtext{s\textit{em}}{\Afootnote{ed  L1572.}} \edtext{óspekt\textit{ir}}{\Afootnote{öspeckt  L679.}} sýndo sér u\textit{m} nætr  \edtext{tíma}{\Afootnote{÷  L679, L1572.}}. Hróm\textit{undr} s\textit{egir} nafn sitt, 
 o\textit{c} þ\textit{at}
m\textit{eþ} a\textit{t} \edtext{Ol\textit{afr} K\textit{óngr} sé þar q\textit{vo}min\textit{n}}{\Afootnote{þar væri kominn Olafr Kóngr  L679.}}.  \edtext{Hálfd\textit{án}}{\Afootnote{+ kóngr  L679.}} heit\textit{ir} á m\textit{enn} sína, o\textit{c} biþr þá  \edtext{v\textit{er}jast}{\Afootnote{berjast  L679.}} o\textit{c} v\textit{er}a  \edtext{orugg\textit{ir}}{\Afootnote{órúggva vard þar hinn hardasti bardægi  L679, örugga  L1572.}}. 
Þeir Svíar  \edtext{brugþo}{\Afootnote{bregda  L679, L2404, L1572.}} \edtext{fljótt}{\Afootnote{skiótt  L1572.}} v\textit{it} o\textit{c} \edtext{lupo}{\Afootnote{hlaupa  L679, L2404.}} t\textit{il} \edtext{vopna}{\Afootnote{+ sinna  L679.}},  o\textit{c} g\textit{jỏr}þo skjaldb\textit{or}g  \edtext{f\textit{yrir}}{\Afootnote{ad  L679.}} K\textit{óng}i sín\textit{om}, v\textit{ar}þ  \edtext{nú 
	þ\textit{ar} hin h\textit{ar}þasta  orrusta o\textit{c}}{\Afootnote{mikid  L679.}} m\textit{an}\textit{n}fall af
N\textit{or}þm\textit{ỏnnom}, t\textit{óc} \edtext{þá}{\Afootnote{nú  L2404.}} Hróm\textit{undr} a\textit{t} þykja  erfidt, a\textit{t} \edtext{sigra}{\Afootnote{eiga vid  L679, vinna  L2404, L1572.}} \edtext{þ\textit{ess}a stofom\textit{enn}}{\Afootnote{÷  L2404.}},  \edtext{þó fá\textit{ir} v\textit{ær}i}{\Afootnote{þá  L2404.}}, þ\textit{ví} þ\textit{ei}r drápo  m\textit{eþ} \edtext{sto\textit{ck}om o\textit{c} stein\textit{om}}{\Afootnote{triám oc stockom  L2404.}},  \edtext{þilj\textit{om} o\textit{c} þv\textit{er}trjám}{\Afootnote{÷  L2404, L1572.}} o\textit{c} ỏllo þ\textit{ví} er f\textit{yrir} v\textit{ar}þ. Þá m\textit{ælir} Hálfd\textit{án} t\textit{il}
Hróm\textit{undar} nú man þér vel  lika a\textit{t} \edtext{á m\textit{ic} snýr
	m\textit{an}\textit{n}fall\textit{it}}{\Afootnote{mannfallinu snijr á mig  L679.}}, f\textit{yrir} \edtext{þ\textit{at}}{\Afootnote{þvj  L679.}}, a\textit{t} viij b\textit{ræþ}r þínir  féllo f\textit{yrir}
o\textit{ss} \edtext{í}{\Afootnote{÷  L1572.}} fyrra vetr. Hróm\textit{undr} m\textit{ælir} ei  \edtext{er}{\Afootnote{+ eg  L679.}} u\textit{m} skaþa  þ\textit{an}\textit{n} a\textit{t} fást,
þ\textit{ví} e\textit{ck}i hrósaþi Helgi  \edtext{lengi}{\Afootnote{laungum sigri  L679, L2404.}} \edtext{ept\textit{ir} þ\textit{at} þ\textit{ei}r 
	féllo}{\Afootnote{÷  L2404.}},  \edtext{stór\textit{om} heiþri}{\Afootnote{÷  L679, L2404.}}. Ol\textit{afr} K\textit{óngr} b\textit{at} þá  \edtext{m\textit{enn} sína}{\Afootnote{sijna menn  L679.}} v\textit{el} duga.  Hróm\textit{undr} \edtext{f\textit{yrir} bauþ}{\Afootnote{bannaþi  L2404.}} all\textit{an} hávaþa, sva b\textit{or}g\textit{ar} m\textit{enn} \edtext{yrþo  lítt v\textit{it} v\textit{ar}ir}{\Afootnote{irdi eij varir vid  L679.}}. 
V\textit{or}o þá ỏll \edtext{þau}{\Afootnote{÷  L679, L2404.}} \edtext{tré}{\Afootnote{+ Svija  L2404.}} t\textit{ec}in\textit{n} o\textit{c} \edtext{f\textit{or}v\textit{or}uþ}{\Afootnote{geimd  L1572.}} s\textit{em} þ\textit{ei}r  \edtext{Sví\textit{ar}}{\Afootnote{÷  L2404.}} útkỏstuþu, t\textit{óc}o þá a\textit{t} þrjóta v\textit{er}jor o\textit{c} skeyti,  
\edtext{t\textit{óc}o þá N\textit{or}þm\textit{enn} a\textit{t} hỏggva stór\textit{om}}{\Afootnote{Norþmenn sókto hart fram  L2404.}}, þ\textit{ar} t\textit{il} \edtext{a\textit{t}}{\Afootnote{÷  L679, L2404.}} all\textit{ir} vóro  \edtext{falln\textit{ir}}{\Afootnote{farnir  L2404.}}, 
nema Hálfd\textit{án} ein\textit{n}. 
v\textit{ar}þist h\textit{ann} þá
\edtext{af}{\Afootnote{÷  L2404.}} \edtext{st\textit{ær}sto  
	príþi m\textit{eþ}}{\Afootnote{÷  L679, allfrækilega  L2404.}} 
\edtext{betsta}{\Afootnote{mesta  L679, L1572, ÷  L2404.}}
\edtext{drengskap}{\Afootnote{÷  L2404.}}, 
sva engin\textit{n} q\textit{vo}m á h\textit{an}n  eino hỏggi. 
J þ\textit{ess}o bili  
\edtext{sjá n\textit{or}þm\textit{enn}, a\textit{t}}{\Afootnote{÷  L679.}} \edtext{m\textit{aþr} q\textit{vo}m}{\Afootnote{kom eirn madr  L679.}} hlaupandi or \edtext{b\textit{or}gin\textit{n}i t\textit{il} ském\textit{m}un\textit{n}\textit{ar}}{\Afootnote{höllinni  L2404.}}, m\textit{eþ} \edtext{náttklæþi}{\Afootnote{náttfelld  L679.}} \edtext{K\textit{ón}gsins}{\Afootnote{kongs  L2404.}}.  Þ\textit{ei}r segja Hróm\textit{undi}
e\textit{n} h\textit{ann} biþr þá draga  \edtext{s\textit{em}}{\Afootnote{sig  L679, L2404, L1572.}} í hlé; m\textit{aþr} þ\textit{ess}i  ||\ledouternote{131r} ljóp a\textit{t} \edtext{dyrun\textit{om}}{\Afootnote{veggnum  L679.}}, o\textit{c} sá vígs u\textit{m} m\textit{er}ki,  \edtext{o\textit{c}}{\Afootnote{÷  L2404.}} þóktist m\textit{ann}avís  v\textit{er}þa; h\textit{ann}
sn\textit{ar}ast fljótt v\textit{it} o\textit{c} vill t\textit{il}
b\textit{ac}a  \edtext{o\textit{c} snýr heim aptor}{\Afootnote{÷  L679.}}. 
Þetta sj\textit{er} Hróm\textit{undr} o\textit{c} \edtext{leypr ept\textit{ir} h\textit{onom} o\textit{c}}{\Afootnote{÷  L679.}} hỏggr t\textit{il} h\textit{an}s,  hỏgg\textit{it} q\textit{vo}m  \edtext{á}{\Afootnote{fyrir  L1572.}} \edtext{apt\textit{an} v\textit{er}þ\textit{an} spjald}{\Afootnote{÷  L2404.}} hriggin\textit{n} o\textit{c} sneid ofan
lend\textit{ar}n\textit{ar} 
\edtext{o\textit{c} þjóin\textit{n}, sva a\textit{t} þau}{\Afootnote{svo þjófinn  L679.}} 
féllo á  \edtext{j\textit{ỏr}þina}{\Afootnote{jórd  L679.}}, 
féll  h\textit{ann} þá  \edtext{o\textit{c} líka}{\Afootnote{daudr  L679.}}; 
t\textit{óc} þá Hróm\textit{undr} a\textit{t} sækja  
\edtext{í meirsta ákafa}{\Afootnote{÷  L679, í ákafa  L1572.}} a\textit{t} \edtext{Hálfd\textit{án}}{\Afootnote{+ kongi i mesta ákafa  L679.}} e\textit{n} hv\textit{or}t þ\textit{at} hỏgg s\textit{em} Hróm\textit{undr} hjó t\textit{il} \edtext{K\textit{óng}s}{\Afootnote{Halfdanar  L679.}},  \edtext{þat}{\Afootnote{÷  L1572.}} q\textit{vo}m flatt á h\textit{er}klæþi  \edtext{K\textit{óng}s}{\Afootnote{hans  L679, L2404.}}, e\textit{n} h\textit{ann} fe\textit{ck} e\textit{ck}ert sár. Þetta  sj\textit{er} Hróm\textit{undr} \edtext{a\textit{t} sva vill lengi g\textit{an}ga}{\Afootnote{og  L2404.}}, kast\textit{ar} \edtext{h\textit{ann} þá}{\Afootnote{÷  L2404.}} Mistilteini,  \edtext{e\textit{n}}{\Afootnote{oc  L2404.}} grípr  \edtext{aptr}{\Afootnote{÷  L2404.}} kylfo sína, o\textit{c} lemr m\textit{eþ} hen\textit{n}i; 
Hálfd\textit{án} v\textit{ar}þist  \edtext{v\textit{el} o\textit{c} ágjætl\textit{e}ga}{\Afootnote{ágjæta vel  L679.}} e\textit{n} u\textit{m} stund, þ\textit{ar} t\textit{il} a\textit{t}
h\textit{ann} mæþdist, skildi Hróm\textit{undr} ei fyrr v\textit{it}
K\textit{óng} e\textit{n} h\textit{ann} hafþi  \edtext{lam\textit{it} í sundr}{\Afootnote{brotid  L679, lamat  L2404.}} í h\textit{onom} \edtext{ỏll}{\Afootnote{hvórt  L679, L2404.}} \edtext{beinin}{\Afootnote{bein  L679, L2404.}}, lauk sva Hálfd\textit{án} \edtext{æfi}{\Afootnote{hreysti  L679.}} sin\textit{n}i m\textit{eþ} \edtext{st\textit{or}ri hreysti}{\Afootnote{miklum dreing skap  L679.}}. Hróm\textit{undr} s\textit{egir} h\textit{ann} \edtext{þ\textit{ann}}{\Afootnote{hinn  L679.}} \edtext{fræknasta}{\Afootnote{hraustasta  L1572.}} m\textit{ann} v\textit{er}it h\textit{afa}, o\textit{c}
óvin\textit{n}\textit{an}l\textit{e}g\textit{an}, hefþi h\textit{ann} í
rúmri h\textit{er}s\-fyl\-kín\-go  \edtext{staþit}{\Afootnote{verid  L679.}}. 
\pend \endnumbering 
\beginnumbering \pstart 
\vspace{5mm} \subsection*{XXI kap\textit{ítuli}} 
 Ept\textit{ir} þ\textit{at} \edtext{a\textit{t}}{\Afootnote{÷  L679.}} Hálfd\textit{án} v\textit{ar} dauþr, t\textit{ac}a  \edtext{þ\textit{ei}r}{\Afootnote{menn  L679.}} h\textit{ann} af re\textit{ck}jon\textit{n}i í  hv\textit{or}ri h\textit{ann} \edtext{v\textit{ar}ist hafþi}{\Afootnote{vardist  L679, lagst hafdi  L1572.}}, o\textit{c} v\textit{ei}ta h\textit{onom} Konúngl\textit{e}ga umbúþ,  \edtext{o\textit{c} settost niþr a\textit{t} t\textit{ac}a hressíng}{\Afootnote{÷  L2404.}}, v\textit{ar} þá vaknat  \edtext{þ\textit{at} liþ}{\Afootnote{lid þad  L679.}} \edtext{s\textit{em}}{\Afootnote{er  L679, L1572.}} í  b\textit{or}gin\textit{n}i v\textit{ar}, fengo
þ\textit{ei}r  \edtext{þá}{\Afootnote{nú  L2404.}} a\textit{t} vita, a\textit{t} K\textit{óngr} v\textit{ar}
fallin\textit{n}, urþo  þá Sví\textit{ar} ráþalus\textit{ir},  \edtext{s\textit{em} v\textit{an}t er}{\Afootnote{÷  L2404.}} \edtext{/: a\textit{t} dauþr er jafnan  hỏfuþlaus
	h\textit{er} :/}{\Afootnote{÷  L2404, því daud er hófudlaus her  L1572.}}\footnote{The entire phrase is in  parenthesis in B11109, but not in L679 and L1572.} 
g\textit{en}go \edtext{þ\textit{ei}r}{\Afootnote{÷  L679.}} þá Hróm\textit{undr} o\textit{c} \edtext{m\textit{enn} h\textit{an}s}{\Afootnote{hans menn  L679.}} í b\textit{or}gina,  o\textit{c} \edtext{rænto}{\Afootnote{÷  L2404.}} \edtext{þ\textit{ei}m}{\Afootnote{÷  L2404, þar  L1572.}} \edtext{grip\textit{om} o\textit{c} gj\textit{ỏr}fil\textit{e}g\textit{om}}{\Afootnote{÷  L679, L2404, L1572.}} \edtext{lut\textit{om}}{\Afootnote{því  L2404, L1572.}} \edtext{er þ\textit{ei}r vildo}{\Afootnote{sem ágætaz var af gersemom  L2404.}},  \edtext{o\textit{c} höfþo mest\textit{ar} mætr á}{\Afootnote{÷  L2404, L1572.}}, e\textit{n} \edtext{m\textit{ỏnnom} gáfo þ\textit{ei}r friþ o\textit{c}
	frelsi}{\Afootnote{gáfo mönnum griþ  L2404.}};  \edtext{t\textit{óc}o þ\textit{ei}r þ\textit{ar} ofrgnægþ
	fjár}{\Afootnote{÷  L2404.}} \edtext{á}{\Afootnote{og  L679, ÷  L2404, L1572.}} \edtext{f\textit{ar}arskjóta sína}{\Afootnote{÷  L2404, L1572.}},  \edtext{o\textit{c} halda  sva}{\Afootnote{heldo þeir síþan  L2404, halda sidan  L1572.}} heim aptr  \edtext{í Hálogal\textit{an}d}{\Afootnote{til Hálogalandz  L2404.}} \edtext{m\textit{eþ} heiþri o\textit{c} vyrþíngo}{\Afootnote{÷  L2404.}} ||\ledouternote{131v} \edtext{o\textit{c} sjálfsögþ\textit{om} sigri}{\Afootnote{÷  L2404, L1572.}}; setjast nú  \edtext{m\textit{enn}}{\Afootnote{÷  L1572.}} u\textit{m} kyrt; t\textit{ec}or þá  Ol\textit{afr}
K\textit{óngr} a\textit{t} draga t\textit{il} faung,  \edtext{o\textit{c} efna}{\Afootnote{÷  L679.}} t\textit{il} v\textit{ei}tslo, lætr h\textit{ann} \edtext{þá}{\Afootnote{÷  L1572.}} \edtext{sam\textit{an} kalla}{\Afootnote{kalla saman  L679.}} Múga o\textit{c} m\textit{ar}gm\textit{enn}i;  \edtext{bændr o\textit{c} hỏfþíngja}{\Afootnote{÷  L1572.}},  \edtext{o\textit{c} hv\textit{or}jo nafni s\textit{em}
	nefnd\textit{ir} v\textit{or}o}{\Afootnote{÷  L2404, L1572.}}; lét  \edtext{þá}{\Afootnote{÷  L1572.}} Hróm\textit{undr} sækja  Gr\textit{eip} F\textit{ỏþr}
sin\textit{n}, o\textit{c} alla sína náúnga  
\edtext{o\textit{c} vini}{\Afootnote{÷  L1572.}};  \edtext{lét þá Sv\textit{an}hvít s\textit{egja}
	Hróm\textit{undi} a\textit{t}}{\Afootnote{Svanhvit oc Rómund  L2404.}} \edtext{láta}{\Afootnote{÷  L679, let oc  L2404.}} sækja Hagal o\textit{c} Haddýsi,  \edtext{v\textit{ar} þá sendt ept\textit{ir} þ\textit{ei}m,
	v\textit{ar} þ\textit{at} jafn sn\textit{ar}t a\textit{t} þau
	q\textit{vo}mo}{\Afootnote{÷  L2404.}}, hafþi Hróm\textit{undr} lát\textit{it}
h\textit{an}dt\textit{ac}a Blind, 
o\textit{c} m\textit{eþ} sér  \edtext{flytja}{\Afootnote{flutt  L2404.}},  \edtext{n\textit{ær} h\textit{ann} heim f\textit{er}þaþist}{\Afootnote{÷  L679, L2404, L1572.}} \edtext{or Svíþjóþ}{\Afootnote{÷  L679, L2404.}}; 
\edtext{v\textit{ar}þ}{\Afootnote{var L679.}} þ\textit{at} a\textit{t} mála lyktum  \edtext{ept\textit{ir}}{\Afootnote{af  L679.}} vilja Hagals o\textit{c} \edtext{kono h\textit{an}s}{\Afootnote{Haddjsar  L679, L2404.}}, a\textit{t} Bl\textit{indr} v\textit{ar} festr á gálga,
o\textit{c} \edtext{lét}{\Afootnote{+ hann  L679, L2404, L1572.}} þ\textit{ar} líf sitt,  \edtext{ept\textit{ir} v\textit{er}þugleik\textit{om},
	o\textit{c} drauma u\textit{m}róti}{\Afootnote{÷  L2404.}};  v\textit{ar}þ þ\textit{at} a\textit{t} mál t\textit{ac}i:
a\textit{t} \edtext{e\textit{ck}i}{\Afootnote{efla  L679.}} skyldi drauma heill meþ  heimam\textit{ỏnnom}, o\textit{c}
f\textit{or} sva m\textit{eþ} Hálfd\textit{án} o\textit{c} Bl\textit{indi} \edtext{a\textit{t} K\textit{óng}i v\textit{ar} þenkt,  e\textit{n} Bl\textit{indr} hreppti}{\Afootnote{÷  L1572.}}, lykt\textit{ar} \edtext{sva hér}{\Afootnote{hier svo  L679, L2404, L1572.}} \edtext{frá}{\Afootnote{af  L1572.}} h\textit{ono}m a\textit{t} segja, e\textit{n} h\textit{an}s 
rétta nafn v\textit{ar},  \edtext{a\textit{t} h\textit{ann} h\textit{ie}t}{\Afootnote{÷  L679, at heita  L2404.}} \edtext{Baniss}{\Afootnote{Bannvijs  L679.}}, e\textit{n} Blindr v\textit{ar} h\textit{ann} kallaþr
f\textit{yrir} þá skuld, a\textit{t} h\textit{ann} 
\edtext{v\textit{ar}þ}{\Afootnote{var L679, L1572.}}
 blindaþr af
K\textit{óng}i m\textit{eþ} útþýþíngo drauma
sin\textit{n}a, sva o\textit{c} líka m\textit{eþ} ofsjón\textit{om} o\textit{c} brỏgþ\textit{om} þ\textit{eir}ra Hagals o\textit{c}
Haddýs\textit{ar};  \edtext{er hér}{\Afootnote{enn hier er  L679, ÷  L2404.}} \edtext{nú  frá a\textit{t} víkja o\textit{c}
	t\textit{il} hins er}{\Afootnote{víkr nú til  L2404.}} \edtext{frá}{\Afootnote{fijrr  L679, ÷  L2404.}} \edtext{v\textit{ar}}{\Afootnote{+ frá L679, ÷  L2404.}} \edtext{h\textit{or}f\textit{it}}{\Afootnote{vekid  L679, ÷  L2404.}},  \edtext{a\textit{t} m\textit{enn} settost}{\Afootnote{÷  L2404.}} \edtext{t\textit{il}}{\Afootnote{undir  L679, L1572, ÷  L2404.}} \edtext{dry\textit{ck}jo o\textit{c}}{\Afootnote{÷  L2404.}} \edtext{vyrþugl\textit{e}gr\textit{ar}}{\Afootnote{virduglega  L679, L1572, ÷  L2404.}} \edtext{v\textit{ei}tslo}{\Afootnote{verslunar  L2404.}},  \edtext{m\textit{eþ} hv\textit{or}skyns gleþi 
	o\textit{c}}{\Afootnote{oc var hon haldin med stæþsto rausn oc hverkyns  L2404, ÷  L1572.}} \edtext{v\textit{er}ald\textit{ar}}{\Afootnote{÷  L679, L1572.}} \edtext{skémton}{\Afootnote{÷  L1572.}}, v\textit{or}o  
\edtext{þ\textit{ar}}{\Afootnote{÷ L679,  L2404, L1572.}} þau Hagall o\textit{c} \edtext{Haþd\textit{ýs}}{\Afootnote{+ þar  L1572.}} í  st\textit{ær}sto vyrþíngo; skip\textit{aþ}i Ol\textit{afr}
K\textit{óngr} an\textit{n}\textit{an}\textit{n} be\textit{ck} o\textit{c}
Hróm\textit{undr} \edtext{n\textit{ær}st}{\Afootnote{næstr  L1572.}} h\textit{onom}, o\textit{c} sva þ\textit{ei}r helsto hỏfþíngj\textit{ar},
e\textit{n} \edtext{hin\textit{n}}{\Afootnote{÷  L2404.}} an\textit{n}an\textit{n} \edtext{sat}{\Afootnote{÷  L679, L2404.}} \edtext{Gr\textit{eipr} o\textit{c} Hagall}{\Afootnote{Hagall og Greipr  L679.}} o\textit{c} \edtext{sva}{\Afootnote{÷  L679, L2404.}} vin\textit{ir} \edtext{Greips}{\Afootnote{hans  L679.}} \edtext{h\textit{onom} ásamt}{\Afootnote{og heima menn  L679, ÷  L1572.}};  \edtext{all\textit{ar} Frúr o\textit{c} Ýngismey\textit{ar}
	sáto}{\Afootnote{sáto allar frúr oc Jngismeyar  L2404, allar meiar og frúr sátu  L1572.}} á  \edtext{stól\textit{om}, s\textit{em} gyllt\textit{ir}}{\Afootnote{forgilltom stólom  L2404.}} ||\ledouternote{132r} oc  \edtext{sylfraþ\textit{ir} v\textit{or}o}{\Afootnote{silfr búnum  L679, L2404.}}; e\textit{n} í  \edtext{miþri v\textit{ei}tslon\textit{n}i}{\Afootnote{miþia veitslo  L2404.}} qvaþdi Hróm\textit{undr} sér hljóþs, o\textit{c} b\textit{at} Ol\textit{af} K\textit{óng}
a\textit{t} \edtext{staþfesta}{\Afootnote{festa  L679.}} giptomál sítt  \edtext{v\textit{it}}{\Afootnote{med  L2404.}} Svanhv\textit{ít} m\textit{eþ} fullkomno Jáyrþi o\textit{c} h\textit{an}dsỏl\textit{om}. 
Stóþ þá Ol\textit{afr} K\textit{óngr} upp, o\textit{c} h\textit{an}dsal\textit{aþ}i Ektamál
o\textit{c} leyfi t\textit{il} h\textit{an}da Hróm\textit{undar} t\textit{il} systr sin\textit{n}\textit{ar} Svanh\textit{vítar}
ept\textit{ir} l\textit{an}dsins venjo o\textit{c} lỏg\textit{om}.  \edtext{K\textit{óngr}}{\Afootnote{Olafur kongr  L679.}} gaf  Hróm\textit{undi} hálf\textit{an} p\textit{ar}t
síns r\textit{ic}is, e\textit{n} Greipi F\textit{ỏþr}
h\textit{an}s gaf han\textit{n} Hers\textit{ir}s nafn  \edtext{o\textit{c} léni}{\Afootnote{÷  L679.}}. Hróm\textit{undr} gaf Hagali  \edtext{k\textit{ar}li}{\Afootnote{÷  L679.}} o\textit{c} kono h\textit{an}s Umsjón\textit{ar} ráþ náqvæmosto  
\edtext{ráþag\textit{jỏr}þ\textit{ar}}{\Afootnote{÷  L679.}} í þ\textit{ví} s\textit{em} \edtext{meirst}{\Afootnote{+ vid  L679.}} þurfti, o\textit{c} vit sitt eg\textit{it} b\textit{or}þ sitja  \edtext{þ\textit{aþ}an af}{\Afootnote{÷  L679.}}, sva  lengi  \edtext{s\textit{em}}{\Afootnote{+ þau  L2404.}} \edtext{lifþo}{\Afootnote{þijrfti  L679.}}. 
\edtext{Stóþ sva þ\textit{ess}i v\textit{ei}tsla m\textit{eþ} st\textit{ær}sta sóma}{\Afootnote{÷  L1572.}},  \edtext{oc}{\Afootnote{ad  L1572.}} \edtext{end\textit{aþ}i}{\Afootnote{endadist  L679, L2404, endadri veitslunni  L1572.}} \edtext{m\textit{eþ} friþi o\textit{c} góþ\textit{om}}{\Afootnote{÷  L1572.}} \edtext{ról\textit{e}gheit\textit{om}}{\Afootnote{fagnaþi  L2404, ÷  L1572.}},  \edtext{o\textit{c}}{\Afootnote{÷  L1572.}} v\textit{or}o all\textit{ir} útleystir  \edtext{m\textit{eþ} vyrþíng o\textit{c} góþ\textit{om}}{\Afootnote{godum og virduglegum  L679, virþolegom  L2404, med godum  L1572.}} \edtext{gjỏf\textit{om}}{\Afootnote{+ og  L679.}};  \edtext{t\textit{óc}ost}{\Afootnote{+ nú  L679.}} \edtext{bráþt}{\Afootnote{+ og godar astir  L679.}} \edtext{m\textit{eþ}}{\Afootnote{+ þeim  L679.}} Hróm\textit{undi} o\textit{c} Svanh\textit{vít} \edtext{góþ\textit{ar} ást\textit{ir}}{\Afootnote{÷  L679.}},  \edtext{ólo þau}{\Afootnote{þau átto  L2404.}} bæþi  \edtext{syni}{\Afootnote{sono  L2404.}} o\textit{c} dætr  \edtext{⸌þeirra Synir Björnólfr oc Hróálfr⸍}{\Afootnote{÷  L679, L2404, L1572.}}\footnote{An asterisk follows the word \enquote{dættur} in B11109 and it refers to the footnote in the manuscript, which contains the addition \enquote{þeirra synir Bjórnólfr oc Hróálfr}. In L1572 the asterisk appears after \enquote{ritud} and the footnote reads \enquote{al: synir þeira Biörnólfr og Hróalfr}.}  \edtext{o\textit{c} e\textit{ro}}{\Afootnote{erú L679, þo  L2404.}} \edtext{hér þ\textit{eir}ra nỏfn}{\Afootnote{þeira nófn hier  L679, L1572, hér sé frá atgiörþom þeira  L2404.}} \edtext{ei}{\Afootnote{÷  L2404.}} \edtext{ritot}{\Afootnote{greind  L679.}},  \edtext{sýnist sva s\textit{em}}{\Afootnote{÷  L2404.}} \edtext{a\textit{t}}{\Afootnote{÷  L679, L2404, L1572.}} \edtext{þau}{\Afootnote{÷  L2404, L1572.}} \edtext{sum  mani}{\Afootnote{hafi sum  L679, ÷  L2404, sum af þeim muni  L1572.}} \edtext{í b\textit{ar}næsko dá\textit{it}}{\Afootnote{÷  L2404.}} \edtext{h\textit{afa}}{\Afootnote{÷  L679, L2404.}},  \edtext{e\textit{n} af sum\textit{om} q\textit{vo}mo}{\Afootnote{oc voru þau in mestu  L2404.}} m\textit{ic}il m\textit{enn}i  \edtext{s\textit{em} í ỏþr\textit{om} Sỏg\textit{om}	fráskýr\textit{ir}}{\Afootnote{oc storar ættir frá þeim komnar þeirra sijnir vóru þeir Biörnólfur oc Hróaldr, Biörnólfr var faþir Arnar faþir Ingólfs landnámamanns, enn son Hrólálfs var Harþmar faþir Hiörleifs er Hiörleifs höfþi er vit kendur  L2404.}}.  
\edtext{Lúk\textit{om}}{\Afootnote{oc lýcor hér  L2404.}} \edtext{v\textit{ær} sva þ\textit{ess}\textit{ar}i}{\Afootnote{÷  L2404.}} Sỏgo  \edtext{af Hróm\textit{undi} Greipssyni}{\Afootnote{Hrómundar Greipssonar  L2404.}}.  
\edtext{Hafi sá góþa þỏkk  s\textit{em} las, e\textit{n} heiþr s\textit{em} hlýþdo, e\textit{n} h\textit{ann} er skrifaþi lof o\textit{c} vinfengi góþra m\textit{anna}; e\textit{n}	all\textit{ir} sameginl\textit{e}ga tím\textit{an}l\textit{e}ga  o\textit{c} eilífa gleþi fyrir utan enda}{\Afootnote{÷  L679, L2404, L1572.}}. \pend \endnumbering 
%Finished proofreading of the entire text 24.05.2022